### 2013-09-23 nengo.tex
First draft of paper

\documentclass{frontiersSCNS}

\usepackage{url,lineno}
% \linenumbers  % TODO uncomment!

\graphicspath{{figures/}}

\copyrightyear{2013}
\pubyear{2013}

\def\journal{Neuroinformatics}
\def\DOI{}
\def\articleType{Methods Article}
\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Bekolay {et~al.}}
\def\Authors{Trevor Bekolay\,$^{1,*}$, James Bergstra\,$^{1}$,
  Xuan Choo\,$^{1}$, Travis DeWolf\,$^{1}$, Eric Hunsberger\,$^{1}$,
  Daniel Rasmussen\,$^{1}$, Terrence C. Stewart\,$^{1}$
  and Chris Eliasmith\,$^{1}$}
\def\Address{$^{1}$Computational Neuroscience Research Group,
  Center for Theoretical Neuroscience,
  University of Waterloo, Waterloo, ON, Canada}
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name
% and city zip code) and email of the corresponding author
\def\corrAuthor{Trevor Bekolay}
\def\corrAddress{David R. Cheriton School of Computer Science,
  University of Waterloo, 200 University Avenue West,
  Waterloo, ON, N2L 3G1, Canada}
\def\corrEmail{tbekolay@uwaterloo.ca}

\begin{document}
\onecolumn
\firstpage{1}

\title[Nengo in Python]{Reimplementing Nengo in Python for speed and simplicity}
\author[\firstAuthorLast ]{\Authors}
\address{}
\correspondance{}
\extraAuth{}
%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2,
%  Organization X2, Street X2, City X2 , State XX2
%  (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}
\topic{Python in Neuroscience II}

\maketitle

\begin{abstract}
%Maximum: 2000 characters
  Nengo is a software package for creating and simulating
  large-scale neural models built with
  the Neural Engineering Framework
  and the Semantic Pointer Architecture.
  Nengo was recently used to build Spaun, the largest
  functional brain model to date.
  However, this version of Nengo simulated Spaun
  9000 times slower than real time,
  and required modellers to know
  both Java and Python to be productive.
  In this paper, we describe a Python reimplementation of Nengo
  that leverages scientific Python packages
  to produce the same large-scale neural models
  with an order of magnitude less code.
  The model definition and simulation steps
  have been decoupled,
  making it possible to run the same model
  on different software simulators
  and different hardware platforms.
  Two software simulators are implemented,
  including an OpenCL simulator
  that is TODO times faster than
  the previous version of Nengo
  for a model computing the circular convolution
  of TODO dimensional vectors.

  \tiny \keyFont{\section{Keywords:} 1 2 3 4 5 (6 7 8)}
\end{abstract}

%max figures+tables: 15
%max legnth: 12000 words
%max pdf length: 12 pages

\section{Introduction}

Nengo is a set of software tools built around
a theoretical framework for building neural models
proposed by Eliasmith \& Anderson \cite{TODO},
and more recently, an architecture
for performing high-level cognitive tasks
in that framework proposed by Eliasmith \cite{TODO}.
Nengo consists of a Java-based simulator,
a Jython-based scripting interface
for common use cases of the simulator,
and a Java-based GUI that
uses both the underlying simulator
and the scripting interface.
This set of tools has been used
to create models of visual attention \cite{TODO},
inductive reasoning \cite{TODO},
and reinforcement learning \cite{TODO},
as well as Spaun, a cognitive agent able
to do these and several other tasks
(\cite{TODO}).
Nengo has been used by dozens of researchers
in labs at the University of Waterloo,
Johns Hopkins University, Yale, Stanford,
and others (TODO: confirm).

Here, we describe a new tool
for creating and simulating NEF-based models
implemented in Python,
which effectively merges two existing tools
into one unified package
that can be extended in
several ways that previous Nengo tools could not.

\subsection{Why reimplement Nengo?}

The original Java simulator was created
with the intention of being a general
neural simulator that included
the methods of the Neural Engineering Framework
(NEF; \cite{TODO}) as one of many different
methods of creating neural models.
However, since creating the simulator in 2007,
it has only been extended minimally,
but has been used extensively for its ability
to create NEF-based models.

The Jython scripting interface,
described in \cite{TODO},
provided a simpler interface
to the NEF-specific portions of the simulator.
However, the Java GUI was developed at the same time
as the Jython scripting interface,
and for that reason primarily
interacts with the simulator directly.
Since large-scale models often require
a scripting language for looping
and parameterization,
this has resulted in a difficult transition
for modellers moving from the
drag-and-drop graphical interface
to the scripting interface.
Additionally, advanced scripts often
contain Jython code that either defines
or manipulates Java objects directly,
meaning that modellers have to know
a certain amount of Python and Java to be productive.

Since Nengo makes it possible to define
very large models,
performance is an important issue.
While Java has good support for threads,
it is difficult to interface with
platform-specific tools such as
graphical processing units (GPUs),
which can quickly perform the types of parallel
computations that these neural models require.
This is due to the
architecture of the Java simulator
and the Java Native Interface,
which results in code that
is difficult to maintain and debug.

The new implementation of Nengo described in this paper
was undertaken with three goals in mind:
\begin{enumerate}
  \item Provide a scripting interface
    similar to the existing Jython interface.
  \item Leverage existing scientific computing tools
    rather than write our own.
  \item Provide the ability to run models with
    different software and hardware simulators.
\end{enumerate}

In order to achieve these goals,
this new version of Nengo uses the Python language
and targets the CPython interpreter,
allowing Nengo to use existing scientific Python
tools for vectorized computations, plotting, and graph manipulation.
As evidence that we have met these goals,
we have implemented two different software simulators,
and have done so with an order of magnitude
less lines of code than the Java simulator
and Jython scripting interface combined.
For the remainder of the paper,
we will use ``Nengo'' to refer to
this new Python rewrite.

\subsection{Neural Engineering Framework (NEF)}

TODO; make sure it meshes with 3.1's operators

\begin{equation} \label{TODO}
  TODO
\end{equation}

%% This should be a paper of its own.

%% \subsection{Semantic Pointer Architecture (SPA)}

%% TODO; should we even include this section?
%% It could probably be a separate paper on its own.

\section{Model definition}

One of the strengths of the NEF
is that it allows neural modellers to think
at a higher level
than is normally associated with neural models.
With the NEF, a modeller decides
what information is being represented
by populations of neurons and how that information
is being transformed through the connections
between populations;
this contrasts with modelling decisions
that are typically thought to be important,
such as the morphology of each neuron,
and the distribution of connection weights
between populations.
Instead, these decisions are made---often
automatically---as a direct result
of the information
being represented and the transformations
being performed on that information.
This situates Nengo as a
``neural compiler'' that can
take a functional model description
and produce a low-level neural model
that can be simulated
with any neural simulation software.

To emphasize Nengo's role as a neural compiler,
Nengo allows models to be defined
in a simulator-agnostic manner,
allowing for functionally identical models to be run
on different software and hardware platforms.

\subsection{Nengo objects} \label{sec:objects}

\begin{figure}
\begin{center}
  Nengo high-level object hierarchy
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{
   Nengo high-level object hierarchy}
\end{figure}

A Nengo model is a collection
of Nengo objects connected together,
and probed in order to collect data
during simulation.
The \texttt{Model} object is primarily a container
for Nengo objects,
but also provides methods
that simplify creating, connecting,
and probing these objects.
Basic Nengo models
can be created solely through interacting
with the \texttt{Model} object.
For more advanced models,
these objects can be instantiated
directly and then added to the model.

There are two types of objects
defined in the main scripting interface.

\paragraph{Ensembles}
The most common is the \texttt{Ensemble},
which is a population of neurons
that represents information
in the form of a real-valued vector.
As described in \eqref{TODO},
this can be done
with any neuron model, $G$.
An ensemble must be given as arguments
its dimensionality
(i.e., the length of the vector it represents),
and an object that describes
the population of neurons.
For example, \texttt{LIF(50, tau\_rc=0.02)}
describes a population
of 50 leaky integrate-and-fire neurons
with an RC time constant of 0.02 seconds.
Neurons are defined in this symbolic manner
so that each simulator can compute
this population's nonlinear function
however it sees fit.

Other attributes of the \texttt{Ensemble},
such as its encoding weights, $\mathbf{e_i}$,
the maximum firing rate of the neurons,
and so on, can also be specified.
This is usually done according
to what is known about
the neurobiology of the area being modelled;
some areas of the brain regularly spike
at rates above 200Hz,
while others rarely spike above 10Hz.
If these attributes are not set,
then they will be randomly selected
from distributions that describe
what is known about pyramidal cells in neocortex.

%% However, not all simulators can support all neuron types,
%% and all of the attributes that can be
%% set on an ensemble.
%% This is particularly the case
%% when interacting with special-purpose hardware.
%% In these cases, simulators should
%% raise appropriate exceptions
%% when unsupported ensembles are encountered.

\paragraph{Nodes}
A \texttt{Node} tracks any information
that is not represented by an ensemble.
In the most common case,
nodes provide input signals
that drive ensembles of sensory neurons.
In a more general sense,
nodes represent the experimental environment
in which a neural model exists.
More practically,
nodes can be used to directly compute
parts of a neural model
for debugging purposes.

A node computes an arbitrary function
of its inputs directly.
Possible input signals include
the simulator timestep,
the decoded output of an ensemble,
or the output of another node.
However, unlike ensembles,
there are no constraints on the type
of function that the node computes.
A node can track any number of variables internally,
and use the state of those variables
when computing its function.
It can interact directly with hardware,
interface with other programs
using shared memory or sockets,
and so on.

Nodes allows many of the special-purpose
routines that are necessary in many real-world
models to be integrated
as core components of a Nengo model.

\paragraph{Connections}

Ensembles and nodes can be connected together
in several ways.
Like other high-level objects,
a connection contains symbolic information
about how two objects are to be connected
together.

The most important type of connection
in Nengo is the \texttt{DecodedConnection}.
This connection implements
the transformation principle equations,
\eqref{TODO}.
In other words, the \texttt{DecodedConnection}
allows ensembles to project
the information they're representing---or
a transformation of that information---to
another ensemble or node.
This functionality is what enables Nengo models
to appear conceptual,
even though the underlying implementation
still connects populations of neurons
together with connection weights.
%TODO: reword? is it important that it's a factored weight matrix

\paragraph{Probes}

Ideally, the entire state of a simulation
would be recorded on each timestep,
and stored with no overhead to the simulator.
In practice, recording data
is a costly part of most simulations,
and often what is simulated must also change
if the signals being probed must be filtered.
For this reason, probes are also
a core component of a Nengo model.

In general, Nengo objects and connections
each have several signals that can be probed.
For example, an ensemble's decoded value,
or the underlying neurons' membrane voltages
can be probed, among other signals.
The underlying logic for the probe
leverages connection logic,
making it easy to probe any part of an object.

Since probes are also described
at a symbolic level,
the underlying implementation
can output probed data in many different formats.
Currently, simulators store probed data
directly in memory,
but adding the ability to store data
in files or to stream probed data
directly to sockets will be added.

\section{Model simulation} \label{sec:simulators}

Once a model has been fully defined,
a simulator can be created to
run that model.
Simulators can be created
that work directly with the
objects described in section \ref{sec:objects}.
However, in the simulators that we have implemented
as reference, the model goes through a
two-stage build process in order to
simulate models as fast as possible.
In the first stage,
the model is ``built;''
i.e., the objects in the model are mapped to a simpler
set of objects that represent the computations
that the simulator must perform.
Those computations are analyzed in order to
schedule them efficiently.
In the second stage,
the simulator performs the computations
that have been scheduled on each timestep.

This two-stage process has enabled us
to implement a more efficient simulator
that is only aware of
the objects that are created
during the build process.
In the future,
new simulators can target either
the high level Nengo objects,
or the low level simulator objects
depending on the constraints of the platform.

\subsection{Reference build process}

\begin{figure}
\begin{center}
  Reference simulator build process
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{
   Reference simulator build process}
\end{figure}

The first stage of the reference implementation
maps all of the Nengo objects
down to a set of signals and operators.

\paragraph{Signals}
A \texttt{Signal} represents any number that
will be used by the simulator.
Each high level Nengo object contains
several signals;
for example, an ensemble contains signals
that represent the high-level input
signal that will be encoded
to input currents (see equation \eqref{TODO}),
and the encoding weights
($\mathbf{e_i}$; see equation \eqref{TODO}).
It also contains a neural population,
which contains signals that represent
input currents
($J(t)$; see equation \eqref{TODO}),
bias currents ($J^{bias}$; \eqref{TODO}),
membrane voltages ($V$; \eqref{TODO}),
and refractory times for each cell.
These signal are symbolic,
and contain only information
about the size of the signal that will be represented,
and optionally an initial value.
Simulators may therefore represent
these signals however they want internally.

In some cases, determining
the initial values of some signals
may involve some computation.
In particular, solving for decoding weights
(equation \eqref{TODO}) requires
solving a least-squares minimization problem.
This computation is done in this stage
of the build process,
as it is computationally intensive,
and may be implemented
differently by some simulators.

An important feature of signals
is that parts of a signal
can be accessed with Python slice syntax,
which results in a view of the original signal.
For example, when connecting to
a subset of a population of neurons,
current would be injected
in a view of the overall input current signal
for that population,
rather than creating a new signal,
or partitioning the population's signal
into smaller chunks.
Grouping related signals together
and exposing them through views
also provides hints to the simulator
that can be used to organize signals
efficiently in memory.

\paragraph{Operators}
Operators represent computations
to be performed on signals on each timestep.
Once the model has been built,
only a small set of mathematical
operations are necessary for simulation.
\begin{enumerate}
  \item \textbf{Reset} assigns a constant value to a signal.
    This is typically used to set a signal to 0
    at the start of a timestep.
  \item \textbf{Copy} assigns the value
    of one signal to another signal.
    This is typically used instead of \textbf{Reset}
    in situations where the signal's value
    at the start of a timestep should
    be the value of another signal.
  \item \textbf{DotInc} increments a signal by
    the dot product of two signals;
    i.e., $Y \leftarrow Y + AB$.
    This is a common operation
    done in many situations,
    such as decoding a signal (see equation \eqref{TODO}).
  \item \textbf{ProdUpdate} scales a signal
    and adds the dot product of two signals to it;
    i.e., $Y \leftarrow BY + AX$.
    This is done when temporally filtering a signal.
  \item \textbf{Nonlinearities} include any
    user-defined function computed by a node,
    and the update equation for each neuron model.
\end{enumerate}

It is important to note that all operators
excepts for the nonlinearities are linear,
and therefore can share an implementation
in an underlying simulator,
allowing those operations to happen in parallel.

Operators are created
by Nengo objects during the build process.
For example, an ensemble creates
a \textbf{Reset} operator to reset
its input signal to 0,
and a \textbf{DotInc} operator
to encode that input signal into
input currents injected into
the underlying population of neurons;
i.e., the \textbf{DotInc} computes
$J = \alpha_i \mathbf{e_i} X$ from equation \eqref{TODO}.

\subsection{Reference simulator}

\begin{figure}
\begin{center}
  Reference simulator simulation (include dependency graph)
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{
   Reference simulator simulation}
\end{figure}

The reference simulator starts by
building NumPy arrays
for every signal created during the build process
described above.
It then builds a graph of the
operators created during the build process,
ordering them such that they can be
executed one after another to get the correct answer.

The ordering is determined
by which signals an operator
sets, increments, reads, and updates.
A signal can only be set or updated once,
but can be read or incremented any number of times.
An update represents the operator that sets
the signal's value at time $t+dt$,
and therefore must be the last operator
that acts on a given signal.
This means that an operator
that updates a signal depends on
all operators that set, increment,
or read that signal.
Reads represent reading the value
of a signal at time $t$,
which is determined by the sets
and increments, and so
an operator that reads a given signal
depends on all operators that
set and increment that signal.
Increments must happen
after a given signal is set,
so an operator that increments a given
signal depends on the operator
that sets that signal, if it exists.

The reference simulator represents
these dependencies by a directed graph,
with edges going from
dependent operators to
the operators on which they depend.
The dependency graph is then topologically sorted,
and that ordering is stored.

The simulator can then be run
for an arbitrary amount of time.
On each timestep,
the simulator goes through each of the
operators in order,
and then makes a copy of any signals
that were probed on this timestep.

\subsection{OpenCL simulator}

OpenCL is a framework for writing
efficient parallel code that runs
on many different platforms,
including modern CPUs and GPUs.
It allows for the kind of parallel execution
that is required to run neural models quickly,
while still being able to run
on multiple platforms.

The OpenCL simulator operates on
the same objects that the reference simulator does;
i.e., it starts with a collection
of signals and operators,
schedules them efficiently,
and runs all operators on each timestep.

However, in order to execute
many operators in parallel,
the OpenCL simulator
organizes computations
differently than the reference simulator.
All of the linear operators are mapped
onto \textbf{MultiProdUpdate} operators,
which perform the general update
$$Y \leftarrow \gamma + \beta Y + AX.$$
It can then schedule many
\textbf{MultiProdUpdate} operations
on a single timestep,
if the dependency graph allows it.

The OpenCL simulator also transparently
attempts to build OpenCL equivalents
for arbitrary Python functions
executed by nodes.
These arbitrary functions
can slow down a simulation significantly,
as data and control has to be passed back
to the Python interpreter
on each timestep.
The OpenCL simulator inspects
the abstract syntax tree
of the function being computed
and substitutes OpenCL equivalents when possible.

\section{Extensibility}

Nengo is designed to be extended
in order to make modelling easier,
and to enable models to be run
on many different platforms.

\subsection{Models}

At model creation time,
everything must be done
in a simulator agnostic manner.
Since ensembles and nodes
are the only objects that exist,
models can become long scripts
that contain many
similar function calls that
create and connect ensembles and nodes.
In order to make modelling scripts
shorter and more expressive,
we have created the \texttt{Network} abstraction.

\paragraph{Networks}
A network is a collection of ensembles and nodes
connected together in a particular way.
It is primarily a way of grouping together
a common set of objects connected in a particular way,
and provides some implicit expressiveness
in the name of the network.

Networks can vary dramatically in size,
and can be composed of other networks.
The \texttt{Integrator} network, for example,
is composed of only one recurrently connected ensemble.
By encapsulating that logic in a network,
the purpose of that ensemble is made explicit.
On the other end of the size spectrum,
the \texttt{BasalGanglia} network,
is composed of five groups of ensembles
connected with several specific functions
that together implement a ``winner take all'' circuit.
Encapsulating that logic
makes this complicated section of code
easy to include in many different models
(see Figure~\ref{TODO}).

\begin{figure}
\begin{center}
  Network code listing
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{
   Network code listing}
\end{figure}

\subsection{Nonlinearities}

Many modellers make use of
different neuron models and
learning rules than are
currently implemented
in the two simulators
in section \ref{sec:simulators}.
Because Nengo endeavors
to compute nonlinear functions efficiently,
these nonlinearities
must be implemented in each simulator,
though they may share the same representation
at the model definition stage.
While this results in simulators
that cannot run all Nengo models,
this cannot be avoided,
and therefore we make this explicit.

In order to implement
a new neuron model or learning rule,
two classes must be created.
The first is a symbolic class
that keeps track of any parameters
that might affect the execution
of the nonlinearity;
for example, for
the leaky integrate-and-fire neuron model,
the refractory time constant,
$\tau^{ref}$, is part of this symbolic class.
An instance of this class
is then used as an argument
to the ensemble constructor.
For learning rules,
an instance of that class
is used as an argument
to functions that connect ensembles.
The second class to be created
is simulator-specific,
and contains the function that will be
executed on each timestep of the simulation.
Individual simulators may have additional
requirements when adding
a new nonlinearity.

\subsection{Model simulation}

The simulator included with Nengo
is provided primarily as a reference
for platform experts designing a simulator
for their platform.
The reference simulator is implemented with
a two-step build process;
this is designed to make
some simulators easier to implement,
as the result of
the first step of the build process
is a set of computations
that most computing devices are able to perform.

However, other simulators---especially
those running on special purpose
hardware---may not be able
to perform all of these computations.
In that case, these simulators
can define their own build process
that takes as input a Nengo model.
This flexibility is possible
due to the decoupling of
the model definition and simulation steps.

\section{Examples and benchmarks}

\subsection{Lorenz attractor network}

Many models in theoretical neuroscience
are based on attractor networks.
The NEF has been used in the past
to implement many different types of
attractor networks \cite{TODO}.
In this example,
we implement the Lorenz chaotic attractor
with a single ensemble
composed of 2000 leaky integrate-and-fire neurons.

\begin{figure}
\begin{center}
  Nengo Lorenz code listing
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{
   Nengo Lorenz code listing}
\end{figure}

In order to compare this Nengo model
with other pieces of neural modelling software,
we have implemented the Lorenz attractor
in PyNN \cite{TODO} in order to simulate it
with the simulators that PyNN supports.

\begin{figure}
\begin{center}
  PyNN Lorenz code listing
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{
   PyNN Lorenz code listing}
\end{figure}

\begin{figure}
\begin{center}
  Lorenz attractor benchmarks
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{
   Lorenz benchmarks}
\end{figure}

\subsection{Circular convolution}

One of the most common operations done
in Nengo models is circular convolution.
Circular convolution is how
the semantic pointer architecture
(SPA; \cite{TODO})
binds two semantic pointers together.
Binding is central to many
of the high level cognitive tasks
that make Spaun unique.
Circular convolution
is also best implemented in a two-layer network,
rather than in a single connection.
This two-layer network construction is simplified
by the existence of the \texttt{CircularConvolution} network.

\begin{figure}
\begin{center}
  Nengo CConv code listing
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{
   Nengo CConv code listing}
\end{figure}

TODO: can we do circconv in PyNN?

\begin{figure}
\begin{center}
  PyNN CConv code listing
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{
   PyNN CConv code listing}
\end{figure}

\begin{figure}
\begin{center}
  CConv benchmarks
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{
   CConv benchmarks}
\end{figure}

\section{Discussion}

Even at its \texttt{0.1} release,
we have met our goals
for this reimplementation of Nengo.
\begin{enumerate}
  \item We have provided a scripting interface
    similar to the Jython interface.
    In many cases, the new scripting interface
    is simpler and more expressive than
    the previous Jython interface.
  \item We have leveraged existing scientific computing tools
    rather than writing our own.
    In the the scripting interface and the reference simulator,
    we use NumPy to simplify array computations.
    In the reference simulator, we use
    NetworkX to perform the topological sort
    of the operator dependency graph.
    In examples, we use Matplotlib
    to plot the results of simulations.
  \item We have provided the ability to run models with
    different software and hardware simulators.
    We have shown that this is possible by
    implementing two software simulators.
\end{enumerate}

While this reimplementation of Nengo
does not completely replace
the existing suite of Java-based Nengo tools,
it is able to cover the vast majority of use cases
with an order of magnitude less
code than the Java-based tools
($\sim$30,000 lines of Java + 10,000 lines of Jython
vs. $\sim$8,000 lines of Python).
We are optimistic that the remaining use cases
can be implemented in this new framework.

TODO: something about documentation, unit tests,
coverage, travis-ci, etc

\subsection{Comparison to similar projects}

There are many other neural simulators
dedicated to building large-scale neural models;
however, Nengo is unique in being built
on a theoretical framework (the NEF, \cite{TODO})
that has already been used to make
large-scale functional brain models.

The most closely related project is PyNN
\cite{TODO},
which provides a high-level scripting
interface similar to the
model creation part of Nengo.
Rather than implementing its
own simulator, PyNN uses existing
simulators, such as Brian \cite{TODO},
NEST \cite{TODO}, and NEURON \cite{TODO}.

The APIs of Nengo and PyNN are similar,
but differ significantly
in how groups of neurons are connected together.
In Nengo, connections commonly describe
the mathematical operation that is performed
through the connection between
two ensembles;
e.g., \texttt{DecodedConnection(A, B,
function=product)} connects ensemble A
to ensemble B, projecting the product of
some dimensions encoded by A to B.
In PyNN, connections commonly describe
features of the connection weight matrix
between two populations;
e.g., \texttt{FixedProbabilityConnector(0.5)}
connects two ensembles together,
with a probability of 0.5
that there will be a connection
between a pair of neurons in the two populations.

As shown in Figure~\ref{TODO},
Nengo is able to simulate many networks
faster than all of PyNN's simulators
(TODO: right?).
This is, in part, because Nengo stores two vectors
whose product is the entire
connection weight matrix between
two ensembles, rather than storing
the entire matrix.

However, PyNN also contains functionality
not currently implemented in Nengo.
PyNN can assign spatial information
to neurons in a population,
which can be used to influence
how those neurons connect to other populations.
PyNN simulates many different learning rules
and neuron types,
and is therefore better suited to
simulate small networks of detailed neuron models
(as would be necessary for modelling
an invertebrate nervous system).

\subsection{Future work}

Nengo is a young project that
has started with a deliberately minimal base
in order to make contributing easier than in
previous implementations.
Additionally, many neural modellers
are already using Python,
as evidenced by the existence of
PyNN, Brian, and Python bindings to other simulators.
Nengo will be able to benefit
from improvements to scientific Python tools.
It is also extendable by modellers with
varying technical skills;
creating a new network is simple,
while creating a new simulator is complex.

By enabling outside contributions,
and providing a user friendly API
and many example models,
we hope to foster a rich community of modellers
contributing networks, nonlinearities,
and simulators to Nengo
and to theoretical neuroscience
as a whole.

\section*{Disclosure/Conflict-of-Interest Statement}

The authors declare that the research was conducted in the absence of
any commercial or financial relationships that could be construed as a
potential conflict of interest.

\section*{Acknowledgement}

We would like to thank
everyone who has contributed
to this reimplementation of Nengo
by providing examples,
unit tests, and bugfixes:
Peter Blouw, Brent Komer, and Bryan Tripp.

\paragraph{Funding\textcolon}
NSERC CRC, CFI, and OIT.

\bibliographystyle{frontiersinSCNS&ENG}
\bibliography{nengo}

\end{document}

### 2013-09-24 nengo.tex
NEF section done, some reorg

@@
 \documentclass{frontiersSCNS}
+
+%%%%% Some customizations
+
+\usepackage{listings}
+\usepackage{inconsolata}
+
+\definecolor{dkgreen}{rgb}{0,0.6,0}
+\definecolor{gray}{rgb}{0.5,0.5,0.5}
+\definecolor{mauve}{rgb}{0.58,0,0.82}
+
+\lstset{
+  language=Python,
+  aboveskip=3mm,
+  belowskip=3mm,
+  showstringspaces=false,
+  columns=flexible,
+  basicstyle={\small\ttfamily},
+  numbers=none,
+  numberstyle=\tiny\color{gray},
+  keywordstyle=\color{blue},
+  commentstyle=\color{dkgreen},
+  stringstyle=\color{mauve},
+  breaklines=true,
+  breakatwhitespace=true,
+  deletendkeywords={filter},
+}
+
+\graphicspath{{figures/}}
+
+%%%%% Frontiers boilerplate
 
@@
 % \linenumbers  % TODO uncomment!
-
-\graphicspath{{figures/}}
 
@@
   University of Waterloo, Waterloo, ON, Canada}
-% The Corresponding Author should be marked with an asterisk
-% Provide the exact contact address (this time including street name
-% and city zip code) and email of the corresponding author
 \def\corrAuthor{Trevor Bekolay}
@@
 
+%%%%% Document begins
+
 \begin{document}
@@
 
-\title[Nengo in Python]{Reimplementing Nengo in Python for speed and simplicity}
+\title[Nengo in Python]{Nengo: A Python tool for
+  building large-scale functional brain models}
 \author[\firstAuthorLast ]{\Authors}
@@
 %Maximum: 2000 characters
-  Nengo is a software package for creating and simulating
-  large-scale neural models built with
-  the Neural Engineering Framework
-  and the Semantic Pointer Architecture.
-  Nengo was recently used to build Spaun, the largest
-  functional brain model to date.
-  However, this version of Nengo simulated Spaun
-  9000 times slower than real time,
-  and required modellers to know
-  both Java and Python to be productive.
-  In this paper, we describe a Python reimplementation of Nengo
-  that leverages scientific Python packages
-  to produce the same large-scale neural models
-  with an order of magnitude less code.
-  The model definition and simulation steps
-  have been decoupled,
-  making it possible to run the same model
-  on different software simulators
-  and different hardware platforms.
-  Two software simulators are implemented,
-  including an OpenCL simulator
-  that is TODO times faster than
-  the previous version of Nengo
-  for a model computing the circular convolution
-  of TODO dimensional vectors.
 
@@
 
-Nengo is a set of software tools built around
-a theoretical framework for building neural models
-proposed by Eliasmith \& Anderson \cite{TODO},
-and more recently, an architecture
-for performing high-level cognitive tasks
-in that framework proposed by Eliasmith \cite{TODO}.
-Nengo consists of a Java-based simulator,
-a Jython-based scripting interface
-for common use cases of the simulator,
-and a Java-based GUI that
-uses both the underlying simulator
-and the scripting interface.
-This set of tools has been used
-to create models of visual attention \cite{TODO},
-inductive reasoning \cite{TODO},
-and reinforcement learning \cite{TODO},
-as well as Spaun, a cognitive agent able
-to do these and several other tasks
-(\cite{TODO}).
-Nengo has been used by dozens of researchers
-in labs at the University of Waterloo,
-Johns Hopkins University, Yale, Stanford,
-and others (TODO: confirm).
-
-Here, we describe a new tool
-for creating and simulating NEF-based models
-implemented in Python,
-which effectively merges two existing tools
-into one unified package
-that can be extended in
-several ways that previous Nengo tools could not.
-
-\subsection{Why reimplement Nengo?}
-
-The original Java simulator was created
+It is often said that neuroscience
+is data-rich but theory-poor.
+Although computational neuroscience
+has grown significantly,
+this phrase continues to be repeated
+due to a lack of satisfying
+theories of the brain.
+Much work has been put into neural simulators
+that attempt to recreate neuroscientific
+data in precise detail,
+but behavior has not emerged
+from data-driven large scale models.
+Similarly, theoretical frameworks
+consisting of a generic algorithm
+replicated many times
+(e.g., the memory-prediction framework \cite{TODO}
+and the pattern recognition theory of mind \cite{TODO})
+have not been shown to solve problems
+that the generic algorithm alone cannot solve.
+
+Nengo is a neural simulator
+based on a theoretical framework proposed
+by Eliasmith \& Anderson \citeyearpar{TODO}
+called the Neural Engineering Framework
+(NEF).
+Nengo was recently used to implement Spaun,
+currently the world's largest functional brain model
+\cite{TODO}.
+Spaun is a network of 2.5 million spiking neurons
+that can perform eight high-level cognitive tasks
+including list memory and inductive reasoning.
+It can perform any of these eight tasks
+at any time by being presented
+the appropriate series of images
+representing the task to be performed
+and the details of that task;
+for example, sequentially presenting images
+containing the characters $A3[1234]$ instructs Spaun
+to remember the list $1234$.
+It then provides its answer by
+sending motor commands to a simulated arm model,
+tracing out digits.
+If asked to recall the remember list,
+it would trace out the digits $1234$.
+
+While the tasks Spaun performs are diverse,
+all tasks use a common set of
+functional modules that we map
+onto the brain areas hypothesized
+to perform those functions.
+Like data-driven large scale models,
+Spaun is able to match experimental data
+across the neuroscientific spectrum,
+from the single cell to the behavioral level.
+However, instead of constructing the model
+based on data and hoping behavior emerges,
+we construct the model
+based on behavior
+and see that the data emerges.
+This is made possible by a theoretical framework
+like the memory-prediction framework \cite{TODO},
+but unlike it and other frameworks,
+the NEF enables Spaun to employ
+many different algorithms
+based on the behavior being performed,
+rather than defining a single algorithm
+and hypothesizing that replicating it
+many times will enable these behaviors.
+
+In implementing Spaun,
+we provided strong support
+that the NEF is beginning to fill the theory void in neuroscience.
+However, in doing so, we pushed Nengo
+to the limits of its ability to construct and simulate
+large-scale neural models.
+In order to continue to evaluate the NEF
+as a theory of the brain
+by building larger and more complicated models,
+we have begun the next generation
+of Nengo with a focus on speed,
+extensibility, and simplicity.
+
+\subsection{JavaNengo limitations}
+
+\begin{table}[!t]
+\processtable{Nengo tools\label{tab:nengo-vers}}
+{\begin{tabular}{p{2.6cm} p{2.7cm} l p{7cm}} \toprule
+\textbf{Name} & \textbf{Implementation language}
+  & \textbf{Code repository} & \textbf{Purpose} \\\midrule
+JavaNengo simulator & Java & \texttt{nengo\_java}
+  & Simulates many types of spiking neuron models.
+  Includes methods to construct and simulate models using the NEF. \\
+JavaNengo GUI & Java & \texttt{nengo\_java\_gui}
+  & Graphical interface to JavaNengo simulator.
+  Allows modellers to drag-and-drop components built with the NEF. \\
+JavaNengo scripting interface & Jython & \texttt{nengo\_jython}
+  & Scripting interface to JavaNengo simulator.
+  Provides shortcuts to classes and functions in the simulator.
+  Also accessible through and interacts with the JavaNengo GUI. \\
+PyNengo & Python & \texttt{nengo}
+  & Constructs and simulates models using the NEF. \\
+PyNengo OpenCL simulator & Python+PyOpenCL
+  & \texttt{nengo\_ocl}
+  & Simulates models constructed with PyNengo. \\\botrule
+\end{tabular}}{Code repositories avilable at \url{http://github.com/ctn-waterloo/}}
+\end{table}
+
+JavaNengo is a suite of tools providing
+a neural simulator and graphical user interface (GUI)
+implemented in Java,
+and a scripting interface implemented in Jython
+(see Table~\ref{tab:nengo-vers} for more detail).
+JavaNengo's simulator was created
 with the intention of being a general
 neural simulator that included
-the methods of the Neural Engineering Framework
-(NEF; \cite{TODO}) as one of many different
-methods of creating neural models.
-However, since creating the simulator in 2007,
-it has only been extended minimally,
-but has been used extensively for its ability
-to create NEF-based models.
-
-The Jython scripting interface,
+the methods of the NEF
+as one of several
+methods for creating neural models.
+As a result, its implementation
+of important NEF components
+are nested in deep class hierarchies,
+adding unnecessary layers of complexity
+for developers wishing
+to extend those NEF components.
+As a representative example,
+the most often used class for making
+NEF models, \texttt{NEFEnsemble},
+extends from five ancestor classes
+and implements seven interfaces.
+Adding support for synaptic plasticity
+in \texttt{NEFEnsemble}
+involved adding or significantly modifying
+25 Java classes \cite{TODO}.
+
+The JavaNengo GUI provided
+a drag-and-drop interface in which to create NEF models.
+However, as it was developed
+before the scripting interface,
+the GUI manipulates simulator objects directly.
+Typically, modellers start by following
+tutorials that use the GUI interface,
+and transition to the scripting interface
+in order to implement more complicated models.
+Since the GUI modifies simulator objects directly,
+modelling concepts learned in the GUI
+have to be relearned in the scripting interface.
+
+The JavaNengo scripting interface,
 described in \cite{TODO},
@@
 to the NEF-specific portions of the simulator.
-However, the Java GUI was developed at the same time
-as the Jython scripting interface,
-and for that reason primarily
-interacts with the simulator directly.
-Since large-scale models often require
-a scripting language for looping
-and parameterization,
-this has resulted in a difficult transition
-for modellers moving from the
-drag-and-drop graphical interface
-to the scripting interface.
-Additionally, advanced scripts often
-contain Jython code that either defines
-or manipulates Java objects directly,
-meaning that modellers have to know
-a certain amount of Python and Java to be productive.
-
-Since Nengo makes it possible to define
-very large models,
-performance is an important issue.
+This greatly enhanced productivity,
+and made it possible
+to create many models with very little code.
+Modellers quickly started using
+the scripting interface
+as their primary tool,
+but still needed to access the GUI
+for interactive model inspection,
+and still needed to modify the simulator
+to add new functionality
+(e.g., to add new neuron models).
+This is because Jython makes it easy
+to inspect and create Java objects,
+but does not make it easy to
+write Java code to
+inspect and create Jython objects.
+Modellers, therefore, had to be
+proficient in Java whether
+they used the scripting interface or not.
+As there are few neural modellers
+that are proficient with both Java and Python syntax,
+the scripting interface still lacks
+convenient ways to record the results of simulations,
+and to define complex experimental setups.
+
+Finally, perhaps the most significant limitation
+of the JavaNengo tools is their performance.
 While Java has good support for threads,
-it is difficult to interface with
+it does not have well-supported libraries
+for doing mathematical operations on large matrices quickly.
+Additionally, it is difficult to interface with
 platform-specific tools such as
 graphical processing units (GPUs),
-which can quickly perform the types of parallel
-computations that these neural models require.
-This is due to the
-architecture of the Java simulator
-and the Java Native Interface,
-which results in code that
-is difficult to maintain and debug.
-
-The new implementation of Nengo described in this paper
-was undertaken with three goals in mind:
+which can perform the types of computations
+that neural models require quickly and in parallel.
+
+Addressing all of these issues
+in the Java codebase would involve
+significant enough changes
+that we started from scratch instead.
+Since the scripting interface
+already used Python syntax,
+and since Python is widely used
+for neuroscientific data analysis,
+we targeted the CPython interpreter
+as the platform for the next generation of Nengo.
+
+\subsection{PyNengo}
+
+PyNengo,
+the subject of the remainder of this paper,
+is a Python package for
+defining and simulating
+neural models using the NEF.
+It targets the CPython interpreter,
+meaning that it integrates seamlessly
+with other scientific Python tools.
+Arbitrary CPython programs
+can use Nengo models,
+opening up possibilities
+for using neurally implemented algorithms
+in applications and games.
+
+PyNengo's object model is simple and minimal,
+making it easy to
+document, test, and modify.
+It also includes a scripting interface to those objects
+based on JavaNengo's scripting interface \cite{TODO},
+but with several modifications
+to allow for easier incorporation
+of user-defined networks,
+new neuron models, and learning rules.
+Additionally, PyNengo's scripting interface
+contains methods to record simulated data
+and access that data after simulation.
+
+Model creation and simulation are decoupled,
+allowing for models to be exported
+to other simulators,
+such as those accessible with PyNN.
+Simulators can also be written
+to leverage specific platforms.
+Currently, we have implemented
+a platform-independent simulator
+that uses NumPy for vectorized computations,
+and a PyOpenCL-based
+simulator that takes full advantage of GPUs
+and multicore CPUs.
+
+PyNengo does not currently
+replace all of JavaNengo's functionality.
+While it can simulate the vast majority
+of models that JavaNengo can simulate,
+some functionality is still in development.
+Most notably, there is no graphical interface
+to PyNengo, and therefore simulations
+cannot be visualized while they are running.
+However, we are confident that these tools
+can be created, and will be much easier
+to implement compared to JavaNengo's GUI.
+
+PyNengo overcomes all of JavaNengo's limitations.
+It can simulate most models
+currently simulated by JavaNengo
+using 11\% as many lines code,
+and depending on
+only three external libraries instead of ten.
+The OpenCL simulator can simulate
+a large model TODO times
+faster than JavaNengo's simulator.
+This, along with a codebase
+that is well tested and
+easy to understand and extend,
+provides a platform
+that can simulate larger and more complex
+models than Spaun,
+and can therefore further
+test the NEF as a theory of the brain.
+
+\section{Neural Engineering Framework (NEF)}
+
+The Neural Engineering Framework (NEF; \cite{TODO})
+proposes three principles
+that enable the construction
+of large-scale neural models.
+
 \begin{enumerate}
-  \item Provide a scripting interface
-    similar to the existing Jython interface.
-  \item Leverage existing scientific computing tools
-    rather than write our own.
-  \item Provide the ability to run models with
-    different software and hardware simulators.
+  \item \textbf{Representation:} A population of neurons
+    collectively represents information in the form of
+    time-varying vectors of real numbers.
+  \item \textbf{Transformation:} Functions on those vectors
+    are computed by the connections between populations.
+  \item \textbf{Dynamics:} The vectors represented
+    by neural populations can be considered state variables
+    in a dynamical system, and can be analyzed using control theory.
 \end{enumerate}
-
-In order to achieve these goals,
-this new version of Nengo uses the Python language
-and targets the CPython interpreter,
-allowing Nengo to use existing scientific Python
-tools for vectorized computations, plotting, and graph manipulation.
-As evidence that we have met these goals,
-we have implemented two different software simulators,
-and have done so with an order of magnitude
-less lines of code than the Java simulator
-and Jython scripting interface combined.
-For the remainder of the paper,
-we will use ``Nengo'' to refer to
-this new Python rewrite.
-
-\subsection{Neural Engineering Framework (NEF)}
-
-TODO; make sure it meshes with 3.1's operators
-
-\begin{equation} \label{TODO}
-  TODO
-\end{equation}
-
-%% This should be a paper of its own.
-
-%% \subsection{Semantic Pointer Architecture (SPA)}
-
-%% TODO; should we even include this section?
-%% It could probably be a separate paper on its own.
-
-\section{Model definition}
-
-One of the strengths of the NEF
-is that it allows neural modellers to think
-at a higher level
-than is normally associated with neural models.
-With the NEF, a modeller decides
-what information is being represented
-by populations of neurons and how that information
-is being transformed through the connections
-between populations;
-this contrasts with modelling decisions
-that are typically thought to be important,
-such as the morphology of each neuron,
-and the distribution of connection weights
-between populations.
-Instead, these decisions are made---often
-automatically---as a direct result
-of the information
-being represented and the transformations
-being performed on that information.
-This situates Nengo as a
-``neural compiler'' that can
-take a functional model description
-and produce a low-level neural model
-that can be simulated
-with any neural simulation software.
-
-To emphasize Nengo's role as a neural compiler,
-Nengo allows models to be defined
-in a simulator-agnostic manner,
-allowing for functionally identical models to be run
-on different software and hardware platforms.
-
-\subsection{Nengo objects} \label{sec:objects}
 
@@
 \begin{center}
-  Nengo high-level object hierarchy
+  \includegraphics[width=0.19\textwidth]{nef_summary_enc}
+  \includegraphics[width=0.19\textwidth]{nef_summary_dec}
+  \includegraphics[width=0.399\textwidth]{nef_summary_trans}
+  \includegraphics[width=0.199\textwidth]{nef_summary_dyn}
 \end{center}
- \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{
-   Nengo high-level object hierarchy}
+ \textbf{\refstepcounter{figure}\label{fig:nef} Figure \arabic{figure}.}{
+   Summary of the Neural Engineering Framework.}
 \end{figure}
 
+\paragraph{Representation}
+Information is encoded by populations of neurons.
+We consider that information
+to be time-varying vectors of real numbers,
+in order to reason with that information
+using conventional mathematics.
+By doing this, we can \textit{encode}
+those vectors by injecting
+specific amounts of current into
+single neuron models based on
+the vector being encoded.
+This drives the neuron,
+causing it to spike.
+We can then estimate
+the originally encoded vector
+through a \textit{decoding} process.
+This idea is similar to population coding
+\cite{TODO}, but generalized
+to vectors of arbitrary length.
+
+Figure~\ref{fig:nef} depicts
+the encoding (A) and decoding (B) process.
+In the encoding process, the input signal drives
+each neuron based on its \textit{tuning curve},
+which describes how much
+that neuron fires for a given input signal.
+The tuning curve summarizes the gain
+of a neuron (how quickly the activity rises),
+the bias (the activity of a neuron given no signal),
+and the encoding weight
+(the part of the vector space
+that causes the neuron to fire most strongly).
+This results in a predictable pattern of firing
+given the input signal.
+
+In the decoding process,
+the spike trains are first filtered
+with an exponentially decaying filter.
+Those filtered spike trains are summed together
+with weights that are determined
+by solving a least-squares minimization problem.
+In the bottom panel of Figure~\ref{fig:nef}B,
+the red line is the filtered spiking activity
+weighted by decoding weights
+that minimize the difference between
+the actual encoded vector
+and the decoded estimate.
+The purple line is the filtered spiking activity
+weighted by decoding weights
+that minimize the different between
+the cosine of the actual encoded vector
+and the decoded estimate.
+Note that these decoding weights are not determined
+based on the input signal;
+instead, the minimization occurs
+on points that are randomly sampled
+in the vector space
+that the population represents.
+
+\paragraph{Transformation}
+Neurons communicate through
+unidirectional connections called synapses.
+When a neuron spikes,
+it releases neurotransmitter across the synapse,
+which causes some amount of current
+to be imparted in the downstream neuron.
+Many factors affect the
+amplitude of the current imparted;
+we summarize those factors
+in a scalar connection weight
+representing the strength
+of the connection between two neurons.
+In order to compute any function,
+we can set the connection weights between
+two populations to be the product of
+the decoding weights for that function
+for neurons in the first population
+and the encoding weights
+for the downstream neuron.
+
+In Figure~\ref{fig:nef}C,
+a sine wave is transformed twice.
+Between populations A and B,
+the signal is negated.
+Between populations B and C,
+the signal is squared.
+
+Note that, in practice, we rarely use
+the full connection weight matrix,
+and instead store
+the encoding and decoding weights
+(i.e., the factors of the connection weight matrix).
+This provides significant
+space and time savings.
+
+\paragraph{Dynamics}
+While doing feedforward transformations
+on vectors is sufficient to model
+many neural systems,
+many systems, notably memory systems,
+require persistent activity through recurrent connections.
+When recurrent connection are introduced,
+the vectors that populations represent
+can be thought of as state variables
+in a dynamical system.
+Neural systems can therefore
+be built and analyzed using
+the methods of control theory.
+
+In Figure~\ref{fig:nef}D,
+a population represents a two-dimensional vector,
+and is recurrently connected
+with negative feedback between dimensions,
+resulting in a harmonic oscillator.
+
+\paragraph{Nengo}
+Large models can be built
+by thinking of the principles of the NEF
+as building blocks that can be put together
+to describe neural systems.
+The goal of Nengo is to allow
+modellers to describe neural systems
+in terms of what information is being represented,
+how that information is transformed,
+and what dynamics are required
+to perform cognitively relevant
+functions with biologically constrained
+neural components.
+Those descriptions are then
+translated to a network
+of interconnected neurons.
+This situates Nengo
+as a ``neural compiler''
+that translates
+a high-level functional model
+to a low-level neural model
+that implements the high-level function.
+
+\section{PyNengo object model}
+
+\subsection{Model}
+
 A Nengo model is a collection
-of Nengo objects connected together,
-and probed in order to collect data
-during simulation.
+of Nengo objects that represent information
+and connections that transform information.
+Objects and connections can be probed
+in order to collect data during simulation.
+
 The \texttt{Model} object is primarily a container
 for Nengo objects,
-but also provides methods
-that simplify creating, connecting,
-and probing these objects.
-Basic Nengo models
+but also doubles as
+a scripting interface
+that simplifies creating, connecting,
+and probing objects.
+Basic models
 can be created solely through interacting
-with the \texttt{Model} object.
-For more advanced models,
-these objects can be instantiated
-directly and then added to the model.
-
-There are two types of objects
-defined in the main scripting interface.
-
-\paragraph{Ensembles}
-The most common is the \texttt{Ensemble},
-which is a population of neurons
+with the \texttt{Model} object
+(see sections \ref{sec:comm-channel}--\ref{sec:lorenz}
+for examples).
+In advanced models,
+objects are instantiated
+and then added to the model
+(see section \ref{sec:cconv} for an example).
+
+\subsection{Ensemble}
+
+An \texttt{Ensemble} is
+a population of neurons
 that represents information
 in the form of a real-valued vector.
-As described in \eqref{TODO},
-this can be done
-with any neuron model, $G$.
 An ensemble must be given as arguments
@@
 the population of neurons.
-For example, \texttt{LIF(50, tau\_rc=0.02)}
+For example, \texttt{nengo.LIF(50, tau\_ref=0.002)}
 describes a population
 of 50 leaky integrate-and-fire neurons
-with an RC time constant of 0.02 seconds.
-Neurons are defined in this symbolic manner
+with a refractory time of 0.002 seconds.
+Neurons are defined symbolically
 so that each simulator can compute
@@
 however it sees fit.
+The neuron model used by the population
+is easily changed, as long as the simulat
 
@@
 
-%% However, not all simulators can support all neuron types,
-%% and all of the attributes that can be
-%% set on an ensemble.
-%% This is particularly the case
-%% when interacting with special-purpose hardware.
-%% In these cases, simulators should
-%% raise appropriate exceptions
-%% when unsupported ensembles are encountered.
-
-\paragraph{Nodes}
+\subsection{Node}
 A \texttt{Node} tracks any information
@@
 
-\paragraph{Connections}
+\subsection{Connections}
 
@@
 together with connection weights.
-%TODO: reword? is it important that it's a factored weight matrix
-
-\paragraph{Probes}
+
+\subsection{Probes}
 
@@
 
-\section{Model simulation} \label{sec:simulators}
+\subsection{Networks}
+A network is a collection of ensembles and nodes
+connected together in a particular way.
+It is primarily a way of grouping together
+a common set of objects connected in a particular way,
+and provides some implicit expressiveness
+in the name of the network.
+
+Networks can vary dramatically in size,
+and can be composed of other networks.
+The \texttt{Integrator} network, for example,
+is composed of only one recurrently connected ensemble.
+By encapsulating that logic in a network,
+the purpose of that ensemble is made explicit.
+On the other end of the size spectrum,
+the \texttt{BasalGanglia} network,
+is composed of five groups of ensembles
+connected with several specific functions
+that together implement a ``winner take all'' circuit.
+Encapsulating that logic
+makes this complicated section of code
+easy to include in many different models
+(see Figure~\ref{TODO}).
+
+\section{PyNengo simulators} \label{sec:simulators}
 
@@
 \begin{center}
-  Reference simulator build process
+  \includegraphics[width=0.6\textwidth]{sim}
 \end{center}
- \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{
-   Reference simulator build process}
+ \textbf{\refstepcounter{figure}\label{fig:sim} Figure \arabic{figure}.}{
+   Sim}
 \end{figure}
@@
 
-\begin{figure}
-\begin{center}
-  Reference simulator simulation (include dependency graph)
-\end{center}
- \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{
-   Reference simulator simulation}
-\end{figure}
-
 The reference simulator starts by
@@
 
-\section{Extensibility}
+\subsection{Extensibility}
 
@@
 
-\subsection{Models}
-
-At model creation time,
-everything must be done
-in a simulator agnostic manner.
-Since ensembles and nodes
-are the only objects that exist,
-models can become long scripts
-that contain many
-similar function calls that
-create and connect ensembles and nodes.
-In order to make modelling scripts
-shorter and more expressive,
-we have created the \texttt{Network} abstraction.
-
-\paragraph{Networks}
-A network is a collection of ensembles and nodes
-connected together in a particular way.
-It is primarily a way of grouping together
-a common set of objects connected in a particular way,
-and provides some implicit expressiveness
-in the name of the network.
-
-Networks can vary dramatically in size,
-and can be composed of other networks.
-The \texttt{Integrator} network, for example,
-is composed of only one recurrently connected ensemble.
-By encapsulating that logic in a network,
-the purpose of that ensemble is made explicit.
-On the other end of the size spectrum,
-the \texttt{BasalGanglia} network,
-is composed of five groups of ensembles
-connected with several specific functions
-that together implement a ``winner take all'' circuit.
-Encapsulating that logic
-makes this complicated section of code
-easy to include in many different models
-(see Figure~\ref{TODO}).
-
-\begin{figure}
-\begin{center}
-  Network code listing
-\end{center}
- \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{
-   Network code listing}
-\end{figure}
-
-\subsection{Nonlinearities}
+\subsubsection{Nonlinearities}
 
@@
 
-\subsection{Model simulation}
+\subsubsection{Model simulation}
 
@@
 
-\section{Examples and benchmarks}
-
-\subsection{Lorenz attractor network}
+\section{Examples} \label{sec:examples}
+
+\subsection{Communication channel} \label{sec:comm-channel}
+
+Faithfully transmit a signal.
+
+\begin{figure}
+\begin{center}
+  \includegraphics[width=.9\textwidth]{comm_channel}
+  \begin{minipage}{0.58\textwidth}
+    \begin{lstlisting}
+      import nengo
+      from nengo.helpers import white_noise
+      model = nengo.Model("Communication Channel")
+      model.make_node("Input", white_noise(1, 5))
+      model.make_ensemble("A", nengo.LIF(30), 1)
+      model.make_ensemble("B", nengo.LIF(30), 1)
+      model.connect("Input", "A")
+      model.connect("A", "B")
+      model.probe("Input")
+      model.probe("A", filter=0.01)
+      model.probe("A.spikes")
+      model.probe("B", filter=0.01)
+      model.probe("B.spikes")
+      sim = model.simulator()
+      sim.run(1)
+      A_data = sim.data("A")
+      A_spikes = sim.data("A.spikes")
+      ...
+    \end{lstlisting}
+  \end{minipage}
+  \begin{minipage}{0.37\textwidth}
+    \includegraphics[width=\textwidth]{comm_channel_res}
+  \end{minipage}
+\end{center}
+ \textbf{\refstepcounter{figure}\label{fig:cchannel} Figure \arabic{figure}.}{
+   Communcation channel model. (Top) Diagram of the model.
+   (Bottom left) Python code to create and run the model.
+   (Bottom right) Results of the model simulation. Lines depict
+   the node's output (Input panel) and the decoded value being represented
+   by ensembles (A and B panels). Light gray lines in A and B panels denote
+   spikes emitted by neurons in the ensembles. Each row depicts spikes
+   for one neuron.}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+  \begin{minipage}{.495\textwidth}
+    \begin{lstlisting}[basicstyle={\footnotesize\ttfamily}]
+      # We use Nengo to generate
+      #     bias: Bias currents
+      #     decoders: Decoding weights
+      #     encoders: Encoding weights
+      # for populations A and B.
+      import pynn.nest as pyNN
+
+      lif_params = {'tau_refrac': 2.0, 'tau_syn_E':100,
+                     'tau_syn_I':100}
+      dt = 0.5
+      pyNN.setup(timestep=dt, min_delay=dt)
+      A = pyNN.Population(30, pyNN.IF_cond_exp, lif_params)
+      B = pyNN.Population(30, pyNN.IF_cond_exp, lif_params)
+      input_signal = 0.5
+      inputA = [pyNN.DCSource(amplitude=val) for val
+                 in A_bias + input_signal * A_encoders]
+      for i, pulse in enumerate(inputA):
+          pulse.inject_into(A[i:i+1])
+      inputB = [pyNN.DCSource(amplitude=val) for val
+                 in B_bias]
+      for i, pulse in enumerate(inputB):
+          pulse.inject_into(B[i:i+1])
+    \end{lstlisting}
+  \end{minipage}
+  \begin{minipage}{.495\textwidth}
+    \begin{lstlisting}[basicstyle={\footnotesize\ttfamily}]
+      weights = []
+      for i in xrange(30):
+          for j in xrange(30):
+              weights.append((i, j,
+                  np.dot(A_decoder[i], B_encoder[j]), 1.0))
+      connection = pyNN.Projection(
+          A, B, pyNN.FromListConnector(weights))
+      A.record('spikes')
+      B.record('spikes')
+      pyNN.run(1000)
+
+      # Decoded value of B
+      B_output = numpy.zeros(1000)
+      for i in xrange(1000):
+          spikes = B[i:i+1].getSpikes()[:,1]
+          spikes = (spikes / dt).astype('int')
+          B_output[spikes] += B_decoder[i]
+      # Filter with tau = 200 ms
+      decay = np.exp(-dt / 200)
+      B_output[0, :] *= (1 - decay)
+      for i in xrange(1, 1000):
+          B_output[i,:] = decay * B_output[i-1,:]
+                           + (1-decay) * B_output[i,:]
+    \end{lstlisting}
+  \end{minipage}
+\end{center}
+ \textbf{\refstepcounter{figure}\label{fig:pynn} Figure \arabic{figure}.}{
+   An implementation of the communication channel in PyNN.}
+\end{figure}
+
+\subsection{Lorenz attractor network} \label{sec:lorenz}
 
@@
 \begin{center}
-  Nengo Lorenz code listing
+  \begin{minipage}{0.43\textwidth}
+    \begin{lstlisting}[basicstyle={\footnotesize\ttfamily}]
+      import nengo
+      tau = 0.1; sigma = 10; beta = 8.0/3; rho = 28
+      model = nengo.Model('Lorenz attractor')
+      model.make_ensemble('State', nengo.LIF(2000),
+                            dimensions=3, radius=60)
+
+      def feedback(x):
+          dx0 = -sigma * x[0] + sigma * x[1]
+          dx1 = -x[0] * x[2] - x[1]
+          dx2 = x[0] * x[1] - beta * (x[2] + rho) - rho
+
+          return [dx0 * tau + x[0],
+                  dx1 * tau + x[1],
+                  dx2 * tau + x[2]]
+
+      model.connect('State', 'State',
+                      function=feedback, filter=tau)
+      model.probe('State', filter=tau)
+      model.probe('State.spikes')
+      sim = model.simulator()
+      sim.run(6)
+    \end{lstlisting}
+  \end{minipage}
+  \begin{minipage}{0.55\textwidth}
+    \includegraphics[width=0.95\textwidth]{lorenz}
+    \includegraphics[width=\textwidth]{lorenz_res}
+  \end{minipage}
 \end{center}
- \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{
-   Nengo Lorenz code listing}
+ \textbf{\refstepcounter{figure}\label{fig:lorenz} Figure \arabic{figure}.}{
+   Lorenz attractor model. (Left) Python code to create and run the model.
+   (Top right) Diagram of the model.
+   (Bottom right) Results of the model simulation. Lines depict
+   the node's output (Input panel) and the decoded value being represented
+   by ensembles (A and B panels). Light gray lines in A and B panels denote
+   spikes emitted by neurons in the ensembles. Each row depicts spikes
+   for one neuron.}
 \end{figure}
@@
 
-\begin{figure}
-\begin{center}
-  PyNN Lorenz code listing
-\end{center}
- \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{
-   PyNN Lorenz code listing}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-  Lorenz attractor benchmarks
-\end{center}
- \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{
-   Lorenz benchmarks}
-\end{figure}
-
-\subsection{Circular convolution}
+\subsection{Circular convolution} \label{sec:cconv}
 
@@
 \begin{center}
-  Nengo CConv code listing
+  \includegraphics[width=0.8\textwidth]{cconv}
+  \begin{minipage}{0.53\textwidth}
+    \begin{lstlisting}[basicstyle={\footnotesize\ttfamily}]
+      import nengo
+      from nengo.networks import CircularConvolution
+      model = nengo.Model("Circular convolution")
+      model.make_node("Input A", [-0.21, 0.5, 0.12, 0.06])
+      model.make_ensemble("A", nengo.LIF(512), 4)
+      model.connect("Input A", "A")
+      model.make_node("Input B", [-0.18, 0.28, 0.18, -0.52])
+      model.make_ensemble("B", nengo.LIF(512), 4)
+      model.connect("Input B", "B")
+      cconv = model.add(CircularConvolution("Circular Convolution",
+          neurons=nengo.LIF(1032), dimensions=4))
+      model.connect("A", cconv.A)
+      model.connect("B", cconv.B)
+      model.make_ensemble("Result", nengo.LIF(512), 4)
+      model.connect(cconv, "Result")
+      model.probe("A", filter=0.02)
+      model.probe("B", filter=0.02)
+      model.probe("Result", filter=0.02)
+      sim = model.simulator()
+      sim.run(0.5)
+    \end{lstlisting}
+  \end{minipage}
+  \begin{minipage}{0.43\textwidth}
+    \includegraphics[width=\textwidth]{cconv_res}
+  \end{minipage}
 \end{center}
- \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{
-   Nengo CConv code listing}
+ \textbf{\refstepcounter{figure}\label{fig:lorenz} Figure \arabic{figure}.}{
+   Circular convolution model. (Top) Diagram of the model.
+   Note that much of the complexity is contained within the
+   circular convolution network; the transform matrices shown
+   are automatically generated by the network depending on
+   input dimensionality.
+   (Bottom left) Python code to create and run the model.
+
+   (Bottom right) Results of the model simulation. Lines depict
+   the node's output (Input panel) and the decoded value being represented
+   by ensembles (A and B panels). Light gray lines in A and B panels denote
+   spikes emitted by neurons in the ensembles. Each row depicts spikes
+   for one neuron.}
 \end{figure}
@@
 TODO: can we do circconv in PyNN?
+
+\section{Benchmarks}
 
@@
 \begin{center}
-  PyNN CConv code listing
+  \includegraphics[width=0.5\textwidth]{ocl_vs_ref}
 \end{center}
- \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{
-   PyNN CConv code listing}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-  CConv benchmarks
-\end{center}
- \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{
-   CConv benchmarks}
+ \textbf{
+     \refstepcounter{figure}\label{fig:ocl_vs_ref}
+     Figure \arabic{figure}.}{
+     Five-hundred-dimensional circular convolution networks were some of the most
+     computationally demanding elements of the Spaun cognitive architecture.
+     Simulations of circular convolution networks of various sizes illustrate the speed of our re-engineered simulators.
+     A straightforward reference implementation in Python is sufficient for small networks,
+     while for larger ones the new OpenCL-powered simulator runs up to 200$\times$ faster,
+     In particular the Radeon 7970 GPU can simulate 500-dimensional convolution networks faster than real time (0.95 seconds for a 1-second simulation).
+     Also, although both are multithreaded, our new implementation delivers better CPU performance than the previous JavaNengo implementation.
+     }
 \end{figure}

### 2013-09-24 nengo.tex
Started implementing Aaron's suggestions

@@
 
+However, the ability to connect neurons together
+with synaptic connection weights is still possible
+with the \texttt{NonlinearityConnection}.
+Both of these connection types
+can be temporally filtered,
+and the weights involved in the connection
+(decoding weights for \texttt{DecodedConnection},
+synaptic connection weights for \texttt{NonlinearityConnection})
+can be modified over time with learning rules.
+
 \subsection{Probes}
+
+A probe is used to denotes a signal that
+whose state will be recorded
+throughout the simulation.
+Nengo objects and connections
+each have several signals that can be probed.
+For an ensemble's decoded value,
+or the underlying neurons' membrane voltages
+can be probed, among others.
 
@@
 and stored with no overhead to the simulator.
-In practice, recording data
+However, in practice, recording data
 is a costly part of most simulations,
@@
 a core component of a Nengo model.
-
-In general, Nengo objects and connections
-each have several signals that can be probed.
-For example, an ensemble's decoded value,
-or the underlying neurons' membrane voltages
-can be probed, among other signals.
-The underlying logic for the probe
-leverages connection logic,
-making it easy to probe any part of an object.
 
@@
 as reference, the model goes through a
-two-stage build process in order to
+two-stage process in order to
 simulate models as fast as possible.

### 2013-09-24 nengo.tex
Added table with different tools

--- 2013-09-23 nengo.tex
@@
 \subsection{Why reimplement Nengo?}
+
+\begin{table}[!t]
+\processtable{Nengo tools\label{tab:nengo-vers}}
+{\begin{tabular}{p{2.6cm} p{2.7cm} l p{7cm}} \toprule
+\textbf{Name} & \textbf{Implementation language}
+  & \textbf{Code repository} & \textbf{Purpose} \\\midrule
+JavaNengo simulator & Java & \texttt{nengo\_java}
+  & Simulates many types of spiking neuron models.
+  Includes methods to construct and simulate models using the NEF. \\
+JavaNengo GUI & Java & \texttt{nengo\_java\_gui}
+  & Graphical interface to JavaNengo simulator.
+  Allows modellers to drag-and-drop components built with the NEF. \\
+JavaNengo scripting interface & Jython & \texttt{nengo\_jython}
+  & Scripting interface to JavaNengo simulator.
+  Provides shortcuts to classes and functions in the simulator.
+  Also accessible through and interacts with the JavaNengo GUI. \\
+PyNengo & Python & \texttt{nengo}
+  & Constructs and simulates models using the NEF. \\
+PyNengo OpenCL simulator & Python+PyOpenCL
+  & \texttt{nengo\_ocl}
+  & Simulates models constructed with PyNengo. \\\botrule
+\end{tabular}}{Footnote?}
+\end{table}
 

### 2013-09-25 nengo.tex
New introduction based on feedback

@@
 
-Nengo is a set of software tools built around
-a theoretical framework for building neural models
-proposed by Eliasmith \& Anderson \cite{TODO},
-and more recently, an architecture
-for performing high-level cognitive tasks
-in that framework proposed by Eliasmith \cite{TODO}.
-Nengo consists of a Java-based simulator,
-a Jython-based scripting interface
-for common use cases of the simulator,
-and a Java-based GUI that
-uses both the underlying simulator
-and the scripting interface.
-This set of tools has been used
-to create models of visual attention \cite{TODO},
-inductive reasoning \cite{TODO},
-and reinforcement learning \cite{TODO},
-as well as Spaun, a cognitive agent able
-to do these and several other tasks
-(\cite{TODO}).
-Nengo has been used by dozens of researchers
-in labs at the University of Waterloo,
-Johns Hopkins University, Yale, Stanford,
-and others (TODO: confirm).
-
-Here, we describe a new tool
-for creating and simulating NEF-based models
-implemented in Python,
-which effectively merges two existing tools
-into one unified package
-that can be extended in
-several ways that previous Nengo tools could not.
-
-\subsection{Why reimplement Nengo?}
+It is often said that neuroscience
+is data-rich but theory-poor.
+Although computational neuroscience
+has grown significantly,
+this phrase continues to be repeated
+due to a lack of satisfying
+theories of the brain.
+Much work has been put into simulators
+that attempt to recreate neural
+data in precise detail,
+but behavior has not emerged
+from data-driven large scale models.
+Similarly, theoretical frameworks
+that involve a generic algorithm
+replicated many times
+(e.g., the memory-prediction framework \cite{TODO}
+and the pattern recognition theory of mind \cite{TODO})
+have not been shown to solve problems
+outside of those that
+the generic algorithm alone can solve.
+
+Nengo is a neural simulator
+based on a theoretical framework proposed
+by Eliasmith \& Anderson \citeyearpar{TODO}
+called the Neural Engineering Framework
+(NEF).
+Nengo was recently used to implement Spaun,
+currently the world's largest functional brain model
+\cite{TODO}.
+Spaun is a network of 2.5 million spiking neurons
+that can perform eight high-level cognitive tasks
+including list memory and inductive reasoning.
+It can perform any of these eight tasks
+at any time by being presented
+the appropriate series of images
+representing the task to be performed
+and the details of that task
+(e.g., $A3[1234]$ instructs Spaun
+to remember the list $1234$).
+It then provides its answer by
+sending motor commands to a simulated arm model,
+tracing out digits
+(e.g., $1234$ if asked to recall the remembered list).
+
+While the tasks Spaun performs are diverse,
+all tasks use a common set of
+functional modules that we map
+onto the brain areas hypothesized
+to perform those functions.
+As a result, Spaun is able to match
+experimental data across the spectrum
+of neuroscientific data,
+from the single cell level
+to the behavioral level,
+despite experimental data
+being a very small part of the model description.
+Similarly, while this is made possible
+by a theoretical framework,
+the NEF enables Spaun to employ
+many different algorithms with spiking neurons,
+rather than defining a single algorithm
+that is replicated many times.
+
+In implementing Spaun,
+we have provided strong support
+that the NEF is beginning to fill the theory void in neuroscience.
+However, we have also pushed Nengo
+to the limits of its ability to construct and simulate
+large-scale neural models.
+In order to continue to evaluate the NEF
+as a theory of the brain
+by building larger and more complicated models,
+we have begun the next generation
+of Nengo tools with a focus on simplicity,
+extensibility, and speed.
+
+\subsection{JavaNengo limitations}
 
@@
   & Simulates models constructed with PyNengo. \\\botrule
-\end{tabular}}{Footnote?}
+\end{tabular}}{}
 \end{table}
 
-The original Java simulator was created
+JavaNengo is a suite of tools providing
+a neural simulator and graphical user interface (GUI)
+implemented in Java,
+and a scripting interface implemented in Jython
+(see Table~\ref{tab:nengo-vers} for more detail).
+JavaNengo's simulator was created
 with the intention of being a general
 neural simulator that included
-the methods of the Neural Engineering Framework
-(NEF; \cite{TODO}) as one of many different
+the methods of the NEF
+as one of many different
 methods of creating neural models.
-However, since creating the simulator in 2007,
-it has only been extended minimally,
-but has been used extensively for its ability
-to create NEF-based models.
-
-The Jython scripting interface,
+As a result, its implementation
+of important NEF components
+are nested in deep class hierarchies,
+adding unnecessary layers of complexity
+for developers wishing
+to extend those NEF components.
+The addition of synaptic plasticity rules,
+for example, involved adding or significantly modifying
+25 Java classes \cite{TODO}.\footnote{Changelog at
+  \url{https://github.com/ctn-waterloo/nengo_1.4/commit/93eee51147545f7ce30ae9919809e3072dc15728}.}
+
+The JavaNengo GUI provided
+a drag-and-drop interface in which to create NEF models.
+However, as it was developed
+before the scripting interface,
+the GUI manipulates simulator objects directly.
+Typically, modellers start by following
+tutorials that use the GUI interface,
+and transition to the scripting interface
+in order to implement more complicated models.
+Since the GUI modifies simulator objects directly,
+modelling concepts learned in the GUI
+have to be relearned in the scripting interface.
+
+The JavaNengo scripting interface,
 described in \cite{TODO},
@@
 to the NEF-specific portions of the simulator.
-However, the Java GUI was developed at the same time
-as the Jython scripting interface,
-and for that reason primarily
-interacts with the simulator directly.
-Since large-scale models often require
-a scripting language for looping
-and parameterization,
-this has resulted in a difficult transition
-for modellers moving from the
-drag-and-drop graphical interface
-to the scripting interface.
-Additionally, advanced scripts often
-contain Jython code that either defines
-or manipulates Java objects directly,
-meaning that modellers have to know
-a certain amount of Python and Java to be productive.
-
-Since Nengo makes it possible to define
-very large models,
-performance is an important issue.
+This greatly enhanced productivity,
+and made it possible
+to create many models with very little code.
+Modellers quickly started using
+the scripting interface
+as their primary tool,
+but still needed to access the GUI
+for interactive model inspection,
+and still needed to modify the simulator
+to add new functionality
+(e.g., new neuron models).
+This is because Jython makes it easy
+to inspect and create Java objects,
+but does not make it easy to
+write Java code to
+inspect and create Jython objects.
+Modellers, therefore, had to be
+proficient in Java whether
+they used the scripting interface or not.
+As the set of neural modellers
+that are proficient with both Java and Python
+syntax is small,
+the scripting interface still lacks
+convenient ways to record the results of simulations,
+and to define complex experimental setups.
+
+Finally, perhaps the most significant limitation
+of the JavaNengo tools is its performance.
 While Java has good support for threads,
-it is difficult to interface with
+it does not have well-supported libraries
+for doing mathematical operations on large matrices quickly.
+Additionally, it is difficult to interface with
 platform-specific tools such as
 graphical processing units (GPUs),
-which can quickly perform the types of parallel
-computations that these neural models require.
-This is due to the
-architecture of the Java simulator
-and the Java Native Interface,
-which results in code that
-is difficult to maintain and debug.
-
-The new implementation of Nengo described in this paper
-was undertaken with three goals in mind:
-\begin{enumerate}
-  \item Provide a scripting interface
-    similar to the existing Jython interface.
-  \item Leverage existing scientific computing tools
-    rather than write our own.
-  \item Provide the ability to run models with
-    different software and hardware simulators.
-\end{enumerate}
-
-In order to achieve these goals,
-this new version of Nengo uses the Python language
-and targets the CPython interpreter,
-allowing Nengo to use existing scientific Python
-tools for vectorized computations, plotting, and graph manipulation.
-As evidence that we have met these goals,
-we have implemented two different software simulators,
-and have done so with an order of magnitude
-less lines of code than the Java simulator
-and Jython scripting interface combined.
-For the remainder of the paper,
-we will use ``Nengo'' to refer to
-this new Python rewrite.
-
-\subsection{Neural Engineering Framework (NEF)}
+which can perform the types of computations
+that these neural models require quickly and in parallel.
+
+Addressing all of these issues
+in the Java codebase would involve
+significant enough changes
+that we decided to start from scratch instead.
+Since the scripting interface
+already used Python syntax,
+and we were using Python
+and its scientific libraries
+for analyzing simulation results,
+we targeted the CPython interpreter
+as the platform for the new generation of Nengo tools.
+
+\subsection{PyNengo}
+
+PyNengo, which we describe in more detail
+in the remainder of this paper,
+is a Python package for
+defining and simulating
+neural models using the NEF.
+It targets the CPython interpreter,
+meaning that it integrates seamlessly
+with other scientific Python tools.
+Nengo models can used
+in arbitrary CPython programs,
+opening up possibilities
+for using neurally inspired algorithms
+in applications and games.
+
+PyNengo's object model has been radically simplified,
+making it easy to
+document, test, and modify those objects.
+It also includes a scripting interface to those objects
+based on JavaNengo's scripting interface \cite{TODO},
+but with several modifications
+to allow for easier incorporation
+of new neuron models and learning rules.
+Additionally, PyNengo's scripting interface
+contains methods to record simulated data
+and access that data after simulation.
+
+Model creation and simulation
+have been completely decoupled,
+allowing for models to be exported
+to other simulators,
+such as those accessible with PyNN.
+Simulators can also be written
+to leverage specific platforms.
+Currently, we have implemented
+a platform-independent simulator
+that uses NumPy for vectorized computations,
+and a PyOpenCL-based
+simulator that takes full advantage of GPUs
+and multicore CPUs.
+
+PyNengo does not currently
+replace all of the JavaNengo tools.
+Most notably, there is no graphical interface
+to PyNengo, and therefore little ability to interact
+with simulations as they are running.
+However, we are confident that these tools
+can be created, and will be much easier
+to implement compared to JavaNengo's GUI.
+
+We believe that PyNengo overcomes
+all of the significant issues
+with JavaNengo.
+It can simulate the majority of models
+currently simulated by JavaNengo,
+using 11\% as many lines code,
+and depending on
+only three external libraries instead of ten.
+The OpenCL simulator can simulate
+a large scale network TODO times
+faster than JavaNengo's simulator.
+This, along with a codebase
+that is easier to understand and extend,
+provides a platform
+that can simulate larger and more complex
+models than Spaun,
+and can therefore further
+test the NEF as a theory of the brain.
+
+\section{Neural Engineering Framework (NEF)}
 

### 2013-09-26 nengo.tex
Added figures for example networks

@@
 \def\corrEmail{tbekolay@uwaterloo.ca}
+
+%%%%% Document begins
 
@@
 to perform those functions.
-As a result, Spaun is able to match
-experimental data across the spectrum
-of neuroscientific data,
-from the single cell level
-to the behavioral level,
-despite experimental data
-being a very small part of the model description.
-Similarly, while this is made possible
-by a theoretical framework,
+Like data-driven large scale models,
+Spaun is able to match experimental data
+across the neuroscientific spectrum,
+from the single cell to the behavioral level.
+However, instead of constructing the model
+based on data and hoping behavior emerges,
+we construct the model
+based on behavior
+and see that the data emerges.
+This is made possible by a theoretical framework
+like the memory-prediction framework \cite{TODO},
+but unlike it and other frameworks,
 the NEF enables Spaun to employ
-many different algorithms with spiking neurons,
+many different algorithms
+based on the behavior being performed,
 rather than defining a single algorithm
-that is replicated many times.
+and hoping that replicating it
+many times will enable these behaviors.
 
@@
 what is known about pyramidal cells in neocortex.
-
-%% However, not all simulators can support all neuron types,
-%% and all of the attributes that can be
-%% set on an ensemble.
-%% This is particularly the case
-%% when interacting with special-purpose hardware.
-%% In these cases, simulators should
-%% raise appropriate exceptions
-%% when unsupported ensembles are encountered.
 
@@
 
-\section{Examples and benchmarks}
+\section{Examples}
+
+\subsection{Communication channel}
+
+Faithfully transmit a signal.
+
+\begin{figure}
+\begin{center}
+  \includegraphics[width=.9\textwidth]{comm_channel}
+  \begin{minipage}{0.58\textwidth}
+    \begin{lstlisting}
+      import nengo
+      from nengo.helpers import white_noise
+      model = nengo.Model("Communication Channel")
+      model.make_node("Input", white_noise(1, 5))
+      model.make_ensemble("A", nengo.LIF(30), 1)
+      model.make_ensemble("B", nengo.LIF(30), 1)
+      model.connect("Input", "A")
+      model.connect("A", "B")
+      model.probe("Input")
+      model.probe("A", filter=0.01)
+      model.probe("A.spikes")
+      model.probe("B", filter=0.01)
+      model.probe("B.spikes")
+      sim = model.simulator()
+      sim.run(1)
+      A_data = sim.data("A")
+      A_spikes = sim.data("A.spikes")
+      ...
+    \end{lstlisting}
+  \end{minipage}
+  \begin{minipage}{0.37\textwidth}
+    \includegraphics[width=\textwidth]{comm_channel_res}
+  \end{minipage}
+\end{center}
+ \textbf{\refstepcounter{figure}\label{fig:cchannel} Figure \arabic{figure}.}{
+   Communcation channel model. (Top) Cartoon depiction of the model.
+   (Bottom left) Python code to create and run the model.
+   (Bottom right) Results of the model simulation. Lines depict
+   the node's output (Input panel) and the decoded value being represented
+   by ensembles (A and B panels). Light gray lines in A and B panels denote
+   spikes emitted by neurons in the ensembles. Each row depicts spikes
+   for one neuron.}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+  \begin{minipage}{.495\textwidth}
+    \begin{lstlisting}[basicstyle={\footnotesize\ttfamily}]
+      # We use Nengo to generate
+      #     bias: Bias currents
+      #     decoders: Decoding weights
+      #     encoders: Encoding weights
+      # for populations A and B.
+      import pynn.nest as pyNN
+
+      lif_params = {'tau_refrac': 2.0, 'tau_syn_E':100,
+                     'tau_syn_I':100}
+      dt = 0.5
+      pyNN.setup(timestep=dt, min_delay=dt)
+      A = pyNN.Population(30, pyNN.IF_cond_exp, lif_params)
+      B = pyNN.Population(30, pyNN.IF_cond_exp, lif_params)
+      input_signal = 0.5
+      inputA = [pyNN.DCSource(amplitude=val) for val
+                 in A_bias + input_signal * A_encoders]
+      for i, pulse in enumerate(inputA):
+          pulse.inject_into(A[i:i+1])
+      inputB = [pyNN.DCSource(amplitude=val) for val
+                 in B_bias]
+      for i, pulse in enumerate(inputB):
+          pulse.inject_into(B[i:i+1])
+    \end{lstlisting}
+  \end{minipage}
+  \begin{minipage}{.495\textwidth}
+    \begin{lstlisting}[basicstyle={\footnotesize\ttfamily}]
+      weights = []
+      for i in xrange(30):
+          for j in xrange(30):
+              weights.append((i, j,
+                  np.dot(A_decoder[i], B_encoder[j]), 1.0))
+      connection = pyNN.Projection(
+          A, B, pyNN.FromListConnector(weights))
+      A.record('spikes')
+      B.record('spikes')
+      pyNN.run(1000)
+
+      # Decoded value of B
+      B_output = numpy.zeros(1000)
+      for i in xrange(1000):
+          spikes = B[i:i+1].getSpikes()[:,1]
+          spikes = (spikes / dt).astype('int')
+          B_output[spikes] += B_decoder[i]
+      # Filter with tau = 200 ms
+      decay = np.exp(-dt / 200)
+      B_output[0, :] *= (1 - decay)
+      for i in xrange(1, 1000):
+          B_output[i,:] = decay * B_output[i-1,:]
+                           + (1-decay) * B_output[i,:]
+    \end{lstlisting}
+  \end{minipage}
+\end{center}
+ \textbf{\refstepcounter{figure}\label{fig:pynn} Figure \arabic{figure}.}{
+   An implementation of the communication channel in PyNN.}
+\end{figure}
 
@@
 \begin{center}
-  Nengo Lorenz code listing
+  \begin{minipage}{0.43\textwidth}
+    \begin{lstlisting}[basicstyle={\footnotesize\ttfamily}]
+      import nengo
+      tau = 0.1; sigma = 10; beta = 8.0/3; rho = 28
+      model = nengo.Model('Lorenz attractor')
+      model.make_ensemble('State', nengo.LIF(2000),
+                            dimensions=3, radius=60)
+
+      def feedback(x):
+          dx0 = -sigma * x[0] + sigma * x[1]
+          dx1 = -x[0] * x[2] - x[1]
+          dx2 = x[0] * x[1] - beta * (x[2] + rho) - rho
+
+          return [dx0 * tau + x[0],
+                  dx1 * tau + x[1],
+                  dx2 * tau + x[2]]
+
+      model.connect('State', 'State',
+                      function=feedback, filter=tau)
+      model.probe('State', filter=tau)
+      model.probe('State.spikes')
+      sim = model.simulator()
+      sim.run(6)
+    \end{lstlisting}
+  \end{minipage}
+  \begin{minipage}{0.55\textwidth}
+    \includegraphics[width=0.95\textwidth]{lorenz}
+    \includegraphics[width=\textwidth]{lorenz_res}
+  \end{minipage}
 \end{center}
- \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{
-   Nengo Lorenz code listing}
+ \textbf{\refstepcounter{figure}\label{fig:lorenz} Figure \arabic{figure}.}{
+   Lorenz attractor model. (Left) Python code to create and run the model.
+   (Top right) Cartoon depiction of the model.
+   (Bottom right) Results of the model simulation. Lines depict
+   the node's output (Input panel) and the decoded value being represented
+   by ensembles (A and B panels). Light gray lines in A and B panels denote
+   spikes emitted by neurons in the ensembles. Each row depicts spikes
+   for one neuron.}
 \end{figure}
@@
 with the simulators that PyNN supports.
-
-\begin{figure}
-\begin{center}
-  PyNN Lorenz code listing
-\end{center}
- \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{
-   PyNN Lorenz code listing}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-  Lorenz attractor benchmarks
-\end{center}
- \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{
-   Lorenz benchmarks}
-\end{figure}
 
@@
 \begin{center}
-  Nengo CConv code listing
+  \includegraphics[width=0.8\textwidth]{cconv}
+  \begin{minipage}{0.53\textwidth}
+    \begin{lstlisting}[basicstyle={\footnotesize\ttfamily}]
+      import nengo
+      from nengo.networks import CircularConvolution
+      model = nengo.Model("Circular convolution")
+      model.make_node("Input A", [-0.21, 0.5, 0.12, 0.06])
+      model.make_ensemble("A", nengo.LIF(512), 4)
+      model.connect("Input A", "A")
+      model.make_node("Input B", [-0.18, 0.28, 0.18, -0.52])
+      model.make_ensemble("B", nengo.LIF(512), 4)
+      model.connect("Input B", "B")
+      cconv = model.add(CircularConvolution("Circular Convolution",
+          neurons=nengo.LIF(1032), dimensions=4))
+      model.connect("A", cconv.A)
+      model.connect("B", cconv.B)
+      model.make_ensemble("Result", nengo.LIF(512), 4)
+      model.connect(cconv, "Result")
+      model.probe("A", filter=0.02)
+      model.probe("B", filter=0.02)
+      model.probe("Result", filter=0.02)
+      sim = model.simulator()
+      sim.run(0.5)
+    \end{lstlisting}
+  \end{minipage}
+  \begin{minipage}{0.43\textwidth}
+    \includegraphics[width=\textwidth]{cconv_res}
+  \end{minipage}
 \end{center}
- \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{
-   Nengo CConv code listing}
+ \textbf{\refstepcounter{figure}\label{fig:lorenz} Figure \arabic{figure}.}{
+   Circular convolution model. (Top) Cartoon depiction of the model.
+   Note that much of the complexity is contained within the
+   circular convolution network; the transform matrices shown
+   are automatically generated by the network depending on
+   input dimensionality.
+   (Bottom left) Python code to create and run the model.
+
+   (Bottom right) Results of the model simulation. Lines depict
+   the node's output (Input panel) and the decoded value being represented
+   by ensembles (A and B panels). Light gray lines in A and B panels denote
+   spikes emitted by neurons in the ensembles. Each row depicts spikes
+   for one neuron.}
 \end{figure}
@@
 TODO: can we do circconv in PyNN?
-
-\begin{figure}
-\begin{center}
-  PyNN CConv code listing
-\end{center}
- \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{
-   PyNN CConv code listing}
-\end{figure}
 

### 2013-09-26 scripts/cconv.py

@@ -1,0 +1,76 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+import nengo
+from nengo.networks import CircularConvolution
+
+def circconv(a, b):
+    A = np.fft.fft(a)
+    B = np.fft.fft(b)
+    return np.fft.ifft(A * B).real
+
+a_in = np.asarray([-0.21, 0.5, 0.12, 0.06])
+b_in = np.asarray([-0.18, 0.28, 0.18, -0.52])
+
+model = nengo.Model("Circular convolution")
+model.make_node("Input A", a_in)
+model.make_ensemble("A", nengo.LIF(512), 4)
+model.connect("Input A", "A")
+
+model.make_node("Input B", b_in)
+model.make_ensemble("B", nengo.LIF(512), 4)
+model.connect("Input B", "B")
+
+cconv = model.add(CircularConvolution(
+    "Circular Convolution", neurons=nengo.LIF(1032), dimensions=4))
+model.connect("A", cconv.A)
+model.connect("B", cconv.B)
+
+model.make_ensemble("Result", nengo.LIF(512), 4)
+model.connect(cconv, "Result")
+
+model.probe("A", filter=0.02)
+model.probe("B", filter=0.02)
+model.probe("Result", filter=0.02)
+
+sim = model.simulator()
+sim.run(0.5)
+
+colors = ['r', 'g', 'b', 'm']
+
+actual_out = circconv(a_in, b_in)
+t = sim.data(model.t)
+
+def plot(actual, sim):
+    plt.gca().set_color_cycle(colors)
+    plt.gca().spines['top'].set_visible(False)
+    plt.gca().spines['left'].set_visible(False)
+    plt.gca().xaxis.set_ticks_position('bottom')
+    plt.gca().yaxis.tick_right()
+    plt.gca().yaxis.set_label_position("right")
+    plt.plot(t, sim)
+    for color, l in zip(colors, actual):
+        plt.axhline(l, color=color)
+    plt.xlim(0, 0.5)
+
+plt.figure(figsize=(5,5))
+plt.subplot(3, 1, 1)
+plot(a_in, sim.data("A"))
+plt.gca().spines['bottom'].set_visible(False)
+plt.xticks(())
+plt.ylabel("A")
+
+plt.subplot(3, 1, 2)
+plot(b_in, sim.data("B"))
+plt.gca().spines['bottom'].set_visible(False)
[…]

### 2013-09-26 scripts/comm_channel.py

@@ -1,0 +1,85 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import nengo
+from nengo.helpers import sorted_neurons, white_noise
+
+model = nengo.Model("Communication Channel")
+
+model.make_node("Input", output=white_noise(1, 5, seed=60))
+
+model.make_ensemble("A", nengo.LIF(30), 1)
+model.make_ensemble("B", nengo.LIF(30), 1)
+
+model.connect("Input", "A")
+model.connect("A", "B")
+
+model.probe("Input")
+model.probe("A", filter=0.01)
+model.probe("A.spikes")
+model.probe("B", filter=0.01)
+model.probe("B.spikes")
+
+sim = model.simulator()
+sim.run(1)
+
+t = sim.data(model.t)
+def sorted_spikes(pop):
+    indices = sorted_neurons(sim.model.get(pop))
+    spikes = [t[sim.data(pop + ".spikes")[:,i] > 0].flatten() for i in indices]
+    for ix in xrange(len(spikes)):
+        if spikes[ix].shape == (0,):
+            spikes[ix] = np.array([-1])
+    return spikes
+
+a_spikes = sorted_spikes("A")
+b_spikes = sorted_spikes("B")
+
+def adjust(ax):
+    ax.spines['top'].set_visible(False)
+    ax.spines['left'].set_visible(False)
+    ax.xaxis.set_ticks_position('bottom')
+    ax.yaxis.tick_right()
+    ax.set_yticks((-1, 0, 1))
+    ax.axis([0, 1, -1.25, 1.25])
+
+plt.figure(figsize=(4,5))
+ax = plt.subplot(3,1,1)
+ax.plot(t, sim.data("Input"), color='k')
+ax.text(0.5, 1.0, "Input", ha='center', va='center', fontsize=20)
+adjust(ax)
+ax.spines['bottom'].set_visible(False)
+ax.set_xticks(())
+
+ax = plt.subplot(3,1,2)
+ax.eventplot(a_spikes, colors=[(0.75,0.75,0.75)], rasterized=True)
+ax.spines['bottom'].set_visible(False)
+ax.spines['top'].set_visible(False)
+ax.spines['left'].set_visible(False)
+ax.set_yticks(())
+ax.axis([0, 1, 0, len(a_spikes)])
+
+ax = ax.twinx()
+ax.plot(t, sim.data("A"), color='k')
[…]

### 2013-09-26 scripts/lorenz.py

@@ -1,0 +1,79 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from mpl_toolkits.mplot3d import Axes3D
+
+import nengo
+from nengo.helpers import sorted_neurons
+
+tau = 0.1
+sigma = 10
+beta = 8.0/3
+rho = 28
+
+model = nengo.Model('Lorenz attractor')
+model.make_ensemble('State', nengo.LIF(2000), 3, radius=60)
+
+def feedback(x):
+    dx0 = -sigma * x[0] + sigma * x[1]
+    dx1 = -x[0] * x[2] - x[1]
+    dx2 = x[0] * x[1] - beta * (x[2] + rho) - rho
+
+    return [dx0 * tau + x[0],
+            dx1 * tau + x[1],
+            dx2 * tau + x[2]]
+
+model.connect('State', 'State',
+              function=feedback, filter=tau)
+model.probe('State', filter=tau)
+model.probe('State.spikes')  # Very expensive!!
+
+sim = model.simulator()
+sim.run(6)
+
+t = sim.data(model.t)
+def sorted_spikes(pop):
+    # 2000 neurons... let's just get 10% (200)
+    indices = sorted_neurons(sim.model.get(pop), iterations=250)[::80]
+    spikes = [t[sim.data(pop + ".spikes")[:,i] > 0].flatten() for i in indices]
+    for ix in xrange(len(spikes)):
+        if spikes[ix].shape == (0,):
+            spikes[ix] = np.array([-1])
+    return spikes
+
+state_spikes = sorted_spikes("State")
+#state_spikes = [np.linspace(0, 6, 6 * 100)]
+
+plt.figure(figsize=(7,4.55))
+ax = plt.subplot2grid((2,3), (0,0), colspan=2, rowspan=2, projection='3d')
+ax.plot(sim.data('State')[:,0], sim.data('State')[:,1], sim.data('State')[:,2],
+        color='k')
+# ax.set_xticklabels(())
+# ax.set_yticklabels(())
+# ax.set_zticklabels(())
+ax.dist = 8.5
+
+ax = plt.subplot2grid((2,3), (0,2))
+ax.plot(t, sim.data('State'))
+ax.set_ylabel("Amplitude")
+ax.spines['top'].set_visible(False)
+ax.spines['left'].set_visible(False)
+ax.spines['bottom'].set_visible(False)
+ax.set_xticks(())
[…]

### 2013-09-26 nengo.tex
Figures mostly done

@@
 
-%% This should be a paper of its own.
-
-%% \subsection{Semantic Pointer Architecture (SPA)}
-
-%% TODO; should we even include this section?
-%% It could probably be a separate paper on its own.
+\begin{figure}
+\begin{center}
+  \includegraphics[width=0.19\textwidth]{nef_summary_enc}
+  \includegraphics[width=0.19\textwidth]{nef_summary_dec}
+  \includegraphics[width=0.399\textwidth]{nef_summary_trans}
+  \includegraphics[width=0.199\textwidth]{nef_summary_dyn}
+\end{center}
+ \textbf{\refstepcounter{figure}\label{fig:nef} Figure \arabic{figure}.}{
+   Summary of the Neural Engineering Framework.}
+\end{figure}
 
@@
 
-\begin{figure}
-\begin{center}
-  Nengo high-level object hierarchy
-\end{center}
- \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{
-   Nengo high-level object hierarchy}
-\end{figure}
-
 A Nengo model is a collection
@@
 (see Figure~\ref{TODO}).
-
-\begin{figure}
-\begin{center}
-  Network code listing
-\end{center}
- \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{
-   Network code listing}
-\end{figure}
 
@@
  \textbf{\refstepcounter{figure}\label{fig:cchannel} Figure \arabic{figure}.}{
-   Communcation channel model. (Top) Cartoon depiction of the model.
+   Communcation channel model. (Top) Diagram of the model.
    (Bottom left) Python code to create and run the model.
@@
    Lorenz attractor model. (Left) Python code to create and run the model.
-   (Top right) Cartoon depiction of the model.
+   (Top right) Diagram of the model.
    (Bottom right) Results of the model simulation. Lines depict
@@
  \textbf{\refstepcounter{figure}\label{fig:lorenz} Figure \arabic{figure}.}{
-   Circular convolution model. (Top) Cartoon depiction of the model.
+   Circular convolution model. (Top) Diagram of the model.
    Note that much of the complexity is contained within the
@@
 
+\section{Benchmarks}
+
+Some stuff about benchmarking
+
 \begin{figure}
 \begin{center}
-  CConv benchmarks
+  \includegraphics[width=0.5\textwidth]{benchmark}
 \end{center}
- \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{
-   CConv benchmarks}
+ \textbf{\refstepcounter{figure}\label{fig:benchmarks} Figure \arabic{figure}.}{
+   Benchmarks}
 \end{figure}

### 2013-09-26 scripts/base.py

@@ -1,0 +1,14 @@
+import numpy as np
+
+from nengo.helpers import sorted_neurons
+
+def sorted_spikes(sim, pop, iterations=100, every=None):
+    t = sim.data(sim.model.t)
+    indices = sorted_neurons(sim.model.get(pop), iterations=iterations)
+    if every is not None:
+        indices = indices[::every]
+    spikes = [t[sim.data(pop + ".spikes")[:,i] > 0].flatten() for i in indices]
+    for ix in xrange(len(spikes)):
+        if spikes[ix].shape == (0,):
+            spikes[ix] = np.array([-1])
+    return spikes

### 2013-09-26 scripts/cconv.py

@@ -36,38 +36,37 @@ model.probe("Result", filter=0.02)
 sim = model.simulator()
 sim.run(0.5)
 
-colors = ['r', 'g', 'b', 'm']
-
 actual_out = circconv(a_in, b_in)
-t = sim.data(model.t)
 
-def plot(actual, sim):
+def plot(actual, ens):
+    color_cycle = plt.gca()._get_lines.color_cycle
+    colors = [next(color_cycle) for _ in xrange(4)]
     plt.gca().set_color_cycle(colors)
     plt.gca().spines['top'].set_visible(False)
     plt.gca().spines['left'].set_visible(False)
     plt.gca().xaxis.set_ticks_position('bottom')
     plt.gca().yaxis.tick_right()
     plt.gca().yaxis.set_label_position("right")
-    plt.plot(t, sim)
+    plt.plot(sim.data(model.t), sim.data(ens))
     for color, l in zip(colors, actual):
         plt.axhline(l, color=color)
     plt.xlim(0, 0.5)
 
 plt.figure(figsize=(5,5))
 plt.subplot(3, 1, 1)
-plot(a_in, sim.data("A"))
+plot(a_in, "A")
 plt.gca().spines['bottom'].set_visible(False)
 plt.xticks(())
 plt.ylabel("A")
 
 plt.subplot(3, 1, 2)
-plot(b_in, sim.data("B"))
+plot(b_in, "B")
 plt.gca().spines['bottom'].set_visible(False)
 plt.xticks(())
 plt.ylabel("B")
 
 plt.subplot(3, 1, 3)
-plot(actual_out, sim.data("Result"))
+plot(actual_out, "Result")
 plt.xlabel("Time (s)")
 plt.ylabel("Result")
 

### 2013-09-26 scripts/comm_channel.py

@@ -1,7 +1,10 @@
 import numpy as np
 import matplotlib.pyplot as plt
+
 import nengo
-from nengo.helpers import sorted_neurons, white_noise
+from nengo.helpers import white_noise
+
+from base import sorted_spikes
 
 model = nengo.Model("Communication Channel")
 
@@ -22,17 +25,8 @@ model.probe("B.spikes")
 sim = model.simulator()
 sim.run(1)
 
-t = sim.data(model.t)
-def sorted_spikes(pop):
-    indices = sorted_neurons(sim.model.get(pop))
-    spikes = [t[sim.data(pop + ".spikes")[:,i] > 0].flatten() for i in indices]
-    for ix in xrange(len(spikes)):
-        if spikes[ix].shape == (0,):
-            spikes[ix] = np.array([-1])
-    return spikes
-
-a_spikes = sorted_spikes("A")
-b_spikes = sorted_spikes("B")
+a_spikes = sorted_spikes(sim, "A")
+b_spikes = sorted_spikes(sim, "B")
 
 def adjust(ax):
     ax.spines['top'].set_visible(False)
@@ -44,7 +38,7 @@ def adjust(ax):
 
 plt.figure(figsize=(4,5))
 ax = plt.subplot(3,1,1)
-ax.plot(t, sim.data("Input"), color='k')
+ax.plot(sim.data(model.t), sim.data("Input"), color='k')
 ax.text(0.5, 1.0, "Input", ha='center', va='center', fontsize=20)
 adjust(ax)
 ax.spines['bottom'].set_visible(False)
@@ -56,10 +50,10 @@ ax.spines['bottom'].set_visible(False)
 ax.spines['top'].set_visible(False)
 ax.spines['left'].set_visible(False)
 ax.set_yticks(())
-ax.axis([0, 1, 0, len(a_spikes)])
+ax.axis([0, 1, -0.5, len(a_spikes)-0.5])
 
 ax = ax.twinx()
-ax.plot(t, sim.data("A"), color='k')
+ax.plot(sim.data(model.t), sim.data("A"), color='k')
 ax.text(0.5, 1.0, "A", ha='center', va='center', fontsize=20,
         bbox=dict(ec='none', fc='w', alpha=0.8))
 ax.spines['bottom'].set_visible(False)
@@ -72,11 +66,11 @@ ax.spines['top'].set_visible(False)
[…]

### 2013-09-26 scripts/lorenz.py

@@ -3,7 +3,8 @@ import matplotlib.pyplot as plt
 from mpl_toolkits.mplot3d import Axes3D
 
 import nengo
-from nengo.helpers import sorted_neurons
+
+from base import sorted_spikes
 
 tau = 0.1
 sigma = 10
@@ -30,17 +31,7 @@ model.probe('State.spikes')  # Very expensive!!
 sim = model.simulator()
 sim.run(6)
 
-t = sim.data(model.t)
-def sorted_spikes(pop):
-    # 2000 neurons... let's just get 10% (200)
-    indices = sorted_neurons(sim.model.get(pop), iterations=250)[::80]
-    spikes = [t[sim.data(pop + ".spikes")[:,i] > 0].flatten() for i in indices]
-    for ix in xrange(len(spikes)):
-        if spikes[ix].shape == (0,):
-            spikes[ix] = np.array([-1])
-    return spikes
-
-state_spikes = sorted_spikes("State")
+state_spikes = sorted_spikes(sim, "State", iterations=250, every=80)
 #state_spikes = [np.linspace(0, 6, 6 * 100)]
 
 plt.figure(figsize=(7,4.55))
@@ -53,7 +44,7 @@ ax.plot(sim.data('State')[:,0], sim.data('State')[:,1], sim.data('State')[:,2],
 ax.dist = 8.5
 
 ax = plt.subplot2grid((2,3), (0,2))
-ax.plot(t, sim.data('State'))
+ax.plot(sim.data(model.t), sim.data('State'))
 ax.set_ylabel("Amplitude")
 ax.spines['top'].set_visible(False)
 ax.spines['left'].set_visible(False)

### 2013-09-26 scripts/nef_summary.py

@@ -1,0 +1,216 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+import nengo
+from nengo.objects import Uniform
+
+from base import sorted_spikes
+
+## Encoding
+
+model = nengo.Model("Encoding")
+model.make_node("Input signal", output=lambda t: t * 2 - 1)
+model.probe("Input signal")
+
+intercepts = np.linspace(-0.9, 0.9, 8)
+encoders = np.asarray([1,-1,1,-1,1,-1,1,-1])
+intercepts *= encoders
+encoders.shape = (8,1)
+model.make_ensemble("A", nengo.LIF(8), 1, intercepts=intercepts,
+                    max_rates=Uniform(80, 100), encoders=encoders)
+model.connect("Input signal", "A")
+model.probe("A.spikes")
+model.probe("A", filter=0.03)
+
+model.make_node("Dummy", output=lambda t: 0)
+cos = model.connect("A", "Dummy", function=np.cos, filter=0.03)
+model.probe(cos)
+
+## Encoding plot
+
+sim = model.simulator()
+sim.run(1)
+
+t = sim.data(model.t)
+spikes = [t[sim.data("A.spikes")[:,i] > 0].flatten()
+          for i in xrange(model.get("A").n_neurons)]
+A = sim.model.get("A")
+A.eval_points.sort(axis=0)
+J = np.dot(A.eval_points, A.encoders.T)
+activities = A.neurons.rates(J)
+
+plt.figure(figsize=(4,9))
+plt.subplot(3,1,1)
+plt.plot(sim.data(model.t)[1:], sim.data("Input signal")[1:], lw=2)
+plt.title("Input signal")
+plt.xlim(0, 1)
+
+plt.subplot(3,1,2)
+plt.plot(A.eval_points, activities, lw=2)
+plt.xlabel("Input signal")
+plt.ylabel("Firing rate (Hz)")
+
+plt.subplot(3,1,3)
+color_cycle = plt.gca()._get_lines.color_cycle
+colors = [next(color_cycle) for _ in xrange(model.get("A").n_neurons)]
+plt.axis([0, 1, model.get("A").n_neurons-0.5, -0.5])
+plt.eventplot(spikes, colors=colors)
+
+plt.tight_layout()
+plt.savefig("../figures/nef_summary_enc.pdf")
+
+## Decoding
[…]

### 2013-09-27 nengo.tex
new benchmark figure

@@
 
-Some stuff about benchmarking
-
 \begin{figure}
 \begin{center}
-  \includegraphics[width=0.5\textwidth]{benchmark}
+  \includegraphics[width=0.5\textwidth]{ocl_vs_ref}
 \end{center}
- \textbf{\refstepcounter{figure}\label{fig:benchmarks} Figure \arabic{figure}.}{
-   Benchmarks}
+ \textbf{
+     \refstepcounter{figure}\label{fig:ocl_vs_ref}
+     Figure \arabic{figure}.}{
+     Five-hundred-dimensional circular convolution networks were some of the most
+     computationally demanding elements of the Spaun cognitive architecture.
+     Simulations of circular convolution networks of various sizes illustrate the speed of our re-engineered simulators.
+     A straightforward reference implementation in Python is sufficient for small networks,
+     while for larger ones the new OpenCL-powered simulator runs up to 200$\times$ faster,
+     In particular the Radeon 7970 GPU can simulate 500-dimensional convolution networks faster than real time (0.95 seconds for a 1-second simulation).
+     Also, although both are multithreaded, our new implementation delivers better CPU performance than the previous JavaNengo implementation.
+     }
 \end{figure}

### 2013-09-27 nengo.tex
Major rewrites

@@
 theories of the brain.
-Much work has been put into simulators
-that attempt to recreate neural
+Much work has been put into neural simulators
+that attempt to recreate neuroscientific
 data in precise detail,
@@
 Similarly, theoretical frameworks
-that involve a generic algorithm
+consisting of a generic algorithm
 replicated many times
@@
 have not been shown to solve problems
-outside of those that
-the generic algorithm alone can solve.
+that the generic algorithm alone cannot solve.
 
@@
 representing the task to be performed
-and the details of that task
-(e.g., $A3[1234]$ instructs Spaun
-to remember the list $1234$).
+and the details of that task;
+for example, sequentially presenting images
+containing the characters $A3[1234]$ instructs Spaun
+to remember the list $1234$.
 It then provides its answer by
 sending motor commands to a simulated arm model,
-tracing out digits
-(e.g., $1234$ if asked to recall the remembered list).
+tracing out digits.
+If asked to recall the remember list,
+it would trace out the digits $1234$.
 
@@
 rather than defining a single algorithm
-and hoping that replicating it
+and hypothesizing that replicating it
 many times will enable these behaviors.
@@
 In implementing Spaun,
-we have provided strong support
+we provided strong support
 that the NEF is beginning to fill the theory void in neuroscience.
-However, we have also pushed Nengo
+However, in doing so, we pushed Nengo
 to the limits of its ability to construct and simulate
@@
 we have begun the next generation
-of Nengo tools with a focus on simplicity,
-extensibility, and speed.
+of Nengo with a focus on speed,
+extensibility, and simplicity.
 
@@
   & Simulates models constructed with PyNengo. \\\botrule
-\end{tabular}}{}
+\end{tabular}}{Code repositories avilable at \url{http://github.com/ctn-waterloo/}}
 \end{table}
@@
 the methods of the NEF
-as one of many different
-methods of creating neural models.
+as one of several
+methods for creating neural models.
 As a result, its implementation
@@
 to extend those NEF components.
-The addition of synaptic plasticity rules,
-for example, involved adding or significantly modifying
-25 Java classes \cite{TODO}.\footnote{Changelog at
-  \url{https://github.com/ctn-waterloo/nengo_1.4/commit/93eee51147545f7ce30ae9919809e3072dc15728}.}
+As a representative example,
+the most often used class for making
+NEF models, \texttt{NEFEnsemble},
+extends from five ancestor classes
+and implements seven interfaces.
+Adding support for synaptic plasticity
+in \texttt{NEFEnsemble}
+involved adding or significantly modifying
+25 Java classes \cite{TODO}.
 
@@
 to add new functionality
-(e.g., new neuron models).
+(e.g., to add new neuron models).
 This is because Jython makes it easy
@@
 they used the scripting interface or not.
-As the set of neural modellers
-that are proficient with both Java and Python
-syntax is small,
+As there are few neural modellers
+that are proficient with both Java and Python syntax,
 the scripting interface still lacks
@@
 Finally, perhaps the most significant limitation
-of the JavaNengo tools is its performance.
+of the JavaNengo tools is their performance.
 While Java has good support for threads,
@@
 which can perform the types of computations
-that these neural models require quickly and in parallel.
+that neural models require quickly and in parallel.
 
@@
 significant enough changes
-that we decided to start from scratch instead.
+that we started from scratch instead.
 Since the scripting interface
 already used Python syntax,
-and we were using Python
-and its scientific libraries
-for analyzing simulation results,
+and since Python is widely used
+for neuroscientific data analysis,
 we targeted the CPython interpreter
-as the platform for the new generation of Nengo tools.
+as the platform for the next generation of Nengo.
 
@@
 
-PyNengo, which we describe in more detail
-in the remainder of this paper,
+PyNengo,
+the subject of the remainder of this paper,
 is a Python package for
@@
 with other scientific Python tools.
-Nengo models can used
-in arbitrary CPython programs,
+Arbitrary CPython programs
+can use Nengo models,
 opening up possibilities
-for using neurally inspired algorithms
+for using neurally implemented algorithms
 in applications and games.
 
-PyNengo's object model has been radically simplified,
+PyNengo's object model is simple and minimal,
 making it easy to
-document, test, and modify those objects.
+document, test, and modify.
 It also includes a scripting interface to those objects
@@
 to allow for easier incorporation
-of new neuron models and learning rules.
+of user-defined networks,
+new neuron models, and learning rules.
 Additionally, PyNengo's scripting interface
@@
 
-Model creation and simulation
-have been completely decoupled,
+Model creation and simulation are decoupled,
 allowing for models to be exported
@@
 PyNengo does not currently
-replace all of the JavaNengo tools.
+replace all of JavaNengo's functionality.
+While it can simulate the vast majority
+of models that JavaNengo can simulate,
+some functionality is still in development.
 Most notably, there is no graphical interface
-to PyNengo, and therefore little ability to interact
-with simulations as they are running.
+to PyNengo, and therefore simulations
+cannot be visualized while they are running.
 However, we are confident that these tools
@@
 
-We believe that PyNengo overcomes
-all of the significant issues
-with JavaNengo.
-It can simulate the majority of models
-currently simulated by JavaNengo,
-using 11\% as many lines code,
+PyNengo overcomes all of JavaNengo's limitations.
+It can simulate most models
+currently simulated by JavaNengo
+using 11\% as many lines of code,
 and depending on
@@
 The OpenCL simulator can simulate
-a large scale network TODO times
+a large model TODO times
 faster than JavaNengo's simulator.
 This, along with a codebase
-that is easier to understand and extend,
+that is well tested and
+easy to understand and extend,
 provides a platform
@@
 
-TODO; make sure it meshes with 3.1's operators
-
-\begin{equation} \label{TODO}
-  TODO
-\end{equation}
+The Neural Engineering Framework (NEF; \cite{TODO})
+proposes three principles
+that enable the construction
+of large-scale neural models.
+
+\begin{enumerate}
+  \item \textbf{Representation:} A population of neurons
+    collectively represents information in the form of
+    time-varying vectors of real numbers.
+  \item \textbf{Transformation:} Functions on those vectors
+    are computed by the connections between populations.
+  \item \textbf{Dynamics:} The vectors represented
+    by neural populations can be considered state variables
+    in a dynamical system, and can be analyzed using control theory.
+\end{enumerate}
 
@@
 
-\section{Model definition}
-
-One of the strengths of the NEF
-is that it allows neural modellers to think
-at a higher level
-than is normally associated with neural models.
-With the NEF, a modeller decides
-what information is being represented
-by populations of neurons and how that information
-is being transformed through the connections
-between populations;
-this contrasts with modelling decisions
-that are typically thought to be important,
-such as the morphology of each neuron,
-and the distribution of connection weights
-between populations.
-Instead, these decisions are made---often
-automatically---as a direct result
-of the information
-being represented and the transformations
-being performed on that information.
-This situates Nengo as a
-``neural compiler'' that can
-take a functional model description
-and produce a low-level neural model
-that can be simulated
-with any neural simulation software.
-
-To emphasize Nengo's role as a neural compiler,
-Nengo allows models to be defined
-in a simulator-agnostic manner,
-allowing for functionally identical models to be run
-on different software and hardware platforms.
-
-\subsection{Nengo objects} \label{sec:objects}
+\paragraph{Representation}
+Information is encoded by populations of neurons.
+We consider that information
+to be time-varying vectors of real numbers,
+in order to reason with that information
+using conventional mathematics.
+By doing this, we can \textit{encode}
+those vectors by injecting
+specific amounts of current into
+single neuron models based on
+the vector being encoded.
+This drives the neuron,
+causing it to spike.
+We can then estimate
+the originally encoded vector
+through a \textit{decoding} process.
+This idea is similar to population coding
+\cite{TODO}, but generalized
+to vectors of arbitrary length.
+
+Figure~\ref{fig:nef} depicts
+the encoding (A) and decoding (B) process.
+In the encoding process, the input signal drives
+each neuron based on its \textit{tuning curve},
+which describes how much
+that neuron fires for a given input signal.
+The tuning curve summarizes the gain
+of a neuron (how quickly the activity rises),
+the bias (the activity of a neuron given no signal),
+and the encoding weight
+(the part of the vector space
+that causes the neuron to fire most strongly).
+This results in a predictable pattern of firing
+given the input signal.
+
+In the decoding process,
+the spike trains are first filtered
+with an exponentially decaying filter.
+Those filtered spike trains are summed together
+with weights that are determined
+by solving a least-squares minimization problem.
+In the bottom panel of Figure~\ref{fig:nef}B,
+the red line is the filtered spiking activity
+weighted by decoding weights
+that minimize the difference between
+the actual encoded vector
+and the decoded estimate.
+The purple line is the filtered spiking activity
+weighted by decoding weights
+that minimize the different between
+the cosine of the actual encoded vector
+and the decoded estimate.
+Note that these decoding weights are not determined
+based on the input signal;
+instead, the minimization occurs
+on points that are randomly sampled
+in the vector space
+that the population represents.
+
+\paragraph{Transformation}
+Neurons communicate through
+unidirectional connections called synapses.
+When a neuron spikes,
+it releases neurotransmitter across the synapse,
+which causes some amount of current
+to be imparted in the downstream neuron.
+Many factors affect the
+amplitude of the current imparted;
+we summarize those factors
+in a scalar connection weight
+representing the strength
+of the connection between two neurons.
+In order to compute any function,
+we can set the connection weights between
+two populations to be the product of
+the decoding weights for that function
+for neurons in the first population
+and the encoding weights
+for the downstream neuron.
+
+In Figure~\ref{fig:nef}C,
+a sine wave is transformed twice.
+Between populations A and B,
+the signal is negated.
+Between populations B and C,
+the signal is squared.
+
+Note that, in practice, we rarely use
+the full connection weight matrix,
+and instead store
+the encoding and decoding weights
+(i.e., the factors of the connection weight matrix).
+This provides significant
+space and time savings.
+
+\paragraph{Dynamics}
+While doing feedforward transformations
+on vectors is sufficient to model
+many neural systems,
+many systems, notably memory systems,
+require persistent activity through recurrent connections.
+When recurrent connection are introduced,
+the vectors that populations represent
+can be thought of as state variables
+in a dynamical system.
+Neural systems can therefore
+be built and analyzed using
+the methods of control theory.
+
+In Figure~\ref{fig:nef}D,
+a population represents a two-dimensional vector,
+and is recurrently connected
+with negative feedback between dimensions,
+resulting in a harmonic oscillator.
+
+\paragraph{Nengo}
+Large models can be built
+by thinking of the principles of the NEF
+as building blocks that can be put together
+to describe neural systems.
+The goal of Nengo is to allow
+modellers to describe neural systems
+in terms of what information is being represented,
+how that information is transformed,
+and what dynamics are required
+to perform cognitively relevant
+functions with biologically constrained
+neural components.
+Those descriptions are then
+translated to a network
+of interconnected neurons.
+This situates Nengo
+as a ``neural compiler''
+that translates
+a high-level functional model
+to a low-level neural model
+that implements the high-level function.
+
+\section{PyNengo object model}
+
+\subsection{Model}
 
 A Nengo model is a collection
-of Nengo objects connected together,
-and probed in order to collect data
-during simulation.
+of Nengo objects that represent information
+and connections that transform information.
+Objects and connections can be probed
+in order to collect data during simulation.
+
 The \texttt{Model} object is primarily a container
 for Nengo objects,
-but also provides methods
-that simplify creating, connecting,
-and probing these objects.
-Basic Nengo models
+but also doubles as a scripting interface
+that simplifies creating, connecting,
+and probing objects.
+Basic models
 can be created solely through interacting
-with the \texttt{Model} object.
-For more advanced models,
-these objects can be instantiated
-directly and then added to the model.
-
-There are two types of objects
-defined in the main scripting interface.
-
-\paragraph{Ensembles}
-The most common is the \texttt{Ensemble},
-which is a population of neurons
+with the \texttt{Model} object
+(see sections \ref{sec:comm-channel}--\ref{sec:lorenz}
+for examples).
+In advanced models,
+objects are instantiated
+and then added to the model
+(see section \ref{sec:cconv} for an example).
+
+\subsection{Ensemble}
+
+An \texttt{Ensemble} is
+a population of neurons
 that represents information
 in the form of a real-valued vector.
-As described in \eqref{TODO},
-this can be done
-with any neuron model, $G$.
 An ensemble must be given as arguments
@@
 however it sees fit.
+The neuron model used by the population
+is easily changed, as long as the simulator
+is aware of the neuron model description object.
 
 Other attributes of the \texttt{Ensemble},
-such as its encoding weights, $\mathbf{e_i}$,
-the maximum firing rate of the neurons,
-and so on, can also be specified.
-This is usually done according
-to what is known about
-the neurobiology of the area being modelled;
-some areas of the brain regularly spike
-at rates above 200Hz,
-while others rarely spike above 10Hz.
+such as its encoding weights,
+the maximum firing rate of its neurons,
+and so on, can also be specified
+either as keyword arguments
+to the \texttt{Ensemble} constructor,
+or by setting an attribute on the instantiated object.
+While an ensemble makes a hypothesis
+about the information being represented by neurons,
+these additional attributes
+also allow modellers to ensure that
+the ensemble stays
+within neurobiological constraints.
 If these attributes are not set,
@@
 from distributions that describe
-what is known about pyramidal cells in neocortex.
-
-\paragraph{Nodes}
+neocortical pyramidal cells.
+
+\subsection{Node}
+
 A \texttt{Node} tracks any information
@@
 in which a neural model exists.
-More practically,
-nodes can be used to directly compute
-parts of a neural model
-for debugging purposes.
 
@@
 
-Nodes allows many of the special-purpose
-routines that are necessary in many real-world
+Nodes allow many of the special-purpose
+routines that are necessary in real-world
 models to be integrated
 as core components of a Nengo model.
-
-\paragraph{Connections}
+This makes Nengo models more explicit,
+and enables simulators
+to control and optimize node execution.
+
+\subsection{Connection}
 
@@
 a connection contains symbolic information
-about how two objects are to be connected
-together.
+about how two objects are to be connected together.
 
@@
 This connection implements
-the transformation principle equations,
-\eqref{TODO}.
+the NEF's transformation principle.
 In other words, the \texttt{DecodedConnection}
@@
 even though the underlying implementation
-still connects populations of neurons
-together with connection weights.
-%TODO: reword? is it important that it's a factored weight matrix
-
-\paragraph{Probes}
-
-Ideally, the entire state of a simulation
-would be recorded on each timestep,
-and stored with no overhead to the simulator.
-In practice, recording data
-is a costly part of most simulations,
-and often what is simulated must also change
-if the signals being probed must be filtered.
-For this reason, probes are also
-a core component of a Nengo model.
-
-In general, Nengo objects and connections
+can translate that connection
+to connection weights.
+
+The ability to connect neurons together
+with synaptic connection weights is still possible
+with the \texttt{NonlinearityConnection}.
+Both of these connection types
+can be temporally filtered,
+and the weights involved in the connection
+(decoding weights for \texttt{DecodedConnection},
+synaptic connection weights for \texttt{NonlinearityConnection})
+can be modified over time with learning rules.
+
+In many cases, the type of the connection
+is not explicitly specified by the modeller.
+Instead, it is inferred by the arguments
+to \texttt{Model}'s \texttt{connect} function.
+However, that function is just a simple wrapper
+of each connection object's constructor,
+so complex models can use
+the appropriate connection,
+or even implement their own.
+
+\subsection{Network}
+
+A network is a collection of ensembles and nodes
+connected together in a particular way.
+It is primarily a way of grouping together
+a common set of connected objects,
+and makes explicit the purpose
+of the network and how it interacts
+with the rest of the model
+(see section \ref{sec:ccon} for a usage example).
+
+Networks are designed to be
+subclasses by users,
+in order to group together
+parts of a model.
+In most cases,
+the code that creates and connects
+several objects in the model can be
+moved into a \texttt{Network}
+subclass' \texttt{make}
+function with only minor changes.
+
+Networks can vary dramatically in size,
+and can be composed of other networks.
+The \texttt{Integrator} network, for example,
+is composed of only one recurrently connected ensemble.
+By encapsulating that logic in a network,
+the purpose of that ensemble is made explicit.
+On the other end of the spectrum,
+the \texttt{BasalGanglia} network,
+is composed of five groups of ensembles
+connected with several specific functions
+that together implement a ``winner take all'' circuit.
+Encapsulating that logic
+makes a complicated section of code
+easy to include in many different models.
+
+\subsection{Probe}
+
+A probe is used to denotes a signal
+that will be recorded
+throughout the simulation.
+Nengo objects and connections
 each have several signals that can be probed.
-For example, an ensemble's decoded value,
-or the underlying neurons' membrane voltages
-can be probed, among other signals.
-The underlying logic for the probe
-leverages connection logic,
-making it easy to probe any part of an object.
-
-Since probes are also described
+An ensemble's decoded value
+and its neurons' spikes
+can be probed, among others.
+
+Like nodes, a probe could be implemented
+outside of the model.
+However, doing so requires detailed knowledge
+of the simulator,
+and can incur significant overhead
+if not implemented carefully.
+For these reasons, we have made probes
+a core component of a Nengo model,
+and are therefore explicit
+and optimizable.
+Further, since probes are described
 at a symbolic level,
@@
 
-\section{Model simulation} \label{sec:simulators}
-
-Once a model has been fully defined,
-a simulator can be created to
-run that model.
-Simulators can be created
-that work directly with the
-objects described in section \ref{sec:objects}.
-However, in the simulators that we have implemented
-as reference, the model goes through a
-two-stage build process in order to
-simulate models as fast as possible.
-In the first stage,
-the model is ``built;''
-i.e., the objects in the model are mapped to a simpler
-set of objects that represent the computations
-that the simulator must perform.
-Those computations are analyzed in order to
-schedule them efficiently.
-In the second stage,
-the simulator performs the computations
-that have been scheduled on each timestep.
-
-This two-stage process has enabled us
-to implement a more efficient simulator
-that is only aware of
-the objects that are created
-during the build process.
-In the future,
-new simulators can target either
-the high level Nengo objects,
-or the low level simulator objects
-depending on the constraints of the platform.
-
-\subsection{Reference build process}
+\section{PyNengo simulators} \label{sec:simulators}
 
@@
  \textbf{\refstepcounter{figure}\label{fig:sim} Figure \arabic{figure}.}{
-   Sim}
+   The first two timesteps of a simple simulation.}
 \end{figure}
 
-The first stage of the reference implementation
-maps all of the Nengo objects
-down to a set of signals and operators.
+Decoupling model creation and simulation
+enables PyNengo simulators
+to allocate memory and schedule computations
+in the most efficient ways possible.
+Simulators are given a \texttt{Model}
+as an argument;
+this \texttt{Model} is a symbolic description,
+which a simulator can assume will not change
+over the course of a simulation.
+The simulator, however,
+can modify the model as it sees fit;
+in most cases, this means that the simulator
+will fill in many of the details
+that the symbolic description
+has allowed to be generated randomly.
+
+We have implemented
+a reference simulator using NumPy
+for vectorized computations,
+and an OpenCL simulator
+using PyOpenCL to parallelize
+computations on GPUs and multicore CPUs.
+In the remainder of this section,
+we will describe
+the reference simulator implementation;
+the OpenCL simulator shares many
+of the reference simulator's architectural choices,
+but is heavily optimized in ways
+that cannot be adequately described
+with the space available.
+Additionally, it should be noted that
+any simulator that takes a \texttt{Model}
+as input and exposes a \texttt{step}
+function can be used in PyNengo;
+the reference simulator
+is provided as an example
+for simulator designers,
+not as a specification.
+
+\subsection{Reference simulator}
+
+Before the first timestep, the reference simulator
+\begin{enumerate}
+  \item translates high-level objects to
+    a set of signals and operators,
+  \item allocates NumPy arrays for each signal, and
+  \item sorts operators based on a dependency graph.
+\end{enumerate}
+On each timestep, the reference simulator
+\begin{enumerate}
+  \item computes each operator in order, and
+  \item copies probed signals to a separate buffer.
+\end{enumerate}
+Figure~\ref{fig:sim} depicts
+the state of the reference simulator
+after two timesteps of a simple model.
 
 \paragraph{Signals}
+
 A \texttt{Signal} represents any number that
@@
 signal that will be encoded
-to input currents (see equation \eqref{TODO}),
-and the encoding weights
-($\mathbf{e_i}$; see equation \eqref{TODO}).
+to input currents,
+and the encoding weights.
 It also contains a neural population,
 which contains signals that represent
-input currents
-($J(t)$; see equation \eqref{TODO}),
-bias currents ($J^{bias}$; \eqref{TODO}),
-membrane voltages ($V$; \eqref{TODO}),
-and refractory times for each cell.
-These signal are symbolic,
-and contain only information
-about the size of the signal that will be represented,
-and optionally an initial value.
-Simulators may therefore represent
-these signals however they want internally.
-
-In some cases, determining
-the initial values of some signals
-may involve some computation.
-In particular, solving for decoding weights
-(equation \eqref{TODO}) requires
-solving a least-squares minimization problem.
-This computation is done in this stage
-of the build process,
-as it is computationally intensive,
-and may be implemented
-differently by some simulators.
-
-An important feature of signals
-is that parts of a signal
-can be accessed with Python slice syntax,
-which results in a view of the original signal.
-For example, when connecting to
-a subset of a population of neurons,
-current would be injected
-in a view of the overall input current signal
-for that population,
-rather than creating a new signal,
-or partitioning the population's signal
-into smaller chunks.
-Grouping related signals together
-and exposing them through views
-also provides hints to the simulator
-that can be used to organize signals
-efficiently in memory.
+input currents, bias currents,
+membrane voltages, and refractory times for each cell.
+
+As can be seen in Figure~\ref{fig:sim},
+the signals used in a Nengo simulation
+can be conceptually grouped into
+those that track low-level neural signals,
+and those that track high-level signals
+defined by the NEF.
+Most neural simulators only track
+low-level signals.
+Operators commonly map
+between related low- and high-level signals.
 
 \paragraph{Operators}
+
 Operators represent computations
@@
 operations are necessary for simulation.
-\begin{enumerate}
-  \item \textbf{Reset} assigns a constant value to a signal.
-    This is typically used to set a signal to 0
-    at the start of a timestep.
-  \item \textbf{Copy} assigns the value
-    of one signal to another signal.
-    This is typically used instead of \textbf{Reset}
-    in situations where the signal's value
-    at the start of a timestep should
-    be the value of another signal.
-  \item \textbf{DotInc} increments a signal by
-    the dot product of two signals;
-    i.e., $Y \leftarrow Y + AB$.
-    This is a common operation
-    done in many situations,
-    such as decoding a signal (see equation \eqref{TODO}).
-  \item \textbf{ProdUpdate} scales a signal
-    and adds the dot product of two signals to it;
-    i.e., $Y \leftarrow BY + AX$.
-    This is done when temporally filtering a signal.
-  \item \textbf{Nonlinearities} include any
-    user-defined function computed by a node,
-    and the update equation for each neuron model.
-\end{enumerate}
-
-It is important to note that all operators
-excepts for the nonlinearities are linear,
-and therefore can share an implementation
-in an underlying simulator,
-allowing those operations to happen in parallel.
-
-Operators are created
-by Nengo objects during the build process.
-For example, an ensemble creates
-a \textbf{Reset} operator to reset
-its input signal to 0,
-and a \textbf{DotInc} operator
-to encode that input signal into
-input currents injected into
-the underlying population of neurons;
-i.e., the \textbf{DotInc} computes
-$J = \alpha_i \mathbf{e_i} X$ from equation \eqref{TODO}.
-
-\subsection{Reference simulator}
-
-The reference simulator starts by
-building NumPy arrays
-for every signal created during the build process
-described above.
-It then builds a graph of the
-operators created during the build process,
-ordering them such that they can be
-executed one after another to get the correct answer.
-
-The ordering is determined
-by which signals an operator
-sets, increments, reads, and updates.
-A signal can only be set or updated once,
-but can be read or incremented any number of times.
-An update represents the operator that sets
-the signal's value at time $t+dt$,
-and therefore must be the last operator
-that acts on a given signal.
-This means that an operator
-that updates a signal depends on
-all operators that set, increment,
-or read that signal.
-Reads represent reading the value
-of a signal at time $t$,
-which is determined by the sets
-and increments, and so
-an operator that reads a given signal
-depends on all operators that
-set and increment that signal.
-Increments must happen
-after a given signal is set,
-so an operator that increments a given
-signal depends on the operator
-that sets that signal, if it exists.
-
-The reference simulator represents
-these dependencies by a directed graph,
-with edges going from
-dependent operators to
-the operators on which they depend.
-The dependency graph is then topologically sorted,
-and that ordering is stored.
-
-The simulator can then be run
-for an arbitrary amount of time.
-On each timestep,
-the simulator goes through each of the
-operators in order,
-and then makes a copy of any signals
-that were probed on this timestep.
-
-\subsection{OpenCL simulator}
-
-OpenCL is a framework for writing
-efficient parallel code that runs
-on many different platforms,
-including modern CPUs and GPUs.
-It allows for the kind of parallel execution
-that is required to run neural models quickly,
-while still being able to run
-on multiple platforms.
-
-The OpenCL simulator operates on
-the same objects that the reference simulator does;
-i.e., it starts with a collection
-of signals and operators,
-schedules them efficiently,
-and runs all operators on each timestep.
-
-However, in order to execute
-many operators in parallel,
-the OpenCL simulator
-organizes computations
-differently than the reference simulator.
-All of the linear operators are mapped
-onto \textbf{MultiProdUpdate} operators,
-which perform the general update
-$$Y \leftarrow \gamma + \beta Y + AX.$$
-It can then schedule many
-\textbf{MultiProdUpdate} operations
-on a single timestep,
-if the dependency graph allows it.
-
-The OpenCL simulator also transparently
-attempts to build OpenCL equivalents
-for arbitrary Python functions
-executed by nodes.
-These arbitrary functions
-can slow down a simulation significantly,
-as data and control has to be passed back
-to the Python interpreter
-on each timestep.
-The OpenCL simulator inspects
-the abstract syntax tree
-of the function being computed
-and substitutes OpenCL equivalents when possible.
-
-\section{Extensibility}
-
-Nengo is designed to be extended
-in order to make modelling easier,
-and to enable models to be run
-on many different platforms.
-
-\subsection{Models}
-
-At model creation time,
-everything must be done
-in a simulator agnostic manner.
-Since ensembles and nodes
-are the only objects that exist,
-models can become long scripts
-that contain many
-similar function calls that
-create and connect ensembles and nodes.
-In order to make modelling scripts
-shorter and more expressive,
-we have created the \texttt{Network} abstraction.
-
-\paragraph{Networks}
-A network is a collection of ensembles and nodes
-connected together in a particular way.
-It is primarily a way of grouping together
-a common set of objects connected in a particular way,
-and provides some implicit expressiveness
-in the name of the network.
-
-Networks can vary dramatically in size,
-and can be composed of other networks.
-The \texttt{Integrator} network, for example,
-is composed of only one recurrently connected ensemble.
-By encapsulating that logic in a network,
-the purpose of that ensemble is made explicit.
-On the other end of the size spectrum,
-the \texttt{BasalGanglia} network,
-is composed of five groups of ensembles
-connected with several specific functions
-that together implement a ``winner take all'' circuit.
-Encapsulating that logic
-makes this complicated section of code
-easy to include in many different models
-(see Figure~\ref{TODO}).
-
-\subsection{Nonlinearities}
-
-Many modellers make use of
-different neuron models and
-learning rules than are
-currently implemented
-in the two simulators
-in section \ref{sec:simulators}.
-Because Nengo endeavors
-to compute nonlinear functions efficiently,
-these nonlinearities
-must be implemented in each simulator,
-though they may share the same representation
-at the model definition stage.
-While this results in simulators
-that cannot run all Nengo models,
-this cannot be avoided,
-and therefore we make this explicit.
-
-In order to implement
-a new neuron model or learning rule,
-two classes must be created.
-The first is a symbolic class
-that keeps track of any parameters
-that might affect the execution
-of the nonlinearity;
-for example, for
-the leaky integrate-and-fire neuron model,
-the refractory time constant,
-$\tau^{ref}$, is part of this symbolic class.
-An instance of this class
-is then used as an argument
-to the ensemble constructor.
-For learning rules,
-an instance of that class
-is used as an argument
-to functions that connect ensembles.
-The second class to be created
-is simulator-specific,
-and contains the function that will be
-executed on each timestep of the simulation.
-Individual simulators may have additional
-requirements when adding
-a new nonlinearity.
-
-\subsection{Model simulation}
-
-The simulator included with Nengo
-is provided primarily as a reference
-for platform experts designing a simulator
-for their platform.
-The reference simulator is implemented with
-a two-step build process;
-this is designed to make
-some simulators easier to implement,
-as the result of
-the first step of the build process
-is a set of computations
-that most computing devices are able to perform.
-
-However, other simulators---especially
-those running on special purpose
-hardware---may not be able
-to perform all of these computations.
-In that case, these simulators
-can define their own build process
-that takes as input a Nengo model.
-This flexibility is possible
-due to the decoupling of
-the model definition and simulation steps.
-
-\section{Examples}
-
-\subsection{Communication channel}
-
-Faithfully transmit a signal.
+
+Many of the computations
+done in a simulation
+are linear (e.g.,
+the decoding and encoding steps
+in Figure~\ref{fig:sim}),
+and therefore can share a common operator;
+this can be helpful for parallelization
+in some simulators.
+
+Nonlinear functions, however,
+require specific operators.
+Each supported neuron model and learning rule
+has an associated operator;
+the simulator explicitly maps
+from symbolic neuron objects in ensembles
+and from symbolic learning rule objects
+in connections to their associated operator.
+
+\section{Examples} \label{sec:examples}
+
+\subsection{Communication channel} \label{sec:comm-channel}
 
@@
 \end{center}
- \textbf{\refstepcounter{figure}\label{fig:cchannel} Figure \arabic{figure}.}{
+ \textbf{\refstepcounter{figure}\label{fig:comm-channel}
+   Figure \arabic{figure}.}{
    Communcation channel model. (Top) Diagram of the model.
@@
 \end{figure}
+
+A communication channel
+is a minimal example demonstrating
+the representation and transformation principles
+of the NEF.
+A communication channel
+represents some signal and transmits it unchanged;
+i.e., it computes the function $f(x) = x$.
+Figure~\ref{fig:comm-channel}
+depicts a scalar communication channel
+representing band-limited Gaussian white noise.
+
+The communication channel
+is a simple enough model
+that it can be readily implemented in PyNN.
+Figure~\ref{fig:pynn} contains the code
+for implementing a communication channel in PyNN.
+This highlights many of the differences
+between Nengo models and conventional neural models;
+we also use the script for benchmarking
+(see section \ref{sec:benchmark}).
 
@@
 
-\subsection{Lorenz attractor network}
-
-Many models in theoretical neuroscience
-are based on attractor networks.
-The NEF has been used in the past
-to implement many different types of
-attractor networks \cite{TODO}.
-In this example,
-we implement the Lorenz chaotic attractor
-with a single ensemble
-composed of 2000 leaky integrate-and-fire neurons.
+\subsection{Lorenz attractor network} \label{sec:lorenz}
 
@@
 
-In order to compare this Nengo model
-with other pieces of neural modelling software,
-we have implemented the Lorenz attractor
-in PyNN \cite{TODO} in order to simulate it
-with the simulators that PyNN supports.
-
-\subsection{Circular convolution}
-
-One of the most common operations done
-in Nengo models is circular convolution.
-Circular convolution is how
-the semantic pointer architecture
-(SPA; \cite{TODO})
-binds two semantic pointers together.
-Binding is central to many
-of the high level cognitive tasks
-that make Spaun unique.
-Circular convolution
-is also best implemented in a two-layer network,
-rather than in a single connection.
-This two-layer network construction is simplified
-by the existence of the \texttt{CircularConvolution} network.
+Many models in theoretical neuroscience
+are based on attractor networks.
+The NEF has been used in the past
+to implement many different types of
+attractor networks \cite{TODO}.
+Figure~\ref{fig:lorez} depicts
+a Nengo implementation of the Lorenz chaotic attractor
+with a single ensemble
+composed of 2000 leaky integrate-and-fire neurons.
+
+We have implemented the Lorenz attractor
+in PyNN for benchmarking purposes
+(see supplementary material).
+
+\subsection{Circular convolution} \label{sec:cconv}
 
@@
 \end{center}
- \textbf{\refstepcounter{figure}\label{fig:lorenz} Figure \arabic{figure}.}{
+ \textbf{\refstepcounter{figure}\label{fig:cconv} Figure \arabic{figure}.}{
    Circular convolution model. (Top) Diagram of the model.
@@
 
-TODO: can we do circconv in PyNN?
+One of the most common operations done
+in Nengo models is circular convolution.
+Circular convolution is how
+the semantic pointer architecture
+(SPA; \cite{TODO})
+binds two semantic pointers together.
+Binding is central to many
+of the high level cognitive tasks
+that make Spaun unique.
+Circular convolution
+is best implemented in a two-layer network,
+rather than in a single connection.
+This two-layer network construction is simplified
+by the existence of the \texttt{CircularConvolution} network.
+The complexity encapsulated in that network
+can be seen in Figure~\ref{fig:cconv}.
+
+Unlike the previous two examples,
+we do not implement
+circular convolution in PyNN.
+The resulting script would be
+too long to be instructive.
 
@@
 
-Even at its \texttt{0.1} release,
-we have met our goals
-for this reimplementation of Nengo.
-\begin{enumerate}
-  \item We have provided a scripting interface
-    similar to the Jython interface.
-    In many cases, the new scripting interface
-    is simpler and more expressive than
-    the previous Jython interface.
-  \item We have leveraged existing scientific computing tools
-    rather than writing our own.
-    In the the scripting interface and the reference simulator,
-    we use NumPy to simplify array computations.
-    In the reference simulator, we use
-    NetworkX to perform the topological sort
-    of the operator dependency graph.
-    In examples, we use Matplotlib
-    to plot the results of simulations.
-  \item We have provided the ability to run models with
-    different software and hardware simulators.
-    We have shown that this is possible by
-    implementing two software simulators.
-\end{enumerate}
-
-While this reimplementation of Nengo
-does not completely replace
-the existing suite of Java-based Nengo tools,
-it is able to cover the vast majority of use cases
-with an order of magnitude less
-code than the Java-based tools
-($\sim$30,000 lines of Java + 10,000 lines of Jython
-vs. $\sim$8,000 lines of Python).
-We are optimistic that the remaining use cases
-can be implemented in this new framework.
-
-TODO: something about documentation, unit tests,
-coverage, travis-ci, etc
-
 \subsection{Comparison to similar projects}
@@
 however, Nengo is unique in being built
-on a theoretical framework (the NEF, \cite{TODO})
+on a theoretical framework
 that has already been used to make
@@
 interface similar to the
-model creation part of Nengo.
+high-level objects in PyNengo.
 Rather than implementing its
 own simulator, PyNN uses existing
-simulators, such as Brian \cite{TODO},
-NEST \cite{TODO}, and NEURON \cite{TODO}.
+simulators with Python bindings,
+such as NEST \cite{TODO}.
 
@@
 Nengo is able to simulate many networks
-faster than all of PyNN's simulators
-(TODO: right?).
-This is, in part, because Nengo stores two vectors
-whose product is the entire
-connection weight matrix between
-two ensembles, rather than storing
-the entire matrix.
+faster than all of PyNN's simulators.
+This is, in part,
+because Nengo stores the factors
+of the connection weight matrix,
+rather than storing the entire matrix.
 
@@
 how those neurons connect to other populations.
-PyNN simulates many different learning rules
+PyNN simulates many learning rules
 and neuron types,
 and is therefore better suited to
-simulate small networks of detailed neuron models
-(as would be necessary for modelling
-an invertebrate nervous system).
+simulate networks of detailed neuron models.
 
@@
 
+\subsection{Conclusion}
+
+PyNengo is the next generation of Nengo.
+Though the project is still young,
+it can already simulate most models
+that have been built using the NEF.
+It does this with 11\% as many lines of code
+as its predecessor,
+and interacts seamlessly with
+other scientific Python tools.
+
+While the reference simulator
+is simple and easy to understand,
+the OpenCL simulator is extremely fast;
+it can simulate a large circular convolution model
+TODO times faster than JavaNengo's simulator.
+This makes the creation and simulation
+of models many times larger than Spaun
+tractable with current hardware;
+these models will further test
+the NEF as a theory of the brain,
+and with PyNengo make those theories
+testable by anyone with a computer.
+
 \section*{Disclosure/Conflict-of-Interest Statement}
@@
 potential conflict of interest.
+
+\section*{Author contributions}
+
+TODO
 

### 2013-09-27 scripts/view_records.py
new: circconv script and data

@@ -1,0 +1,56 @@
+#!/usr/bin/env python
+
+#
+# Usage: ./view_records.py circconv_records/*
+#
+import sys
+import cPickle
+from collections import defaultdict
+
+by_name = defaultdict(list)
+
+for recfile in sys.argv[1:]:
+    records = cPickle.load(open(recfile))
+    for rec in records:
+        rec['filename'] = recfile
+        by_name[rec['name']].append(rec)
+
+import matplotlib.pyplot as plt
+nr = by_name.items()
+nr.sort(key=lambda (name, recs):
+        [-rec['runtime'] for rec in recs if rec['dim'] == 50][0])
+
+nr[1:1] = [('JavaNengo', None)]
+
+for name, recs in nr:
+    if name == 'JavaNengo':
+        plt.plot([100, 200, 500], [9, 18, 45], '.-', markersize=30, label='JavaNengo')
+        continue
+
+    print name.strip()
+    if name.strip() == 'Tahiti':
+        name = 'ATI Radeon HD 7970'
+    if name.strip() == 'ref':
+        name = 'NumPy Reference'
+    if name.strip().startswith("Tesla"):
+        name = 'NVidia ' + name.strip()
+    if '5540' in name.strip():
+        name = 'Intel Xeon E5540 @ 2.53GHz'
+    if '2620' in name:
+        name = 'Intel Xeon E5-2620 @ 2.00GHz'
+    if 'Core' in name.strip():
+        name = 'Intel Core i7-3770 @ 3.40GHz'
+
+    oks = [rec for rec in recs if rec['status'] == 'ok']
+    dims = [rec['dim'] for rec in oks]
+    runtimes = [rec['runtime'] for rec in oks]
+    filenames = [rec['filename'] for rec in oks]
+    for dim, rt, fname in zip(dims, runtimes, filenames):
+        print '  ', dim, rt, fname
+    plt.plot(dims, runtimes, '.-', markersize=30, label=name.strip())
+
+plt.xlabel('n. dimensions convolved')
+plt.ylabel('simulation time (seconds)')
+plt.ylim(0, 20)
+plt.legend(loc=2)
+plt.show()

### 2013-09-27 scripts/cconv.py
Fixed up figures, esp. NEF

@@ -70,6 +70,5 @@ plot(actual_out, "Result")
 plt.xlabel("Time (s)")
 plt.ylabel("Result")
 
-plt.tight_layout()
-plt.subplots_adjust(hspace=0.12)
+plt.tight_layout(h_pad=0.12)
 plt.savefig("../figures/cconv_res.pdf")

### 2013-09-27 scripts/comm_channel.py

@@ -75,5 +75,5 @@ ax.text(0.5, 1.0, "B", ha='center', va='center', fontsize=20,
         bbox=dict(ec='none', fc='w', alpha=0.8))
 adjust(ax)
 
-plt.tight_layout()
+plt.tight_layout(h_pad=0.15)
 plt.savefig("../figures/comm_channel_res.pdf")

### 2013-09-27 scripts/lorenz.py

@@ -65,6 +65,6 @@ ax.xaxis.set_ticks_position('bottom')
 ax.yaxis.tick_right()
 ax.yaxis.set_label_position("right")
 
-plt.tight_layout()
-plt.subplots_adjust(left=0, hspace=0.12)
+plt.tight_layout(h_pad=0.12)
+plt.subplots_adjust(left=0)
 plt.savefig("../figures/lorenz_res.pdf")

### 2013-09-27 scripts/nef_summary.py

@@ -27,6 +27,7 @@ cos = model.connect("A", "Dummy", function=np.cos, filter=0.03)
 model.probe(cos)
 
 ## Encoding plot
+figsize = (3.5, 7)
 
 sim = model.simulator()
 sim.run(1)
@@ -39,24 +40,40 @@ A.eval_points.sort(axis=0)
 J = np.dot(A.eval_points, A.encoders.T)
 activities = A.neurons.rates(J)
 
-plt.figure(figsize=(4,9))
+plt.figure(figsize=figsize)
+
 plt.subplot(3,1,1)
-plt.plot(sim.data(model.t)[1:], sim.data("Input signal")[1:], lw=2)
-plt.title("Input signal")
-plt.xlim(0, 1)
+plt.title("Encoding")
+plt.plot(A.eval_points, activities, lw=2)
+plt.ylabel("Firing rate (Hz)")
+plt.gca().spines['top'].set_visible(False)
+plt.gca().spines['right'].set_visible(False)
+plt.gca().xaxis.set_ticks_position('bottom')
+plt.gca().yaxis.set_ticks_position('left')
 
 plt.subplot(3,1,2)
-plt.plot(A.eval_points, activities, lw=2)
-plt.xlabel("Input signal")
-plt.ylabel("Firing rate (Hz)")
+plt.plot(sim.data(model.t)[1:], sim.data("Input signal")[1:], lw=2)
+plt.gca().spines['top'].set_visible(False)
+plt.gca().spines['bottom'].set_visible(False)
+plt.gca().spines['right'].set_visible(False)
+plt.gca().yaxis.set_ticks_position('left')
+plt.xlim(0, 1)
+plt.xticks(())
+plt.ylabel("Input signal")
 
 plt.subplot(3,1,3)
 color_cycle = plt.gca()._get_lines.color_cycle
 colors = [next(color_cycle) for _ in xrange(model.get("A").n_neurons)]
 plt.axis([0, 1, model.get("A").n_neurons-0.5, -0.5])
 plt.eventplot(spikes, colors=colors)
+plt.gca().spines['top'].set_visible(False)
+plt.gca().spines['right'].set_visible(False)
+plt.gca().xaxis.set_ticks_position('none')
+plt.gca().yaxis.set_ticks_position('none')
+plt.xlabel("Time (s)")
+plt.ylabel("Neuron")
 
-plt.tight_layout()
+plt.tight_layout(pad=0.1, h_pad=0.1)
[…]

### 2013-10-02 nengo.tex
Fixed up figures, esp. NEF
Figures done

@@
 \def\Authors{Trevor Bekolay\,$^{1,*}$, James Bergstra\,$^{1}$,
-  Xuan Choo\,$^{1}$, Travis DeWolf\,$^{1}$, Eric Hunsberger\,$^{1}$,
-  Daniel Rasmussen\,$^{1}$, Terrence C. Stewart\,$^{1}$
+  Eric Hunsberger\,$^{1}$, Travis DeWolf\,$^{1}$, Terrence C. Stewart\,$^{1}$,
+  Daniel Rasmussen\,$^{1}$, Xuan Choo\,$^{1}$, Aaron Voelker\,$^{1}$,
   and Chris Eliasmith\,$^{1}$}
@@
 \begin{abstract}
-%Maximum: 2000 characters
-
-  \tiny \keyFont{\section{Keywords:} 1 2 3 4 5 (6 7 8)}
+  Neuroscience currently lacks
+  a comprehensive theory of
+  how the brain implements cognitive processes
+  with biological parts.
+  The Neural Engineering Framework (NEF)
+  proposes one such theory,
+  but has not has not gathered
+  significant empirical support
+  due, in part, to the technical
+  challenge of building and simulating
+  large-scale models with the NEF.
+  Nengo is a tool that builds and simulates
+  functional brain models with the NEF.
+  Nengo is the primary resource
+  for teaching how the NEF
+  is used to model neural systems,
+  and for researchers using the NEF
+  to build and simulate
+  functional brain models.
+  A previous version of Nengo was used
+  to create Spaun,
+  the world's largest functional brain model.
+  The implementation of Spaun
+  identified limitations
+  in Nengo's ability to
+  construct models with simple syntax,
+  simulate them quickly,
+  and collect data for subsequent analysis.
+  We have implemented a new version of Nengo
+  in Python that overcomes these limitations.
+  It uses simple and extendable syntax,
+  simulates a benchmark model on the scale of Spaun
+  200 times faster than the previous implementation,
+  and has a flexible mechanism
+  for collecting simulation results.
+
+  \tiny \keyFont{\section{Keywords:} Python, Neural Engineering
+    Framework, theoretical neuroscience, computational neuroscience,
+    control theory, software, simulation}
 \end{abstract}
@@
 \begin{table}[!t]
-\processtable{Nengo tools\label{tab:nengo-vers}}
+\processtable{Tools bearing the name Nengo.\label{tab:nengo-vers}}
 {\begin{tabular}{p{2.6cm} p{2.7cm} l p{7cm}} \toprule
@@
   & Simulates models constructed with PyNengo. \\\botrule
-\end{tabular}}{Code repositories avilable at \url{http://github.com/ctn-waterloo/}}
+\end{tabular}}{Code repositories avilable at \url{http://github.com/ctn-waterloo/}.}
 \end{table}
@@
 \begin{center}
-  \includegraphics[width=0.19\textwidth]{nef_summary_enc}
-  \includegraphics[width=0.19\textwidth]{nef_summary_dec}
-  \includegraphics[width=0.399\textwidth]{nef_summary_trans}
-  \includegraphics[width=0.199\textwidth]{nef_summary_dyn}
+  \includegraphics[width=0.245\textwidth]{nef_summary_enc}
+  \includegraphics[width=0.245\textwidth]{nef_summary_dec}
+  \includegraphics[width=0.245\textwidth]{nef_summary_trans}
+  \includegraphics[width=0.245\textwidth]{nef_summary_dyn}
 \end{center}
  \textbf{\refstepcounter{figure}\label{fig:nef} Figure \arabic{figure}.}{
-   Summary of the Neural Engineering Framework.}
+   Summary of the three principles of the Neural Engineering Framework
+   (NEF). (A) By the representation principle, signals are encoded
+   in neural populations based on the \textit{tuning curve}
+   of each neuron (top). The tuning curve describes
+   how active a neuron is given some input signal.
+   If we drive the eight neurons in the top panel
+   with the input signal in the middle panel,
+   we see the firing pattern in the bottom panel.
+   (B) By the representation principle,
+   the spiking activity of a neural population
+   can be decoded to recover the original input signal,
+   or some transformation of that input signal.
+   First, the firing pattern shown in the top panel
+   is filtered with a decaying exponential filter (middle panel).
+   The filtered activity is then summed together
+   with a set of weights that approximates
+   the input signal (bottom panel, green)
+   and the cosine of the input signal (bottom panel, purple).
+   (C) By the transformation principle,
+   populations of neurons can send signals
+   to another population by decoding
+   the desired function from the first population
+   and then encoding the decoded estimate
+   into the second population.
+   These two steps can be expanded into a single step
+   by calculating a set of weights
+   that describe the strength of the connection
+   between each neuron in the first population
+   and each neuron in the second population.
+   In (C), a sine wave is encoded by population A (top panel);
+   the negative of that signal is projected
+   to population B (middle panel)
+   and the square of that signal is projected
+   to population C (bottom panel).
+   (D) By the dynamcs principle,
+   signals being represented by population of neurons
+   can be thought of
+   as state variables in a dynamical system.
+   In this example, the dynamical system
+   has negative feedback across its two dimensions,
+   resulting in a harmonic oscillator.
+   The oscillator is plotted across time (top)
+   and in state space (bottom).}
 \end{figure}
@@
 \begin{center}
-  \includegraphics[width=0.6\textwidth]{sim}
+  \includegraphics[width=\textwidth]{sim}
 \end{center}
  \textbf{\refstepcounter{figure}\label{fig:sim} Figure \arabic{figure}.}{
-   The first two timesteps of a simple simulation.}
+   Detailed breakdown of the PyNengo reference simulator
+   running a simple model for two timesteps.
+   (A) Code describing the model being simulated.
+   It consists of ensemble A project the sine of its
+   encoded signal to ensemble B,
+   which is recurrently connected.
+   (B) Diagram depicting the model being simulated.
+   (C) A detailed diagram of how the reference simulator
+   simulates this model. Signals, which are the values
+   tracked in the simulation, are colored blue.
+   Operators, which are the computations done on signals,
+   are colored red.
+   Signals can be grouped as low-level neural signals
+   that are used to compute the nonlinear functions
+   underlying neuron models,
+   and high-level NEF signals that are used to
+   drive neurons and track the signals
+   that the neurons are representing.
+   The operators that implement the decoding
+   and encoding steps map between
+   the low-level neural signals
+   and the high-level NEF signals.
+   (D) The signals tracked at the low level
+   can be interpreted as a model
+   commonly seen in computational neuroscience literature;
+   a population of leaky integrate-and-fire neurons
+   is driven by some time-varying input current, $J(t)$.
+   These neurons project to a population
+   of recurrently connected neurons.
+   The connection weight between the two populations,
+   and from the second population to itself,
+   can be computed by the NEF's transformation
+   principle, bypassing the need for
+   the high-level NEF signals
+   used by the reference simulator
+   for speed and data collection purposes.
+   (E) The signals tracked at the high level
+   can be interpreted as a dynamical system
+   that can be analyzed with control theory.
+   A simply represents its input,
+   and passes its state to a sin function
+   which becomes the input to B.
+   B is a simple linear system
+   that can be described with the typical
+   $\dot{x}(t) = A x(t) + B u(t)$ equation.
+   These dynamical systems can be simulated
+   directly, without the use of spiking neurons,
+   in order to quickly analyze system behavior.}
 \end{figure}
@@
 \begin{center}
-  \includegraphics[width=.9\textwidth]{comm_channel}
-  \begin{minipage}{0.58\textwidth}
+  \includegraphics[width=.85\textwidth]{comm_channel}
+  \begin{minipage}{0.48\textwidth}
     \begin{lstlisting}
@@
   \end{minipage}
-  \begin{minipage}{0.37\textwidth}
+  \begin{minipage}{0.39\textwidth}
     \includegraphics[width=\textwidth]{comm_channel_res}
@@
    Figure \arabic{figure}.}{
-   Communcation channel model. (Top) Diagram of the model.
-   (Bottom left) Python code to create and run the model.
-   (Bottom right) Results of the model simulation. Lines depict
-   the node's output (Input panel) and the decoded value being represented
-   by ensembles (A and B panels). Light gray lines in A and B panels denote
-   spikes emitted by neurons in the ensembles. Each row depicts spikes
-   for one neuron.}
+   A communication channel implemented with Nengo.
+   (A) Diagram depicting the model. Ensemble A
+   projects its encoded signal to ensemble B unchanged.
+   (B) Nengo code to build and simulate the model
+   for 1 second.
+   (C) The results of the simulation.
+   The input signal (top panel) is white noise limited to 0 to 5 Hz.
+   The signal is well represented by both ensemble A (middle panel)
+   and ensemble B (bottom panel) despite the neural firing patterns
+   (underlaid in middle and bottom panels) being different.}
 \end{figure}
@@
 \begin{center}
-  \begin{minipage}{.495\textwidth}
+  \begin{minipage}{.435\textwidth}
     \begin{lstlisting}[basicstyle={\footnotesize\ttfamily}]
@@
   \end{minipage}
-  \begin{minipage}{.495\textwidth}
+  \begin{minipage}{.46\textwidth}
     \begin{lstlisting}[basicstyle={\footnotesize\ttfamily}]
@@
  \textbf{\refstepcounter{figure}\label{fig:pynn} Figure \arabic{figure}.}{
-   An implementation of the communication channel in PyNN.}
+   Implementation of the communication channel in PyNN.
+   Note that several quantities are generated from Nengo;
+   the functions that generate these quantities in Nengo
+   can be decoupled, but are not included here for brevity.
+   The implementation of the Lorenz attractor
+   is not shown, but is very similar to the communication channel.}
 \end{figure}
@@
  \textbf{\refstepcounter{figure}\label{fig:lorenz} Figure \arabic{figure}.}{
-   Lorenz attractor model. (Left) Python code to create and run the model.
-   (Top right) Diagram of the model.
-   (Bottom right) Results of the model simulation. Lines depict
-   the node's output (Input panel) and the decoded value being represented
-   by ensembles (A and B panels). Light gray lines in A and B panels denote
-   spikes emitted by neurons in the ensembles. Each row depicts spikes
-   for one neuron.}
+   A Lorenz attractor implemented with Nengo.
+   (A) Nengo code to build and simulate the model
+   for 6 seconds.
+   (B) Diagram depicting the model. The state ensemble
+   is recurrently connected with complex dynamics
+   implemented the chaotic Lorenz attractor.
+   Not that this populations does not receive
+   any input that might drive its initial value;
+   instead, the initial value is determined by
+   the baseline firing of the 2000 leaky integrate-and-fire
+   neurons that make up the state ensemble.
+   (C) The trajectory that the state ensemble takes
+   in its three-dimensional state space.
+   For the parameters chosen, the trajectory takes
+   the well-known butterfly shape.
+   (D) The state vector plotted over time.
+   (E) The spikes emitted by a random sample of 25
+   neurons from the state ensemble.
+   Some neurons fire uniformly across the 6 second simulation,
+   but most change depending on the state being tracked
+   due to the recurrent connection.}
 \end{figure}
@@
   \includegraphics[width=0.8\textwidth]{cconv}
-  \begin{minipage}{0.53\textwidth}
+  \begin{minipage}{0.46\textwidth}
     \begin{lstlisting}[basicstyle={\footnotesize\ttfamily}]
@@
   \end{minipage}
-  \begin{minipage}{0.43\textwidth}
+  \begin{minipage}{0.46\textwidth}
     \includegraphics[width=\textwidth]{cconv_res}
@@
 \end{center}
- \textbf{\refstepcounter{figure}\label{fig:cconv} Figure \arabic{figure}.}{
-   Circular convolution model. (Top) Diagram of the model.
-   Note that much of the complexity is contained within the
-   circular convolution network; the transform matrices shown
+ \textbf{\refstepcounter{figure}\label{fig:cconv} Figure
+   \arabic{figure}.}{
+   Circular convolution implemented with Nengo.
+   (A) Diagram depicting the model.
+   The input vectors, A and B, represent four-dimensional signal
+   which are mapped onto six ensembles within the
+   Circular Convolution network through
+   a complicated transformation matrix.
+   Each ensemble within the network represents a
+   two-dimensional signal. The product of those two signals
+   is projected through another complicated transformation matrix
+   to compute the final four-dimensional result.
+   Note that the complicated parts of the model
+   are contained within the network;
+   the number of ensembles and the transform matrices shown
    are automatically generated by the network depending on
-   input dimensionality.
-   (Bottom left) Python code to create and run the model.
-
-   (Bottom right) Results of the model simulation. Lines depict
-   the node's output (Input panel) and the decoded value being represented
-   by ensembles (A and B panels). Light gray lines in A and B panels denote
-   spikes emitted by neurons in the ensembles. Each row depicts spikes
-   for one neuron.}
+   the dimensionality of the input vectors.
+   (B) Nengo code to build and simulate the model
+   for 0.5 seconds.
+   (C) The result of the simulation.
+   Straight horizontal lines represent
+   the target values that each ensemble
+   should represent.
+   Wavy lines represent the decoded values
+   for each dimension represented by the
+   A, B, and Result ensembles
+   (top, middle, bottom panels respectively).
+   The ensembles represent the correct values,
+   after a startup transient of less than 0.1 seconds.}
 \end{figure}
@@
 \begin{center}
-  \includegraphics[width=0.5\textwidth]{ocl_vs_ref}
+  \includegraphics[width=0.32\textwidth]{bench_cchannel}
+  \includegraphics[width=0.32\textwidth]{bench_lorenz}
+  \includegraphics[width=0.32\textwidth]{bench_cconv}
 \end{center}
  \textbf{
-     \refstepcounter{figure}\label{fig:ocl_vs_ref}
+     \refstepcounter{figure}\label{fig:benchmarks}
      Figure \arabic{figure}.}{
-     Five-hundred-dimensional circular convolution networks were some of the most
-     computationally demanding elements of the Spaun cognitive architecture.
-     Simulations of circular convolution networks of various sizes illustrate the speed of our re-engineered simulators.
-     A straightforward reference implementation in Python is sufficient for small networks,
-     while for larger ones the new OpenCL-powered simulator runs up to 200$\times$ faster,
-     In particular the Radeon 7970 GPU can simulate 500-dimensional convolution networks faster than real time (0.95 seconds for a 1-second simulation).
-     Also, although both are multithreaded, our new implementation delivers better CPU performance than the previous JavaNengo implementation.
+     For (A) and (B), all of the simulators
+     except the PyNengo OpenCL simulator
+     was run on an Intel Core i7-965.
+     JavaNengo used all 4 cores of this processor;
+     all other simulator used only 1 core.
+     The PyNengo OpenCL simulator
+     was run on an NVidia GTX280 GPU.
+     For (C), if not noted, the simulator used
+     was the PyNengo OpenCL simulator.
+     The CPU used for JavaNengo and
+     the PyNengo reference simulator
+     was an Intel Core i7-3770;
+     all 4 cores were used by JavaNengo,
+     while PyNengo's reference simulator
+     only used one core.
      }
 \end{figure}
+
+While benchmark models are not indicative
+of performance on all models,
+increasing simulation speed
+was a primary motivators in creating PyNengo.
+To validate that performance has improved,
+we ran the models described in section~\ref{sec:examples}
+for various numbers of neurons and dimensions
+for each ensemble.
+
+The communication channel and Lorenz attractor
+are small models that demonstrate
+the principles of the NEF.
+Their small size enables us to write
+PyNN scripts that implement roughly
+the same functionality
+with Brian, NEURON, NEST, and TODO PCSim.
+We run each parameter set five times
+on the same machine,
+and plot the mean.
+In most cases, the coefficient of variation
+is well below 0.1, except for two
+outliers with coefficients of 0.18 and 0.22,
+indicating that these means are robust.
+The results, shown in Figure~\ref{fig:benchmarks},
+suggest that all versions of Nengo are significantly
+faster than the simulators accessible
+through PyNN, especially
+as the size of the model increases.
+This is likely due to Nengo's
+use of factorized weight matrices,
+rather than storing and computing with
+the entire weight matrix
+on each timestep.
+While NEST and NEURON were not
+run with MPI,
+the reference simulator of PyNengo
+also only uses one CPU core.
+The results also suggest that PyNengo's
+OpenCL simulator is faster
+than the reference simulator
+and JavaNengo's simulator,
+though the difference
+is not overwhelming in these models.
+
+As a real-world example,
+we have also benchmarked
+the circular convolution model.
+Circular convolution is an important test case,
+as a significant portion of Spaun's
+2.5 million neurons are used to
+implement circular convolution.
+In this case, only versions of Nengo
+were tested, and were tested rigorously.
+Instead of running each simulation multiple times,
+we instead run the simulator for 10 timesteps
+in order to fill various levels of the cache,
+and then run the simulator for 1000 more timesteps;
+there is very little variance using this method.
+As can be seen in Figure~\ref{fig:benchmarks},
+for large models, the OpenCL simulator
+performs much faster than JavaNengo;
+in particular, a Radeon 7970 GPU performs 500-dimensional
+circular convolution faster than real time,
+and 200 times faster than JavaNengo.
+Additionally, although both JavaNengo
+and the OpenCL simulator on a CPU
+both use all available CPU cores,
+PyNengo's OpenCL simulator is significantly faster.
 

### 2013-10-02 scripts/bench_cchannel.py

@@ -1,0 +1,144 @@
+"""
+Simulating communication channel for 10 seconds with 1ms timestep
+"""
+
+from collections import OrderedDict
+
+import numpy as np
+import matplotlib.pyplot as plt
+from mpl_toolkits.axes_grid1.inset_locator import zoomed_inset_axes
+from mpl_toolkits.axes_grid1.inset_locator import mark_inset
+
+data = OrderedDict()
+data['Brian'] = OrderedDict()
+data['NEURON'] = OrderedDict()
+data['NEST'] = OrderedDict()
+data['JavaNengo'] = OrderedDict()
+data['PyNengo\nOpenCL'] = OrderedDict()
+data['PyNengo\nreference'] = OrderedDict()
+colors = ['#238B45', '#984EA3', '#FF7F00', 'k', '#08519C', '#E41A1C']
+
+data['PyNengo\nOpenCL'][(100, 1)] = [
+    0.794183015823,0.797194957733,0.798620939255,0.796317100525,0.799592971802]
+data['PyNengo\nOpenCL'][(250, 1)] = [
+    0.836793899536,0.837684869766,0.840667963028,0.837545871735,0.837010145187]
+data['PyNengo\nOpenCL'][(500, 10)] = [
+    1.11762309074,1.11406683922,1.11392784119,1.11361694336,1.11517691612]
+data['PyNengo\nOpenCL'][(1000, 50)] = [
+    4.37966895103,4.36161804199,4.36030006409,4.36526703835,4.36328911781]
+
+data['PyNengo\nreference'][(100, 1)] = [
+    1.79912400246,1.72637414932,1.77085995674,1.72308516502,1.75330209732]
+data['PyNengo\nreference'][(250, 1)] = [
+    1.97227215767,1.98903083801,1.96172499657,2.02151703835,2.0146241188]
+data['PyNengo\nreference'][(500, 10)] = [
+    2.4279820919,2.41257190704,2.44041395187,2.39347505569,2.40128898621]
+data['PyNengo\nreference'][(1000, 50)] = [
+    3.12998580933,3.08235788345,3.18121695518,3.12098288536,3.13230895996]
+
+data['JavaNengo'][(100, 1)] = [
+    4.13199996948,4.29999995232,3.9509999752,3.97099995613,4.41100001335]
+data['JavaNengo'][(250, 1)] = [
+    5.25999999046,5.44099998474,4.79100012779,4.75,4.85299992561]
+data['JavaNengo'][(500, 10)] = [
[…]

### 2013-10-02 scripts/bench_cconv.py

@@ -1,0 +1,113 @@
+from collections import OrderedDict
+
+import matplotlib.pyplot as plt
+from mpl_toolkits.axes_grid1.inset_locator import zoomed_inset_axes
+from mpl_toolkits.axes_grid1.inset_locator import mark_inset
+
+neurons = {
+    5:4224, 10:8448, 20:17408, 50:44288, 100:89088, 200:178688, 500:447488
+}
+
+data = OrderedDict()
+
+data['JavaNengo'] = OrderedDict()
+data['PyNengo\nreference'] = OrderedDict()
+data['Intel Core\ni7-3770'] = OrderedDict()
+data['Intel Xeon\nE5540'] = OrderedDict()
+data['Intel Xeon\nE5-2620'] = OrderedDict()
+data['NVidia\nQuadro\nK4000'] = OrderedDict()
+data['NVidia Tesla\nC2050'] = OrderedDict()
+data['ATI Radeon\nHD7970'] = OrderedDict()
+colors = ['k', '#E41A1C', '#023858', '#045A8D', '#0570B0',
+          '#3690C0', '#74A9CF', '#A6BDDB']
+
+data['JavaNengo'][100] = 9
+data['JavaNengo'][200] = 18
+data['JavaNengo'][500] = 45
+data['PyNengo\nreference'][5] = 2.62999796867
+data['PyNengo\nreference'][10] = 5.21562504768
+data['PyNengo\nreference'][20] = 10.5295498371
+data['PyNengo\nreference'][50] = 27.0069601536
+data['Intel Core\ni7-3770'][5] = 0.168757915497
+data['Intel Core\ni7-3770'][10] = 0.256566047668
+data['Intel Core\ni7-3770'][20] = 0.482143878937
+data['Intel Core\ni7-3770'][50] = 1.10272312164
+data['Intel Core\ni7-3770'][100] = 2.19662594795
+data['Intel Core\ni7-3770'][200] = 5.19148278236
+data['Intel Core\ni7-3770'][500] = 17.1793811321
+data['ATI Radeon\nHD7970'][5] = 0.0806539058685
+data['ATI Radeon\nHD7970'][10] = 0.0720150470734
+data['ATI Radeon\nHD7970'][20] = 0.0996639728546
+data['ATI Radeon\nHD7970'][50] = 0.127480983734
+data['ATI Radeon\nHD7970'][100] = 0.172482967377
+data['ATI Radeon\nHD7970'][200] = 0.290589809418
[…]

### 2013-10-02 scripts/bench_lorenz.py

@@ -1,0 +1,138 @@
+"""
+Simulating Lorenz attractor for 10 seconds with 1ms timestep
+"""
+
+from collections import OrderedDict
+
+import numpy as np
+import matplotlib.pyplot as plt
+from mpl_toolkits.axes_grid1.inset_locator import zoomed_inset_axes
+from mpl_toolkits.axes_grid1.inset_locator import mark_inset
+
+data = OrderedDict()
+data['Brian'] = OrderedDict()
+data['NEST'] = OrderedDict()
+# data['NEURON'] = OrderedDict()
+data['JavaNengo'] = OrderedDict()
+data['PyNengo\nreference'] = OrderedDict()
+data['PyNengo\nOpenCL'] = OrderedDict()
+colors = ['#238B45', '#FF7F00', 'k', '#E41A1C', '#08519C'] # '#984EA3',
+
+data['PyNengo\nOpenCL'][100] = [
+    0.823611974716,0.827681064606,0.824057102203,0.826804876328,0.825788021088]
+data['PyNengo\nOpenCL'][500] = [
+    1.03150510788,1.03527617455,1.03847002983,1.04004907608,1.08543992043]
+data['PyNengo\nOpenCL'][1000] = [
+    1.27273511887,1.27891302109,1.26637387276,1.27474498749,1.26921701431]
+data['PyNengo\nOpenCL'][2000] = [
+    1.73656892776,1.73761916161,1.72962117195,1.72728586197,1.73043680191]
+
+data['PyNengo\nreference'][100] = [
+    1.09255409241,1.09260296822,1.09535098076,1.04782700539,1.09734082222]
+data['PyNengo\nreference'][500] = [
+    1.37857794762,1.36459708214,1.35432505608,1.36031198502,1.35320615768]
+data['PyNengo\nreference'][1000] = [
+    1.62529683113,1.61883807182,1.64285898209,1.61444401741,1.63075304031]
+data['PyNengo\nreference'][2000] = [
+    2.17722296715,2.18516111374,2.20153999329,2.20053505898,2.19265985489]
+
+data['JavaNengo'][100] = [
+    3.18400001526,3.3599998951,3.22500014305,3.28499984741,3.30400013924]
+data['JavaNengo'][500] = [
+    5.78500008583,5.55000019073,5.67199993134,5.05799984932,5.44799995422]
+data['JavaNengo'][1000] = [
+    8.76200008392,8.84599995613,8.36500000954,7.40699982643,8.35899996758]
+data['JavaNengo'][2000] = [
[…]

### 2013-10-02 scripts/view_records.py

@@ -1,56 +1,0 @@
-#!/usr/bin/env python
-
-#
-# Usage: ./view_records.py circconv_records/*
-#
-import sys
-import cPickle
-from collections import defaultdict
-
-by_name = defaultdict(list)
-
-for recfile in sys.argv[1:]:
-    records = cPickle.load(open(recfile))
-    for rec in records:
-        rec['filename'] = recfile
-        by_name[rec['name']].append(rec)
-
-import matplotlib.pyplot as plt
-nr = by_name.items()
-nr.sort(key=lambda (name, recs):
-        [-rec['runtime'] for rec in recs if rec['dim'] == 50][0])
-
-nr[1:1] = [('JavaNengo', None)]
-
-for name, recs in nr:
-    if name == 'JavaNengo':
-        plt.plot([100, 200, 500], [9, 18, 45], '.-', markersize=30, label='JavaNengo')
-        continue
-
-    print name.strip()
-    if name.strip() == 'Tahiti':
-        name = 'ATI Radeon HD 7970'
-    if name.strip() == 'ref':
-        name = 'NumPy Reference'
-    if name.strip().startswith("Tesla"):
-        name = 'NVidia ' + name.strip()
-    if '5540' in name.strip():
-        name = 'Intel Xeon E5540 @ 2.53GHz'
-    if '2620' in name:
-        name = 'Intel Xeon E5-2620 @ 2.00GHz'
-    if 'Core' in name.strip():
-        name = 'Intel Core i7-3770 @ 3.40GHz'
-
-    oks = [rec for rec in recs if rec['status'] == 'ok']
-    dims = [rec['dim'] for rec in oks]
-    runtimes = [rec['runtime'] for rec in oks]
-    filenames = [rec['filename'] for rec in oks]
-    for dim, rt, fname in zip(dims, runtimes, filenames):
-        print '  ', dim, rt, fname
-    plt.plot(dims, runtimes, '.-', markersize=30, label=name.strip())
-
-plt.xlabel('n. dimensions convolved')
-plt.ylabel('simulation time (seconds)')
-plt.ylim(0, 20)
-plt.legend(loc=2)
-plt.show()

### 2013-10-02 nengo.tex
Edits from Aaron's feedback

@@
 Networks are designed to be
-subclasses by users,
+subclassed by users,
 in order to group together
@@
 
-A probe is used to denotes a signal
-that will be recorded
+A probe marks that a signal
+will be recorded
 throughout the simulation.
@@
 not as a specification.
-
-\subsection{Reference simulator}
 
@@
 between a pair of neurons in the two populations.
+This difference reflects the
+fundamental difference that Nengo
+is built on a theoretical framework
+that enables modellers to think
+about information processing in the brain
+at a conceptual level.
 
@@
 as a whole.
+Internally, we will bring the functionality
+of PyNengo to parity with JavaNengo
+by implementing the use cases
+not currently covered by PyNengo,
+and create a graphical interface
+to build models and
+interactively inspect simulations.
 

### 2013-10-02 nengo.tex
Edits from Travis's feedback

@@
 and to define complex experimental setups.
+Additionally, Jython is always several versions
+behind CPython, making it impossible
+to use new language features
+like dictionary comprehensions.
+Jython also cannot access libraries
+that contain any C code,
+including all scientific libraries
+that use NumPy.
 
@@
 currently simulated by JavaNengo
-using 11\% as many lines of code,
+using 11\% as many lines of code,\footnote{
+  JavaNengo is 40,000 lines of code, not including the GUI,
+  and PyNengo is 5000 lines of code. TODO}
 and depending on
@@
 in several ways.
-Like other high-level objects,
-a connection contains symbolic information
-about how two objects are to be connected together.
+Like other objects in PyNengo's object model,
+a connection contains symbolic information;
+in this case, a connection tracks
+how two objects are to be connected together.
 
@@
 On the other end of the spectrum,
-the \texttt{BasalGanglia} network,
+the \texttt{BasalGanglia} network
 is composed of five groups of ensembles
 connected with several specific functions
-that together implement a ``winner take all'' circuit.
-Encapsulating that logic
+that together implement a ``winner-take-all'' circuit.
+Encapsulating the code to create those ensembles
+and connections in a network subclass
 makes a complicated section of code

### 2013-10-02 nengo.tex
Edits from Bryan's feedback

@@
 describes a population
-of 50 leaky integrate-and-fire neurons
+of 50 leaky integrate-and-fire neurons \cite{TODO} % LIF cite
 with a refractory time of 0.002 seconds.
-Neurons are defined symbolically
+The \texttt{nengo.LIF} class defines
+the parameters of the LIF neurons symbolically
 so that each simulator can compute
-this population's nonlinear function
-however it sees fit.
-The neuron model used by the population
-is easily changed, as long as the simulator
-is aware of the neuron model description object.
+the LIF nonlinearity efficiently.
+The neuron model used by the ensemble
+is changed by passing in a different symbolic neuron object;
+however, the simulator used must be aware
+of that neuron model.
 
@@
 these additional attributes
-also allow modellers to ensure that
-the ensemble stays
-within neurobiological constraints.
+allow modellers to fit
+neural parameters to \textit{in vivo}
+electrophysiology data.
 If these attributes are not set,
-then they will be randomly selected
-from distributions that describe
+Nengo attempts to maintain
+neurobiological constraints
+by selecting neural parameters
+from distributions consistent with
 neocortical pyramidal cells.
@@
 
-A \texttt{Node} tracks any information
-that is not represented by an ensemble.
-In the most common case,
-nodes provide input signals
-that drive ensembles of sensory neurons.
-In a more general sense,
-nodes represent the experimental environment
-in which a neural model exists.
-
-A node computes an arbitrary function
-of its inputs directly.
-Possible input signals include
+A \texttt{Node} contains a user-defined
+Python function that directly calculates
+the node's outputs from its inputs at each timestep.
+Available inputs include
 the simulator timestep,
@@
 
+A node tracks any information
+that is not represented by an ensemble.
+As a simple example,
+a node can be used to model sensory stimuli
+that are predefined functions of time.
+As a more sophisticated example,
+a node can be used to model
+a complex experimental environment that
+both provides input to the neural model
+and responds to the neural model's output.
 Nodes allow many of the special-purpose
@@
 can translate that connection
-to connection weights.
+to synaptic weights.
 
@@
 
-A probe marks that a signal
+A \texttt{Probe} marks that a signal
 will be recorded
 throughout the simulation.
-Nengo objects and connections
-each have several signals that can be probed.
-An ensemble's decoded value
-and its neurons' spikes
-can be probed, among others.
+Nengo models contain many variables
+that change over time,
+including membrane potentials,
+spike events, and represented variables.
+It is resource intensive to store the values of
+large numbers of variables
+at each time step, and it is also not necessary,
+as typically only a small fraction
+of these variables are analyzed after a simulation.
+The modeller chooses which variables to
+record by connecting a probe to an object.
 
@@
 Many models in theoretical neuroscience
-are based on attractor networks.
+are based on attractor networks \cite{TODO}. % Amit, Rolls, Deco
 The NEF has been used in the past

### 2013-10-02 nengo.tex
Edits from Dan's feedback

@@
 
-A node tracks any information
-that is not represented by an ensemble.
+A node represents information
+that cannot be decoded from an ensemble.
 As a simple example,
@@
 
-One of the most common operations done
-in Nengo models is circular convolution.
+A common function computed by Nengo models
+is circular convolution.
 Circular convolution is how
@@
 e.g., \texttt{DecodedConnection(A, B,
-function=product)} connects ensemble A
-to ensemble B, projecting the product of
-some dimensions encoded by A to B.
+function=square)} connects ensemble A
+to ensemble B, projecting the square of
+the value represented by A to B.
 In PyNN, connections commonly describe

### 2013-10-02 nengo.tex
Edits from Emily's feedback

@@
 that attempt to recreate neuroscientific
-data in precise detail,
-but behavior has not emerged
+data in precise detail
+with the thought that
+cognition will emerge by connecting
+detailed neuron models according
+to the statistics of biological connections \cite{TODO}; % blue brain
+cognition has not yet emerged
 from data-driven large scale models.
 Similarly, theoretical frameworks
-consisting of a generic algorithm
-replicated many times
-(e.g., the memory-prediction framework \cite{TODO}
-and the pattern recognition theory of mind \cite{TODO})
-have not been shown to solve problems
-that the generic algorithm alone cannot solve.
+that use a generic algorithm
+replicated many times % reword based on papers below
+are unable to solve more complex problems
+than the generic algorithm alone
+(e.g., \cite{TODO} \cite{TODO}). % Hawkins, Kurzweil
 
@@
 that can perform eight high-level cognitive tasks
-including list memory and inductive reasoning.
+including memorizing lists and inductive reasoning.
 It can perform any of these eight tasks
@@
 tracing out digits.
-If asked to recall the remember list,
+If asked to recall the remembered list,
 it would trace out the digits $1234$.
@@
 the NEF enables Spaun to employ
-many different algorithms
-based on the behavior being performed,
+different sets of algorithms
+depending on the behavioral task performed,
 rather than defining a single algorithm
@@
 and a scripting interface implemented in Jython
-(see Table~\ref{tab:nengo-vers} for more detail).
+(as detailed in Table~\ref{tab:nengo-vers}).
 JavaNengo's simulator was created
@@
 before the scripting interface,
-the GUI manipulates simulator objects directly.
+the GUI manipulates simulator objects directly,
+rather than constructing a script
+that manipulates those simulator objects.
 Typically, modellers start by following
@@
 and transition to the scripting interface
-in order to implement more complicated models.
-Since the GUI modifies simulator objects directly,
+in order to implement more complicated models;
+since the GUI modifies simulator objects directly,
 modelling concepts learned in the GUI
@@
 Finally, perhaps the most significant limitation
-of the JavaNengo tools is their performance.
+of the JavaNengo tools is performance.
 While Java has good support for threads,
@@
 
-PyNengo,
-the subject of the remainder of this paper,
-is a Python package for
+PyNengo is a Python package for
 defining and simulating
@@
 
-PyNengo's object model is simple and minimal,
-making it easy to
+PyNengo has a simple object model,
+which makes it easy to
 document, test, and modify.
-It also includes a scripting interface to those objects
-based on JavaNengo's scripting interface \cite{TODO},
-but with several modifications
-to allow for easier incorporation
-of user-defined networks,
-new neuron models, and learning rules.
-Additionally, PyNengo's scripting interface
-contains methods to record simulated data
-and access that data after simulation.
+PyNengo's scripting interface to those objects
+is modified from JavaNengo's scripting interface \cite{TODO}
+to easily incorporate user-defined networks,
+new neuron models and learning rules,
+and multiple simulators.
+Additionally, the scripting interface
+contains functions to record data during a simulation,
+and access that data after the simulation.
 
@@
 Simulators can also be written
-to leverage specific platforms.
+to leverage special-purpose hardware.
 Currently, we have implemented
@@
 
-PyNengo overcomes all of JavaNengo's limitations.
+PyNengo overcomes JavaNengo's limitations.
 It can simulate most models
@@
   and PyNengo is 5000 lines of code. TODO}
-and depending on
+and depends on
 only three external libraries instead of ten.
@@
 Information is encoded by populations of neurons.
-We consider that information
-to be time-varying vectors of real numbers,
-in order to reason with that information
+We represent information
+with time-varying vectors of real numbers
+in order to manipulate that information
 using conventional mathematics.
@@
 each neuron based on its \textit{tuning curve},
-which describes how much
-that neuron fires for a given input signal.
+which describes how likely
+that neuron is to fire for a given input signal.
 The tuning curve summarizes the gain
@@
 (the part of the vector space
-that causes the neuron to fire most strongly).
+that causes the neuron to be the most active).
 This results in a predictable pattern of firing
@@
 In order to compute any function,
-we can set the connection weights between
+we set the connection weights between
 two populations to be the product of
@@
 Large models can be built
-by thinking of the principles of the NEF
-as building blocks that can be put together
-to describe neural systems.
+by using of the principles of the NEF
+as connectable components
+that describe neural systems,
+just as a circuit diagram
+describes an electronic circuit.
 The goal of Nengo is to allow
@@
 of 50 leaky integrate-and-fire neurons \cite{TODO} % LIF cite
-with a refractory time of 0.002 seconds.
+with a refractory time of 2 milliseconds.
 The \texttt{nengo.LIF} class defines
@@
 
+% Also acknowledge Bryan for making JavaNengo, and edits
+
 \paragraph{Funding\textcolon}

### 2013-10-02 nengo.tex
Edits from Eric's feedback

@@
 containing the characters $A3[1234]$ instructs Spaun
-to remember the list $1234$.
+to memorize the list $1234$.
 It then provides its answer by
@@
 tracing out digits.
-If asked to recall the remembered list,
+If asked to recall the memorized list,
 it would trace out the digits $1234$.
@@
 all tasks use a common set of
-functional modules that we map
-onto the brain areas hypothesized
-to perform those functions.
+functional modules.
+Each functional module corresponds
+to a brain area that has been hypothesized
+to perform those functions
+in neuroscientific literature
+(e.g., motor commands are represented
+in primary motor cortex).
 Like data-driven large scale models,
@@
 and see that the data emerges.
-This is made possible by a theoretical framework
-like the memory-prediction framework \cite{TODO},
-but unlike it and other frameworks,
-the NEF enables Spaun to employ
+This is possible due to the NEF,
+which, unlike the memory prediction framework
+and theoretical frameworks,
+allows Spaun to employ
 different sets of algorithms
@@
 
-JavaNengo is a suite of tools providing
+JavaNengo is a suite of tools composed of
 a neural simulator and graphical user interface (GUI)
@@
 and transition to the scripting interface
-in order to implement more complicated models;
+in order to implement more complex models;
 since the GUI modifies simulator objects directly,
@@
 This greatly enhanced productivity,
-and made it possible
-to create many models with very little code.
+allowing models to be described
+by short scripts.
 Modellers quickly started using
@@
 Addressing all of these issues
-in the Java codebase would involve
-significant enough changes
-that we started from scratch instead.
+in the Java codebase would require
+such extensive changes
+that we instead opted
+to write a new version of Nengo in Python.
 Since the scripting interface
@@
 we targeted the CPython interpreter
-as the platform for the next generation of Nengo.
+as the platform for the new version of Nengo.
 
@@
   \item \textbf{Representation:} A population of neurons
-    collectively represents information in the form of
-    time-varying vectors of real numbers.
+    collectively represents a time-varying vector of real numbers.
   \item \textbf{Transformation:} Functions on those vectors
@@
 require persistent activity through recurrent connections.
-When recurrent connection are introduced,
+When recurrent connections are introduced,
 the vectors that populations represent
@@
 describes an electronic circuit.
-The goal of Nengo is to allow
-modellers to describe neural systems
-in terms of what information is being represented,
-how that information is transformed,
-and what dynamics are required
-to perform cognitively relevant
-functions with biologically constrained
-neural components.
-Those descriptions are then
-translated to a network
-of interconnected neurons.
-This situates Nengo
-as a ``neural compiler''
+The goal of Nengo is to enable
+modellers to create those components
+and connect them.
+Components describe
+what information is being represented,
+and connections describe
+how that information is transformed
+over time.
+Nengo translates those descriptions
+to a network of interconnected neurons,
+situating it as a ``neural compiler''
 that translates
 a high-level functional model
-to a low-level neural model
-that implements the high-level function.
+to a low-level neural model.
 
@@
 
+To describe the high-level functional model,
+PyNengo defines six core objects.
+\begin{enumerate}
+  \item The \texttt{Model} encapsulates the model and
+    exposes a simplified interface to other objects.
+  \item The \texttt{Ensemble} contains a group of neurons
+    that encodes a time-varying vector of real numbers.
+  \item The \texttt{Node} represents non-neural information,
+    such as sensory inputs.
+  \item The \texttt{Connection} describes connections
+    between nodes and ensembles.
+  \item The \texttt{Network} is a group of interconnected
+    nodes and ensembles.
+  \item The \texttt{Probe} gathers data during a simulation
+    for later analysis.
+\end{enumerate}
+
 \subsection{Model}
@@
 A Nengo model is a collection
-of Nengo objects that represent information
-and connections that transform information.
+of Nengo objects, which represent information,
+and connections, which transform information.
 Objects and connections can be probed
@@
 in the form of a real-valued vector.
-An ensemble must be given as arguments
-its dimensionality
+When creating an ensemble,
+the user must provide the dimensionality
 (i.e., the length of the vector it represents),
 and an object that describes
-the population of neurons.
-For example, \texttt{nengo.LIF(50, tau\_ref=0.002)}
-describes a population
-of 50 leaky integrate-and-fire neurons \cite{TODO} % LIF cite
-with a refractory time of 2 milliseconds.
+a population of neurons.
+For example, \texttt{Ensemble('A', nengo.LIF(50, tau\_ref=0.002), 1)}
+describes an ensemble representing a one-dimensional vector
+with 50 leaky integrate-and-fire neurons \cite{TODO} % LIF cite
+that have a refractory period of 2 milliseconds.
 The \texttt{nengo.LIF} class defines
@@
 however, the simulator used must be aware
-of that neuron model.
+of that type of neuron.
 
@@
 
-Networks are designed to be
-subclassed by users,
+\textt{Network} is a base class designed to be
+subclassed by users
 in order to group together
@@
 several objects in the model can be
-moved into a \texttt{Network}
-subclass' \texttt{make}
-function with only minor changes.
+made into a Network subclass
+with only minor changes.
 
@@
 Decoupling model creation and simulation
-enables PyNengo simulators
+has been done previously
+by PyNN \cite{TODO}.
+In PyNN, the same Python script
+can be used to run a model
+on four different simulators.
+Nengo follows this programming model,
+decoupling model creation and simulation,
+which enables PyNengo simulators
 to allocate memory and schedule computations
@@
 of the reference simulator's architectural choices,
-but is heavily optimized in ways
-that cannot be adequately described
-with the space available.
+but the details of its implementation
+include OpenCL-specific optimizations
+that are beyond the scope of this paper.
 Additionally, it should be noted that
@@
 not as a specification.
+
+The reference simulator
+uses a reduced set of objects
+that describe computations
+occurring in the model.
+Specifically, the simulator
+contains signals, which represent values,
+and operators, which represent computations
+on signals.
+Figure~\ref{fig:sim} shows the signals
+and operators used in a simple model.
+
+\paragraph{Signals}
+
+A \texttt{Signal} represents any number that
+will be used by the simulator.
+Each high level Nengo object contains
+several signals;
+for example, an ensemble contains signals
+that represent the high-level input
+signal that will be encoded
+to input currents,
+and the encoding weights.
+It also contains a neural population,
+which contains signals that represent
+input currents, bias currents,
+membrane voltages, and refractory times for each cell.
+
+As can be seen in Figure~\ref{fig:sim},
+the signals used in a Nengo simulation
+can be conceptually grouped into
+those that track low-level neural signals,
+and those that track high-level signals
+defined by the NEF.
+Other neural simulators only track
+low-level signals.
+Operators commonly map
+between related low- and high-level signals.
+
+\paragraph{Operators}
+
+Operators represent computations
+to be performed on signals on each timestep.
+Once the model has been built,
+only a small set of mathematical
+operations are necessary for simulation.
+
+Many of the computations
+done in a simulation
+are linear transformations (e.g.,
+the decoding and encoding steps
+in Figure~\ref{fig:sim}),
+and therefore can share a common operator;
+this can be helpful for parallelizing computations.
+
+Nonlinear functions, however,
+require specific operators.
+Each supported neuron model and learning rule
+has an associated operator;
+the simulator explicitly maps
+from symbolic neuron objects in ensembles
+and from symbolic learning rule objects
+in connections to their associated operator.
+
+\paragraph{Reference simulator description}
 
@@
 the state of the reference simulator
-after two timesteps of a simple model.
-
-\paragraph{Signals}
-
-A \texttt{Signal} represents any number that
-will be used by the simulator.
-Each high level Nengo object contains
-several signals;
-for example, an ensemble contains signals
-that represent the high-level input
-signal that will be encoded
-to input currents,
-and the encoding weights.
-It also contains a neural population,
-which contains signals that represent
-input currents, bias currents,
-membrane voltages, and refractory times for each cell.
-
-As can be seen in Figure~\ref{fig:sim},
-the signals used in a Nengo simulation
-can be conceptually grouped into
-those that track low-level neural signals,
-and those that track high-level signals
-defined by the NEF.
-Most neural simulators only track
-low-level signals.
-Operators commonly map
-between related low- and high-level signals.
-
-\paragraph{Operators}
-
-Operators represent computations
-to be performed on signals on each timestep.
-Once the model has been built,
-only a small set of mathematical
-operations are necessary for simulation.
-
-Many of the computations
-done in a simulation
-are linear (e.g.,
-the decoding and encoding steps
-in Figure~\ref{fig:sim}),
-and therefore can share a common operator;
-this can be helpful for parallelization
-in some simulators.
-
-Nonlinear functions, however,
-require specific operators.
-Each supported neuron model and learning rule
-has an associated operator;
-the simulator explicitly maps
-from symbolic neuron objects in ensembles
-and from symbolic learning rule objects
-in connections to their associated operator.
-
-\section{Examples} \label{sec:examples}
+after two timesteps of a simple model;
+all subsequent timesteps perform the same
+operations as the first two.
+
+\section{Example scripts} \label{sec:examples}
+
+The scripting interface provides
+a simple way to add PyNengo objects to a model,
+simulate that model,
+and extract data collecting during the simulation.
+Rather than list the functions in the scripting interface,
+we instead provide three concrete example scripts
+that highlight the types of models that can be
+build with PyNengo.
+We have also implemented two of these three examples
+in PyNN to provide a comparison
+for the length and clarity of the code
+describing the models.
 

### 2013-10-02 nengo.tex
Edits from James's feedback

@@
 
-It is often said that neuroscience
-is data-rich but theory-poor.
-Although computational neuroscience
-has grown significantly,
-this phrase continues to be repeated
-due to a lack of satisfying
-theories of the brain.
+Modeling the brain of a human,
+or just about any animal at all,
+is one of the greatest
+scientific challenges of our time.
+However, despite decades of detailed
+experimental work in neurobiology
+and the recent growth of computational neuroscience,
+we lack a mathematical account of
+how biological and electrical components
+implement cognitive functions such as sensory processing,
+memory formation, reasoning, and motor control.
 Much work has been put into neural simulators
 that attempt to recreate neuroscientific
-data in precise detail
-with the thought that
+data in precise detail with the thought that
 cognition will emerge by connecting
 detailed neuron models according
-to the statistics of biological connections \cite{TODO}; % blue brain
-cognition has not yet emerged
+to the statistics of biological connections \cite{TODO}. % blue brain
+Cognition has not yet emerged
 from data-driven large scale models.
-Similarly, theoretical frameworks
-that use a generic algorithm
-replicated many times % reword based on papers below
-are unable to solve more complex problems
-than the generic algorithm alone
-(e.g., \cite{TODO} \cite{TODO}). % Hawkins, Kurzweil
+On the other end of the spectrum,
+cognitive architectures \cite{TODO} % act-r
+and theoretical frameworks \cite{TODO} % Hawkins, Kurzweil
+have seen success solving statistical problems,
+but do so without respecting biological constraints.
+Additionally, the statistical problems
+solved by these frameworks are better solved
+by machine learning approaches.
 
@@
 (NEF).
-Nengo was recently used to implement Spaun,
-currently the world's largest functional brain model
-\cite{TODO}.
+The NEF is a large-scale modeling approach
+that can leverage any neuron model
+to build neural networks with
+demonstrable cognitive abilities.
+Nengo and the NEF has been used to build
+increasingly sophisticated neural subsystems
+for the last decade (e.g., subsystem \cite{TODO}),
+culminating recently with Spaun,
+currently the world's
+largest functional brain model \cite{TODO}.
 Spaun is a network of 2.5 million spiking neurons
@@
 it would trace out the digits $1234$.
-
 While the tasks Spaun performs are diverse,
@@
 to perform those functions
-in neuroscientific literature
-(e.g., motor commands are represented
-in primary motor cortex).
-Like data-driven large scale models,
-Spaun is able to match experimental data
-across the neuroscientific spectrum,
-from the single cell to the behavioral level.
-However, instead of constructing the model
-based on data and hoping behavior emerges,
-we construct the model
-based on behavior
-and see that the data emerges.
-This is possible due to the NEF,
-which, unlike the memory prediction framework
-and theoretical frameworks,
-allows Spaun to employ
-different sets of algorithms
-depending on the behavioral task performed,
-rather than defining a single algorithm
-and hypothesizing that replicating it
-many times will enable these behaviors.
-
-In implementing Spaun,
-we provided strong support
-that the NEF is beginning to fill the theory void in neuroscience.
-However, in doing so, we pushed Nengo
-to the limits of its ability to construct and simulate
-large-scale neural models.
+in neuroscientific literature.
+
+The NEF shows how to
+construct a neural model
+from anatomical constraints,
+cognitive objectives, and
+dynamical systems theory.
+Perhaps surprisingly,
+constructing models from this starting point,
+rather than from single cell electrophysiology
+and connectivity statistics
+has not meant sacrificing
+explanatory power.
+Instead, single cell activity,
+response timing, and behavioral errors
+of NEF-designed models match
+psychological and physiological findings
+without being built specifically into the design.
+These quirks seem to fall directly out
+of the need to satisfy competing objectives
+within anatomical and neurobiological constraints.
+The opposite phenomenon in which
+cognitively useful behaviors emerge
+from networks of detailed neurons
+connected with plausible connectivity statistics
+has not yet been observed.
+
+It has often been said that neuroscience
+is data-rich and theory-poor.
+The modeling success of
+large functional networks such as Spaun
+indicates that the NEF
+is beginning to fill the theory void.
+However, in simulating these large networks,
+we pushed our Java-powered simulator
+to its limits in terms of memory consumption,
+simulation speed, and extensibility.
 In order to continue to evaluate the NEF
@@
 extensibility, and simplicity.
+
+This paper introduces a new version
+of Nengo (PyNengo) written in Python.
+The remainder of this introduction outlines
+the design issues in the previous version (JavaNengo)
+that motivated the reimplementation.
+PyNengo is not currently meant
+to completely replace JavaNengo;
+JavaNengo is a GUI application
+and it remains a capable visual NEF
+simulator and valuable teaching tool.
+PyNengo's focus is on speed
+and a simplified internal design
+to support ongoing research in
+large adaptive functional models.
+The remainder of the paper outlines PyNengo's design,
+scripting interface, and performance relative to
+alternatives such as Brian and NEST.
 
@@
 While Java has good support for threads,
-it does not have well-supported libraries
+it lacks native high-performance numerical libraries
 for doing mathematical operations on large matrices quickly.
@@
 graphical processing units (GPUs),
-which can perform the types of computations
-that neural models require quickly and in parallel.
+which are well-suited to the computations
+required of neural simulator.
 
@@
 to write a new version of Nengo in Python.
-Since the scripting interface
+JavaNengo's Jython scripting interface
 already used Python syntax,
 and since Python is widely used
-for neuroscientific data analysis,
+for neural simulation and data analysis,
 we targeted the CPython interpreter
@@
 neural models using the NEF.
-It targets the CPython interpreter,
-meaning that it integrates seamlessly
-with other scientific Python tools.
+Its interfaces are based on NumPy,
+so it integrates seamlessly
+with other scientific Python tools
+for interactive exploration \cite{TODO}, %Ipython
+scientific computation \cite{TODO}, %scipy
+and plotting \cite{TODO}. % Matlab
 Arbitrary CPython programs
@@
 for using neurally implemented algorithms
-in applications and games.
+in web services, games, and other applications.
 
@@
 document, test, and modify.
-PyNengo's scripting interface to those objects
-is modified from JavaNengo's scripting interface \cite{TODO}
-to easily incorporate user-defined networks,
-new neuron models and learning rules,
-and multiple simulators.
-Additionally, the scripting interface
-contains functions to record data during a simulation,
-and access that data after the simulation.
+These objects are accessible
+through a scripting interface
+similar to JavaNengo's \cite{TODO}, % frontiers 2009
+with improvements
+for easier incorporation of user-defined networks,
+neuron models and learning rules.
+Critically, and unlike JavaNengo,
+PyNengo's scripting interface
+contains functions to record simulated data
+and access that data after simulation.
 
@@
 allowing for models to be exported
-to other simulators,
-such as those accessible with PyNN.
-Simulators can also be written
-to leverage special-purpose hardware.
-Currently, we have implemented
-a platform-independent simulator
-that uses NumPy for vectorized computations,
-and a PyOpenCL-based
-simulator that takes full advantage of GPUs
-and multicore CPUs.
+to other simulation packages
+(e.g., those accessible by PyNN)
+or for faster simulators to be used
+as drop-in replacements for PyNengo's
+platform-independent reference simulator.
+To date, we have implemented one other simulator
+that uses PyOpenCL to take advantage
+of a GPU or multicore CPU.
 
@@
 replace all of JavaNengo's functionality.
-While it can simulate the vast majority
-of models that JavaNengo can simulate,
-some functionality is still in development.
-Most notably, there is no graphical interface
-to PyNengo, and therefore simulations
-cannot be visualized while they are running.
-However, we are confident that these tools
-can be created, and will be much easier
-to implement compared to JavaNengo's GUI.
-
-PyNengo overcomes JavaNengo's limitations.
+While it can simulate most
+models that JavaNengo can simulate,
+there is no graphical interface to PyNengo.
+Model components must be created
+using Python scripts rather than
+the drag-and-drop interface presented by JavaNengo.
+Some of the means of interacting
+with a running model in JavaNengo
+can be provided using packages
+such as IPython and Matplotlib,
+but such tools have not been our focus thus far.
+It is possible to build a graphical program
+like JavaNengo's GUI on top of PyNengo's object model,
+but that remains future work.
+
+To summarize, PyNengo overcomes
+JavaNengo's limitations as exposed by large-scale modelling research.
 It can simulate most models
 currently simulated by JavaNengo
-using 11\% as many lines of code,\footnote{
-  JavaNengo is 40,000 lines of code, not including the GUI,
-  and PyNengo is 5000 lines of code. TODO}
-and depends on
-only three external libraries instead of ten.
-The OpenCL simulator can simulate
-a large model TODO times
-faster than JavaNengo's simulator.
-This, along with a codebase
-that is well tested and
-easy to understand and extend,
-provides a platform
-that can simulate larger and more complex
-models than Spaun,
-and can therefore further
-test the NEF as a theory of the brain.
+using a well-tested codebase of
+TODO lines of code,
+compared to JavaNengo's TODO
+(which does not include the JavaNengo GUI).
+A new internal representation of neural models
+allows the OpenCL simulator to simulate
+large models on the scale of Spaun
+hundreds of times faster than JavaNengo's simulator
+using inexpensive commodity hardware.
+Compatibility with CPython makes scientific Python software
+available for preparing and analyzing data.
+In all, PyNengo provides a platform for
+simulating larger and more complex models than Spaun,
+and can therefore further test NEF
+as a theory of the brain.
 
@@
   \item \textbf{Transformation:} Functions on those vectors
-    are computed by the connections between populations.
+    are computed by the connections between populations of neurons.
   \item \textbf{Dynamics:} The vectors represented
@@
 causing it to spike.
-We can then estimate
-the originally encoded vector
+With enough neurons,
+we can estimate the originally encoded vector
 through a \textit{decoding} process.
@@
 which describes how likely
-that neuron is to fire for a given input signal.
-The tuning curve summarizes the gain
+that neuron is to respond to a given input signal.
+The tuning curve is a function of the gain
 of a neuron (how quickly the activity rises),
 the bias (the activity of a neuron given no signal),
-and the encoding weight
-(the part of the vector space
+and the \textit{encoding weight}
+(the direction in the input vector space
 that causes the neuron to be the most active).
-This results in a predictable pattern of firing
-given the input signal.
+% This results in a predictable pattern of firing
+% given the input signal.
 
@@
 and the decoded estimate.
-Note that these decoding weights are not determined
+Note that these decoding weights
+need not be determined
 based on the input signal;
-instead, the minimization occurs
-on points that are randomly sampled
-in the vector space
+instead, we typically perform
+this minimization on points
+randomly sampled from
+the portion of the vector space
 that the population represents.
 
-\paragraph{Transformation}
+\paragraph{Transformation}  % NEF as neural model...
 Neurons communicate through
@@
 for the downstream neuron.
+In other words, the NEF hypothesizes
+that synaptic weight matrices
+have low rank, and can be factored.
 
@@
 and instead store
-the encoding and decoding weights
-(i.e., the factors of the connection weight matrix).
+the encoding and decoding weights separately
+(i.e., the two factors of the connection weight matrix).
 This provides significant
-space and time savings.
+space and time savings in the simulation loop.
 
 \paragraph{Dynamics}
-While doing feedforward transformations
-on vectors is sufficient to model
-many neural systems,
+While feedforward vector transformations
+suffice to describe
+some neural systems,
 many systems, notably memory systems,
@@
 When recurrent connections are introduced,
-the vectors that populations represent
+the vectors represented by neural populations
 can be thought of as state variables
 in a dynamical system.
-Neural systems can therefore
-be built and analyzed using
-the methods of control theory.
+The equations governing dynamics
+in such a system
+can be designed and analyzed
+using the methods of control theory,
+and translated into neural circuitry
+using the principles
+of representation and transformation.
 
@@
 
-\paragraph{Nengo}
+\paragraph{NEF and Nengo}
+% The NEF makes it possible
+% to design large systems
+% because its three principles
+% show how to compose functional subsystems
+% into a coherent whole.
 Large models can be built
@@
 
-\textt{Network} is a base class designed to be
+\texttt{Network} is a base class designed to be
 subclassed by users
@@
 creating a new network is simple,
-while creating a new simulator is complex.
+while creating a new simulator is complex, but possible.
 
 By enabling outside contributions,
-and providing a user friendly API
+and providing a user-friendly API
 and many example models,

### 2013-10-03 nengo.tex
Full edit of paper for flow.

@@
   proposes one such theory,
-  but has not has not gathered
-  significant empirical support
-  due, in part, to the technical
+  but has not yet gathered
+  significant empirical support,
+  partly due to the technical
   challenge of building and simulating
@@
   Nengo is a tool that builds and simulates
-  functional brain models with the NEF.
+  large-scale models with the NEF.
   Nengo is the primary resource
@@
   is used to model neural systems,
-  and for researchers using the NEF
-  to build and simulate
-  functional brain models.
-  A previous version of Nengo was used
+  and for doing research
+  comparing NEF models
+  to experimental data.
+  A previous version of Nengo
+  implemented in Java was used
   to create Spaun,
   the world's largest functional brain model.
-  The implementation of Spaun
+  Simulating Spaun
   identified limitations
@@
   and collect data for subsequent analysis.
-  We have implemented a new version of Nengo
-  in Python that overcomes these limitations.
+  This paper describes a new version of Nengo
+  implemented in Python that overcomes these limitations.
   It uses simple and extendable syntax,
@@
 
-Modeling the brain of a human,
-or just about any animal at all,
+Modeling the human brain
 is one of the greatest
 scientific challenges of our time.
-However, despite decades of detailed
-experimental work in neurobiology
-and the recent growth of computational neuroscience,
-we lack a mathematical account of
+Computational neuroscience
+has made significant advancements;
+however, we still lack a mathematical account of
 how biological and electrical components
@@
 detailed neuron models according
-to the statistics of biological connections \cite{TODO}. % blue brain
+to the statistics of biological synapses \cite{TODO}. % blue brain
 Cognition has not yet emerged
@@
 and theoretical frameworks \cite{TODO} % Hawkins, Kurzweil
-have seen success solving statistical problems,
-but do so without respecting biological constraints.
-Additionally, the statistical problems
-solved by these frameworks are better solved
-by machine learning approaches.
+have solved high-dimensional statistical problems,
+but do so without respecting biological constraints,
+and with worse performance than machine learning approaches.
 
@@
 increasingly sophisticated neural subsystems
-for the last decade (e.g., subsystem \cite{TODO}),
+for the last decade (e.g., \cite{TODO}),
 culminating recently with Spaun,
@@
 to memorize the list $1234$.
-It then provides its answer by
-sending motor commands to a simulated arm model,
-tracing out digits.
 If asked to recall the memorized list,
-it would trace out the digits $1234$.
+it sends motor commands to a simulated arm model,
+tracing out the digits $1234$.
 While the tasks Spaun performs are diverse,
@@
 rather than from single cell electrophysiology
-and connectivity statistics
-has not meant sacrificing
-explanatory power.
-Instead, single cell activity,
+and connectivity statistics,
+still produces simulated data that explains
+and predicts experimental results.
+Single cell activity,
 response timing, and behavioral errors
@@
 within anatomical and neurobiological constraints.
-The opposite phenomenon in which
+The opposite phenomenon, in which
 cognitively useful behaviors emerge
 from networks of detailed neurons
-connected with plausible connectivity statistics
+connected with plausible connectivity statistics,
 has not yet been observed.
@@
 is beginning to fill the theory void.
-However, in simulating these large networks,
+However, in simulating this large model,
 we pushed our Java-powered simulator
@@
 to completely replace JavaNengo;
-JavaNengo is a GUI application
-and it remains a capable visual NEF
-simulator and valuable teaching tool.
+JavaNengo remains a capable visual NEF
+simulator and a valuable teaching tool
+due to its graphical interface.
 PyNengo's focus is on speed
@@
 scripting interface, and performance relative to
-alternatives such as Brian and NEST.
+JavaNengo and alternatives such as Brian and NEST.
 
@@
 to extend those NEF components.
-As a representative example,
+For example,
 the most often used class for making
@@
 
-The JavaNengo GUI provided
+The JavaNengo GUI provides
 a drag-and-drop interface in which to create NEF models.
@@
 described in \cite{TODO},
-provided a simpler interface
+provides a simpler interface
 to the NEF-specific portions of the simulator.
@@
 Finally, perhaps the most significant limitation
-of the JavaNengo tools is performance.
+of JavaNengo is its performance.
 While Java has good support for threads,
 it lacks native high-performance numerical libraries
-for doing mathematical operations on large matrices quickly.
+for doing mathematical operations on large matrices.
 Additionally, it is difficult to interface with
-platform-specific tools such as
-graphical processing units (GPUs),
+hardware such as graphical processing units (GPUs),
 which are well-suited to the computations
-required of neural simulator.
+required in neural simulations.
 
@@
 in the Java codebase would require
-such extensive changes
+so many extensive changes
 that we instead opted
@@
 To summarize, PyNengo overcomes
-JavaNengo's limitations as exposed by large-scale modelling research.
+JavaNengo's limitations.
 It can simulate most models
@@
 large models on the scale of Spaun
-hundreds of times faster than JavaNengo's simulator
+hundreds of times faster than JavaNengo
 using inexpensive commodity hardware.
@@
 simulating larger and more complex models than Spaun,
-and can therefore further test NEF
+and can therefore further test the NEF
 as a theory of the brain.
@@
    with the input signal in the middle panel,
-   we see the firing pattern in the bottom panel.
+   we see the spike trains in the bottom panel.
    (B) By the representation principle,
@@
    and the cosine of the input signal (bottom panel, purple).
-   (C) By the transformation principle,
+   (C) A sine wave is encoded by population A (top panel);
+   the negative of that signal is projected
+   to population B (middle panel)
+   and the square of that signal is projected
+   to population C (bottom panel).
+   By the transformation principle,
    populations of neurons can send signals
@@
    into the second population.
-   These two steps can be expanded into a single step
+   These two steps can be combined into a single step
    by calculating a set of weights
-   that describe the strength of the connection
+   that describe the strength of the connections
    between each neuron in the first population
    and each neuron in the second population.
-   In (C), a sine wave is encoded by population A (top panel);
-   the negative of that signal is projected
-   to population B (middle panel)
-   and the square of that signal is projected
-   to population C (bottom panel).
-   (D) By the dynamcs principle,
-   signals being represented by population of neurons
-   can be thought of
-   as state variables in a dynamical system.
-   In this example, the dynamical system
+   (D) A neurally implemented dynamical system
    has negative feedback across its two dimensions,
@@
    The oscillator is plotted across time (top)
-   and in state space (bottom).}
+   and in state space (bottom).
+   By the dynamcs principle,
+   signals being represented by population of neurons
+   can be thought of as state variables in a dynamical system.
+   }
 \end{figure}
@@
 Figure~\ref{fig:nef} depicts
-the encoding (A) and decoding (B) process.
+the encoding (A) and decoding (B) processes.
 In the encoding process, the input signal drives
@@
 by solving a least-squares minimization problem.
-In the bottom panel of Figure~\ref{fig:nef}B,
-the red line is the filtered spiking activity
-weighted by decoding weights
-that minimize the difference between
-the actual encoded vector
-and the decoded estimate.
-The purple line is the filtered spiking activity
-weighted by decoding weights
-that minimize the different between
-the cosine of the actual encoded vector
-and the decoded estimate.
 Note that these decoding weights
-need not be determined
-based on the input signal;
+do not necessarily depend on the input signal;
 instead, we typically perform
@@
 randomly sampled from
-the portion of the vector space
+the vector space
 that the population represents.
@@
 it releases neurotransmitter across the synapse,
-which causes some amount of current
+which typically causes some amount of current
 to be imparted in the downstream neuron.
 Many factors affect the
-amplitude of the current imparted;
+amplitude of the imparted current;
 we summarize those factors
@@
 the decoding weights for that function
-for neurons in the first population
+in the first population
 and the encoding weights
-for the downstream neuron.
-In other words, the NEF hypothesizes
+for the downstream population.
+
+This implies the NEF hypothesizes
 that synaptic weight matrices
 have low rank, and can be factored.
-
-In Figure~\ref{fig:nef}C,
-a sine wave is transformed twice.
-Between populations A and B,
-the signal is negated.
-Between populations B and C,
-the signal is squared.
-
 Note that, in practice, we rarely use
@@
 This provides significant
-space and time savings in the simulation loop.
+space and time savings during simulation.
 
@@
 some neural systems,
-many systems, notably memory systems,
-require persistent activity through recurrent connections.
+many require persistent activity through recurrent connections.
 When recurrent connections are introduced,
@@
 
-In Figure~\ref{fig:nef}D,
-a population represents a two-dimensional vector,
-and is recurrently connected
-with negative feedback between dimensions,
-resulting in a harmonic oscillator.
-
 \paragraph{NEF and Nengo}
-% The NEF makes it possible
-% to design large systems
-% because its three principles
-% show how to compose functional subsystems
-% into a coherent whole.
 Large models can be built
@@
 The goal of Nengo is to enable
-modellers to create those components
-and connect them.
+modellers to create and connect those components.
 Components describe
@@
 and connections describe
-how that information is transformed
-over time.
+how that information is transformed.
 Nengo translates those descriptions
@@
 \begin{enumerate}
-  \item The \texttt{Model} encapsulates the model and
+  \item The \texttt{Model} encapsulates the entire model and
     exposes a simplified interface to other objects.
@@
     such as sensory inputs.
-  \item The \texttt{Connection} describes connections
-    between nodes and ensembles.
+  \item The \texttt{Connection} describes how
+    nodes and ensembles are connected.
   \item The \texttt{Network} is a group of interconnected
@@
 \subsection{Model}
-
-A Nengo model is a collection
-of Nengo objects, which represent information,
-and connections, which transform information.
-Objects and connections can be probed
-in order to collect data during simulation.
 
@@
 with the \texttt{Model} object
-(see sections \ref{sec:comm-channel}--\ref{sec:lorenz}
+(see sections \ref{sec:comm-channel} \& \ref{sec:lorenz}
 for examples).
@@
 these additional attributes
-allow modellers to fit
-neural parameters to \textit{in vivo}
+allow modellers to set
+neural parameters according to \textit{in vivo}
 electrophysiology data.
@@
 when computing its function.
-It can interact directly with hardware,
-interface with other programs
-using shared memory or sockets,
-and so on.
+For example, it can interact directly with hardware,
+and interface with other programs
+using shared memory or sockets.
 
@@
 in several ways.
-Like other objects in PyNengo's object model,
+Like other PyNengo objects,
 a connection contains symbolic information;
 in this case, a connection tracks
-how two objects are to be connected together.
+how two objects are connected together.
 
@@
 allows ensembles to project
-the information they're representing---or
+encoded information---or
 a transformation of that information---to
@@
 to synaptic weights.
-
-The ability to connect neurons together
-with synaptic connection weights is still possible
-with the \texttt{NonlinearityConnection}.
+By default, ensembles are connected
+with a \texttt{DecodedConnection}.
+
+However, neurons in an ensemble can be directly connected
+to neurons in another ensemble
+with synaptic connection weights
+using the \texttt{NonlinearityConnection}.
 Both of these connection types
@@
 
-In many cases, the type of the connection
-is not explicitly specified by the modeller.
-Instead, it is inferred by the arguments
-to \texttt{Model}'s \texttt{connect} function.
-However, that function is just a simple wrapper
-of each connection object's constructor,
-so complex models can use
-the appropriate connection,
-or even implement their own.
-
 \subsection{Network}
 
-A network is a collection of ensembles and nodes
-connected together in a particular way.
-It is primarily a way of grouping together
-a common set of connected objects,
-and makes explicit the purpose
-of the network and how it interacts
-with the rest of the model
-(see section \ref{sec:ccon} for a usage example).
-
+A network is a collection of interconnected ensembles and nodes.
+It is primarily a way of grouping
+a set of connected objects
+that work together to perform a complex function.
+Encapsulating them in a network
+makes its purpose explicit
+and hides the complexity of the function
+(see section \ref{sec:cconv} for an example).
 \texttt{Network} is a base class designed to be
-subclassed by users
-in order to group together
-parts of a model.
-In most cases,
-the code that creates and connects
-several objects in the model can be
-made into a Network subclass
+subclassed by modellers.
+The code that creates and connects
+several objects in a model can be
+grouped into a \texttt{Network} subclass
 with only minor changes.
@@
 
-A \texttt{Probe} marks that a signal
-will be recorded
-throughout the simulation.
+A \texttt{Probe} denotes that
+a particular part of another object
+will be recorded throughout a simulation.
 Nengo models contain many variables
@@
 including membrane potentials,
-spike events, and represented variables.
+spike events, and encoded vectors.
 It is resource intensive to store the values of
 large numbers of variables
-at each time step, and it is also not necessary,
+at each timestep, and it is also not necessary,
 as typically only a small fraction
@@
 directly in memory,
-but adding the ability to store data
-in files or to stream probed data
+but the ability to store data
+in files or to stream data
 directly to sockets will be added.
-
-\section{PyNengo simulators} \label{sec:simulators}
 
@@
    (A) Code describing the model being simulated.
-   It consists of ensemble A project the sine of its
+   It consists of ensemble A projecting the sine of its
    encoded signal to ensemble B,
@@
    (C) A detailed diagram of how the reference simulator
-   simulates this model. Signals, which are the values
-   tracked in the simulation, are colored blue.
-   Operators, which are the computations done on signals,
-   are colored red.
+   organizes this model. Signals (blue) represent the values
+   tracked in the simulation.
+   Operators (red) represent the computations done on signals.
    Signals can be grouped as low-level neural signals
@@
    of recurrently connected neurons.
-   The connection weight between the two populations,
+   The connection weights between the two populations,
    and from the second population to itself,
@@
    that can be analyzed with control theory.
-   A simply represents its input,
-   and passes its state to a sin function
+   State variable A simply represents its input,
+   and passes its state to a sine function
    which becomes the input to B.
-   B is a simple linear system
+   State variable B is a simple linear system
    that can be described with the typical
@@
 
+\section{PyNengo simulators} \label{sec:simulators}
+
 Decoupling model creation and simulation
@@
 on four different simulators.
-Nengo follows this programming model,
+Nengo follows this programming model by
 decoupling model creation and simulation,
@@
 this \texttt{Model} is a symbolic description,
-which a simulator can assume will not change
-over the course of a simulation.
+which a simulator can assume will not change.
 The simulator, however,
@@
 will fill in many of the details
-that the symbolic description
-has allowed to be generated randomly.
+not specified at the symbolic level.
 
@@
 and operators, which represent computations
-on signals.
+performed on signals.
 Figure~\ref{fig:sim} shows the signals
@@
 will be used by the simulator.
-Each high level Nengo object contains
+Each high-level Nengo object contains
 several signals;
@@
 
+\clearpage
+
 \paragraph{Operators}
@@
 this can be helpful for parallelizing computations.
-
 Nonlinear functions, however,
@@
   \item computes each operator in order, and
-  \item copies probed signals to a separate buffer.
+  \item copies probed signals to memory.
 \end{enumerate}
@@
 simulate that model,
-and extract data collecting during the simulation.
+and extract data collected during the simulation.
 Rather than list the functions in the scripting interface,
@@
 that highlight the types of models that can be
-build with PyNengo.
+built with PyNengo.
 We have also implemented two of these three examples
@@
 A communication channel
-is a minimal example demonstrating
+is a simple example demonstrating
 the representation and transformation principles
@@
    the functions that generate these quantities in Nengo
-   can be decoupled, but are not included here for brevity.
-   The implementation of the Lorenz attractor
-   is not shown, but is very similar to the communication channel.}
+   can be decoupled, but are not included here for brevity.}
 \end{figure}
@@
    (B) Diagram depicting the model. The state ensemble
-   is recurrently connected with complex dynamics
-   implemented the chaotic Lorenz attractor.
-   Not that this populations does not receive
+   is recurrently connected with a complex function
+   implementing the dynamics of the Lorenz attractor.
+   Note that this populations does not receive
    any input that might drive its initial value;
@@
 attractor networks \cite{TODO}.
-Figure~\ref{fig:lorez} depicts
+Figure~\ref{fig:lorenz} depicts
 a Nengo implementation of the Lorenz chaotic attractor
@@
 composed of 2000 leaky integrate-and-fire neurons.
-
 We have implemented the Lorenz attractor
 in PyNN for benchmarking purposes
-(see supplementary material).
+(code not shown).
 
@@
    (A) Diagram depicting the model.
-   The input vectors, A and B, represent four-dimensional signal
+   The input vectors, A and B, represent four-dimensional vectors
    which are mapped onto six ensembles within the
    Circular Convolution network through
-   a complicated transformation matrix.
+   complicated transformation matrices.
    Each ensemble within the network represents a
-   two-dimensional signal. The product of those two signals
+   two-dimensional vector. The product of the two dimensions
    is projected through another complicated transformation matrix
@@
    A, B, and Result ensembles
-   (top, middle, bottom panels respectively).
+   (top, middle and bottom panels, respectively).
    The ensembles represent the correct values,
@@
 Circular convolution is how
-the semantic pointer architecture
-(SPA; \cite{TODO})
-binds two semantic pointers together.
-Binding is central to many
+Spaun binds two concepts together.
+Binding is central to solving many
 of the high level cognitive tasks
@@
 rather than in a single connection.
-This two-layer network construction is simplified
-by the existence of the \texttt{CircularConvolution} network.
+This two-layer organization is simplified
+the \texttt{CircularConvolution} network subclass.
 The complexity encapsulated in that network
@@
 
-\section{Benchmarks}
+\section{Benchmarks} \label{sec:benchmark}
 
@@
      Figure \arabic{figure}.}{
-     For (A) and (B), all of the simulators
+     Benchmark results for several simulators
+     on the example models described
+     in section~\ref{sec:examples}.
+     In (A) and (B), all of the simulators
      except the PyNengo OpenCL simulator
-     was run on an Intel Core i7-965.
+     were run on an Intel Core i7-965.
      JavaNengo used all 4 cores of this processor;
@@
      was run on an NVidia GTX280 GPU.
-     For (C), if not noted, the simulator used
+     (A) Benchmark results from simulating
+     the communication channel for 10 simulated seconds
+     at a 1 millisecond timestep.
+     For all model sizes, PyNengo simulators
+     are faster than JavaNengo,
+     which is significantly faster than NEURON and NEST,
+     which are significantly faster than Brian.
+     The full Brian results are not shown;
+     for the largest model, Brian takes $\sim$768 seconds.
+     (B) Benchmark results from simulating
+     the Lorenz attractor for 10 simulated seconds
+     at a 1 millisecond timestep.
+     For most model sizes,
+     the results are the same as (A),
+     except that NEURON is notably faster.
+     However, when analyzed, the NEURON model
+     emits almost no spikes after the first
+     simulated second, which may be
+     why the simulation is relatively fast.
+     The full results for Brian and NEST are not shown;
+     for the largest model, they take $\sim$1467
+     and $\sim$601 seconds respectively.
+     (C) Benchmark results from simulating
+     circular convolution for 1 simulated second
+     at a 1 millisecond timestep.
+     For the blue lines, the simulator used
      was the PyNengo OpenCL simulator.
@@
      only used one core.
+     For all model sizes,
+     the OpenCL simulator is faster
+     than the JavaNengo simulator,
+     which is faster than the PyNengo
+     reference simulator.
+     The reference simulator was only run
+     up to 50 dimensions.
+     The full results for JavaNengo
+     are not shown; for the largest model,
+     it takes $\sim$45 seconds.
      }
@@
 increasing simulation speed
-was a primary motivators in creating PyNengo.
+was a primary motivator in creating PyNengo.
 To validate that performance has improved,
@@
 indicating that these means are robust.
-The results, shown in Figure~\ref{fig:benchmarks},
+The results, shown in Figure~\ref{fig:benchmarks}A and B,
 suggest that all versions of Nengo are significantly
@@
 through PyNN, especially
-as the size of the model increases.
+as the size of models increase.
 This is likely due to Nengo's
@@
 The results also suggest that PyNengo's
-OpenCL simulator is faster
-than the reference simulator
-and JavaNengo's simulator,
-though the difference
-is not overwhelming in these models.
+simulators are faster than JavaNengo's simulator.
 
@@
 In this case, only versions of Nengo
-were tested, and were tested rigorously.
+were tested.
 Instead of running each simulation multiple times,
 we instead run the simulator for 10 timesteps
-in order to fill various levels of the cache,
+in order to fill various levels of CPU or GPU cache,
 and then run the simulator for 1000 more timesteps;
 there is very little variance using this method.
-As can be seen in Figure~\ref{fig:benchmarks},
+As can be seen in Figure~\ref{fig:benchmarks}C,
 for large models, the OpenCL simulator
@@
 and the OpenCL simulator on a CPU
-both use all available CPU cores,
+use all available CPU cores,
 PyNengo's OpenCL simulator is significantly faster.
@@
 on a theoretical framework
-that has already been used to make
-large-scale functional brain models.
+that has been validated
+through the Spaun model.
 
@@
 interface similar to the
-high-level objects in PyNengo.
+high-level object model in PyNengo.
 Rather than implementing its
-own simulator, PyNN uses existing
+own simulator, PyNN uses four existing
 simulators with Python bindings,
@@
 function=square)} connects ensemble A
-to ensemble B, projecting the square of
+to ensemble B, transmitting the square of
 the value represented by A to B.
@@
 
-As shown in Figure~\ref{TODO},
-Nengo is able to simulate many networks
-faster than all of PyNN's simulators.
+As shown in Figure~\ref{fig:benchmarks},
+Nengo is able to simulate many models
+faster than each of PyNN's simulators.
 This is, in part,
@@
 rather than storing the entire matrix.
-
-However, PyNN also contains functionality
-not currently implemented in Nengo.
-PyNN can assign spatial information
-to neurons in a population,
-which can be used to influence
-how those neurons connect to other populations.
-PyNN simulates many learning rules
-and neuron types,
-and is therefore better suited to
-simulate networks of detailed neuron models.
+However, PyNN is able to simulate
+many detailed neuron models
+and learning rules,
+and has access to the wealth of models
+already created for each of its
+accessible simulators.
+PyNN is therefore better suited to
+simulate detailed neural models.
 
@@
 
-Nengo is a young project that
+PyNengo is a young project that
 has started with a deliberately minimal base
-in order to make contributing easier than in
+in order to make development easier than in
 previous implementations.
@@
 we hope to foster a rich community of modellers
-contributing networks, nonlinearities,
-and simulators to Nengo
-and to theoretical neuroscience
-as a whole.
-Internally, we will bring the functionality
-of PyNengo to parity with JavaNengo
-by implementing the use cases
+contributing networks, neuron types, learning rules,
+and simulators to Nengo.
+Internally, we will
+implementing the JavaNengo use cases
 not currently covered by PyNengo,
@@
 other scientific Python tools.
-
 While the reference simulator
@@
 it can simulate a large circular convolution model
-TODO times faster than JavaNengo's simulator.
+200 times faster than JavaNengo.
 This makes the creation and simulation
-of models many times larger than Spaun
-tractable with current hardware;
-these models will further test
-the NEF as a theory of the brain,
-and with PyNengo make those theories
-testable by anyone with a computer.
+of models that are many times larger than Spaun
+tractable with current hardware.
+These models will further test
+the NEF as a theory of the brain;
+PyNengo will make those models
+accessible to anyone with a modern computer.
 
@@
 
-TODO
+Trevor Bekolay led development of the
+PyNengo object model and scripting interface,
+wrote the text of the paper,
+and prepared all of the figures.
+James Bergstra led development of
+the PyNengo reference simulator and OpenCL simulator,
+edited text, created an early version
+of Figure~\ref{fig:sim},
+and ran the benchmarks shown in
+Figure~\ref{fig:benchmarks}C.
+Eric Hunsberger contributed significantly
+to PyNengo and both of its simulators, and edited text.
+Travis DeWolf led development
+of a Theano-backed
+version of PyNengo that
+identified issues with Theano,
+and provided the base for
+the version of PyNengo described in this paper.
+Terrence C. Stewart contributed
+to PyNengo and helped implement
+the PyNN scripts used in
+sections~\ref{sec:examples} and \ref{sec:benchmark}.
+Daniel Rasmussen contributed
+to PyNengo and the reference simulator,
+and edited text.
+Xuan Choo contributed to PyNengo
+and the reference simulator.
+Aaron Voelker ran the benchmarks
+shown in Figure~\ref{fig:benchmarks}A and B.
+Chris Eliasmith oversaw all development,
+contributed to PyNengo,
+and co-created the NEF with Charles Anderson.
 

### 2013-10-03 nengo.tex
Added references

@@
 detailed neuron models according
-to the statistics of biological synapses \cite{TODO}. % blue brain
+to the statistics of biological synapses \citep{markram2006}.
 Cognition has not yet emerged
@@
 On the other end of the spectrum,
-cognitive architectures \cite{TODO} % act-r
-and theoretical frameworks \cite{TODO} % Hawkins, Kurzweil
+cognitive architectures \citep{anderson2004}
+and theoretical frameworks \citep{hawkins2006}
 have solved high-dimensional statistical problems,
@@
 based on a theoretical framework proposed
-by Eliasmith \& Anderson \citeyearpar{TODO}
+by \citet{eliasmith2003}
 called the Neural Engineering Framework
@@
 increasingly sophisticated neural subsystems
-for the last decade (e.g., \cite{TODO}),
+for the last decade
+(e.g., path integration [\citealp{conklin2005}],
+working memory [\citealp{singh2006}],
+list memory [\citealp{choo2010}],
+inductive reasoning [\citealp{rasmussen2011}],
+motor control [\citealp{dewolf2011}],
+decision making [\citealp{stewart2012}])
 culminating recently with Spaun,
 currently the world's
-largest functional brain model \cite{TODO}.
+largest functional brain model \citep{eliasmith2012}.
 Spaun is a network of 2.5 million spiking neurons
@@
 involved adding or significantly modifying
-25 Java classes \cite{TODO}.
+25 Java classes \citep{bekolay2010}.
 
@@
 The JavaNengo scripting interface,
-described in \cite{TODO},
+described in \citet{stewart2009},
 provides a simpler interface
@@
 with other scientific Python tools
-for interactive exploration \cite{TODO}, %Ipython
-scientific computation \cite{TODO}, %scipy
-and plotting \cite{TODO}. % Matlab
+for interactive exploration \citep{perez2007},
+scientific computation \citep{oliphant2007},
+and plotting \citep{hunter2007}.
 Arbitrary CPython programs
@@
 through a scripting interface
-similar to JavaNengo's \cite{TODO}, % frontiers 2009
+similar to JavaNengo's \citep{stewart2009},
 with improvements
@@
 
-The Neural Engineering Framework (NEF; \cite{TODO})
+The Neural Engineering Framework (NEF; \citealp{eliasmith2003})
 proposes three principles
@@
 This idea is similar to population coding
-\cite{TODO}, but generalized
+\citep{georgopoulos1986}, but generalized
 to vectors of arbitrary length.
@@
 describes an ensemble representing a one-dimensional vector
-with 50 leaky integrate-and-fire neurons \cite{TODO} % LIF cite
+with 50 leaky integrate-and-fire neurons \citep{lapicque1907}
 that have a refractory period of 2 milliseconds.
@@
 has been done previously
-by PyNN \cite{TODO}.
+by PyNN \citep{davison2008}.
 In PyNN, the same Python script
@@
 \begin{center}
-  \includegraphics[width=.85\textwidth]{comm_channel}
-  \begin{minipage}{0.48\textwidth}
-    \begin{lstlisting}
+  \includegraphics[width=.73\textwidth]{comm_channel}
+  \begin{minipage}{0.37\textwidth}
+    \begin{lstlisting}[basicstyle={\footnotesize\ttfamily}]
 import nengo
@@
   \end{minipage}
-  \begin{minipage}{0.39\textwidth}
+  \begin{minipage}{0.3\textwidth}
     \includegraphics[width=\textwidth]{comm_channel_res}
@@
 \end{figure}
+
+\clearpage
 
@@
 Many models in theoretical neuroscience
-are based on attractor networks \cite{TODO}. % Amit, Rolls, Deco
+are based on attractor networks \citep{amit1992, deco2003}.
 The NEF has been used in the past
 to implement many different types of
-attractor networks \cite{TODO}.
+attractor networks \citep{eliasmith2005}.
 Figure~\ref{fig:lorenz} depicts
@@
 too long to be instructive.
+
+\clearpage
 
@@
 the same functionality
-with Brian, NEURON, NEST, and TODO PCSim.
+with Brian \citep{goodman2009}, NEURON \citep{hines2009},
+NEST \citep{eppler2008}, and PCSim \citep{pecevski2009}.
 We run each parameter set five times
@@
 The most closely related project is PyNN
-\cite{TODO},
+\citep{davison2008},
 which provides a high-level scripting
@@
 simulators with Python bindings,
-such as NEST \cite{TODO}.
+such as NEST.
 

### 2013-10-03 scripts/bench_cchannel.py
Improvements to benchmark figures

@@ -74,7 +74,7 @@ data['NEST'][(1000, 50)] = [
 
 cvs = []
 
-plt.figure(figsize=(4,4))
+plt.figure(figsize=(4,5))
 ax = plt.subplot(1,1,1)
 
 x = data.values()[0].keys()
@@ -100,6 +100,8 @@ plt.xticks(xval, xlabels)
 plt.xlim(70, 1130)
 plt.ylim(0, 200)
 plt.ylabel('Simulation time (s)')
+plt.text(720, 206.5, 'Communication channel benchmarks',
+         ha='center', va='center', fontsize='large')
 ax.spines['top'].set_visible(False)
 ax.spines['right'].set_visible(False)
 ax.xaxis.set_ticks_position('bottom')
@@ -107,7 +109,7 @@ ax.xaxis.set_tick_params(direction='out')
 ax.yaxis.set_ticks_position('left')
 
 inset = zoomed_inset_axes(ax, 7,
-                          bbox_to_anchor=(0.955, 0.99),
+                          bbox_to_anchor=(0.955, 0.935),
                           bbox_transform=plt.gcf().transFigure)
 
 xval = [n for (n, d) in x]
@@ -119,9 +121,9 @@ for ix, name in enumerate(data.keys()):
     inset.plot(xval, runtimes, '.-', ms=15, color=colors[ix], label=name)
     yoffset = 0
     if name == "PyNengo\nOpenCL":
-        yoffset = 1.2
+        yoffset = 0.85
     elif name == "PyNengo\nreference":
-        yoffset = -1.2
+        yoffset = -0.85
     if 'Nengo' in name:
         plt.text(xval[-1] + 7, runtimes[-1] + yoffset, name, color=colors[ix],
                  ha='left', va='center', fontsize=14)
@@ -137,7 +139,7 @@ inset.set_xticks(())
 
 mark_inset(ax, inset, loc1=3, loc2=4, fc="none", ec="0.5")
 
-plt.subplots_adjust(left=0.15, bottom=0.12, top=0.97, right=0.75)
+plt.subplots_adjust(left=0.15, bottom=0.1, top=0.935, right=0.75)
 plt.savefig("../figures/bench_cchannel.pdf")
 
 print "Max CV:", max(cvs)

### 2013-10-03 scripts/bench_cconv.py

@@ -71,7 +71,7 @@ data['Intel Xeon\nE5-2620'][100] = 1.68300008774
 data['Intel Xeon\nE5-2620'][200] = 3.46999979019
 data['Intel Xeon\nE5-2620'][500] = 12.3960001469
 
-plt.figure(figsize=(4,4))
+plt.figure(figsize=(4,5))
 ax = plt.subplot(1,1,1)
 
 for ix, name in enumerate(data.keys()):
@@ -79,15 +79,17 @@ for ix, name in enumerate(data.keys()):
     plt.plot(n, data[name].values(), '.-', ms=15, color=colors[ix])
     yoffset = 0
     if "3770" in name:
-        yoffset = 1.2
+        yoffset = 0.68
     elif "5540" in name:
-        yoffset = -0.6
+        yoffset = -0.7
     elif "Quadro" in name:
-        yoffset = 0.5
+        yoffset = 0.45
     elif "Tesla" in name:
-        yoffset = 0.3
+        yoffset = 0.2
     elif "Radeon" in name:
         yoffset = 0.2
+    elif "reference" in name:
+        yoffset = -0.68
 
     if name == "JavaNengo":
         plt.text(n[1] + 12000, data[name].values()[1], name, color=colors[ix],
@@ -102,6 +104,8 @@ xlabels = ["%d neurons\n%d dimensions" % (neurons[d], d) for d in dims]
 plt.xticks(xval, xlabels)
 plt.xlim(0, 500000)
 plt.ylim(0, 27.5)
+plt.text(280000, 28.5, 'Circular convolution benchmarks',
+         ha='center', va='center', fontsize='large')
 ax.tick_params(axis='x', which='major', labelsize=11)
 plt.ylabel('Simulation time (s)')
 ax.spines['top'].set_visible(False)
@@ -109,5 +113,5 @@ ax.spines['right'].set_visible(False)
 ax.xaxis.set_ticks_position('none')
 ax.yaxis.set_ticks_position('left')
 
-plt.subplots_adjust(left=0.12, bottom=0.11, top=0.97, right=0.82)
+plt.subplots_adjust(left=0.125, bottom=0.085, top=0.93, right=0.82)
 plt.savefig("../figures/bench_cconv.pdf")

### 2013-10-03 scripts/bench_lorenz.py

@@ -12,11 +12,11 @@ from mpl_toolkits.axes_grid1.inset_locator import mark_inset
 data = OrderedDict()
 data['Brian'] = OrderedDict()
 data['NEST'] = OrderedDict()
-# data['NEURON'] = OrderedDict()
+data['NEURON'] = OrderedDict()
 data['JavaNengo'] = OrderedDict()
 data['PyNengo\nreference'] = OrderedDict()
 data['PyNengo\nOpenCL'] = OrderedDict()
-colors = ['#238B45', '#FF7F00', 'k', '#E41A1C', '#08519C'] # '#984EA3',
+colors = ['#238B45', '#FF7F00', '#984EA3', 'k', '#E41A1C', '#08519C']
 
 data['PyNengo\nOpenCL'][100] = [
     0.823611974716,0.827681064606,0.824057102203,0.826804876328,0.825788021088]
@@ -54,14 +54,14 @@ data['Brian'][1000] = [
 data['Brian'][2000] = [
     1448.42626905,1461.23177814,1475.119735,1485.89479995,1462.42504406]
 
-# data['NEURON'][100] = [
-#     0.393055915833,0.394006967545,0.392350912094,0.398588895798,0.395780801773]
-# data['NEURON'][500] = [
-#     2.99810099602,2.97593998909,2.95858383179,2.93473315239,2.96808600426]
-# data['NEURON'][1000] = [
-#     9.42938303947,9.283867836,9.34721088409,9.22296404839,9.32386898994]
-# data['NEURON'][2000] = [
-#     20.4385299683,20.4057679176,20.3683900833,20.3296558857,20.2264928818]
+data['NEURON'][100] = [
+    0.393055915833,0.394006967545,0.392350912094,0.398588895798,0.395780801773]
+data['NEURON'][500] = [
+    2.99810099602,2.97593998909,2.95858383179,2.93473315239,2.96808600426]
+data['NEURON'][1000] = [
+    9.42938303947,9.283867836,9.34721088409,9.22296404839,9.32386898994]
+data['NEURON'][2000] = [
+    20.4385299683,20.4057679176,20.3683900833,20.3296558857,20.2264928818]
 
 data['NEST'][100] = [
     23.4582211971,22.8383898735,23.7844650745,24.908190012,24.6342151165]
@@ -74,7 +74,7 @@ data['NEST'][2000] = [
 
 cvs = []
 
-plt.figure(figsize=(4,4))
[…]

### 2013-10-04 nengo.tex
Edits from Eric's feedback, round 2

@@
   Nengo is a tool that builds and simulates
-  large-scale models with the NEF.
-  Nengo is the primary resource
+  large-scale models with the NEF;
+  it is the primary resource
   for teaching how the NEF
@@
 a population of neurons.
-For example, \texttt{Ensemble('A', nengo.LIF(50, tau\_ref=0.002), 1)}
+For example,
+\begin{quote}
+  \texttt{Ensemble('A', nengo.LIF(50, tau\_ref=0.002), 1)}
+\end{quote}
 describes an ensemble representing a one-dimensional vector
@@
 in several ways.
-Like other PyNengo objects,
-a connection contains symbolic information;
-in this case, a connection tracks
-how two objects are connected together.
+A \texttt{Connection} contains symbolic information
+about how two objects are connected.
+That information either includes
+a factored or full connection weight matrix,
+or includes enough information
+in order to generate weights during simulation.
 
@@
 using the \texttt{NonlinearityConnection}.
-Both of these connection types
+Both connection types
 can be temporally filtered,
@@
 
+\clearpage
+
 \paragraph{Signals}
@@
 between related low- and high-level signals.
-
-\clearpage
 
@@
 \begin{center}
-  \begin{minipage}{.435\textwidth}
-    \begin{lstlisting}[basicstyle={\footnotesize\ttfamily}]
-# We use Nengo to generate
-#     bias: Bias currents
-#     decoders: Decoding weights
-#     encoders: Encoding weights
-# for populations A and B.
-import pynn.nest as pyNN
-
-lif_params = {'tau_refrac': 2.0, 'tau_syn_E':100,
-               'tau_syn_I':100}
-dt = 0.5
-pyNN.setup(timestep=dt, min_delay=dt)
-A = pyNN.Population(30, pyNN.IF_cond_exp, lif_params)
-B = pyNN.Population(30, pyNN.IF_cond_exp, lif_params)
-input_signal = 0.5
-inputA = [pyNN.DCSource(amplitude=val) for val
-           in A_bias + input_signal * A_encoders]
-for i, pulse in enumerate(inputA):
-    pulse.inject_into(A[i:i+1])
-inputB = [pyNN.DCSource(amplitude=val) for val
-           in B_bias]
-for i, pulse in enumerate(inputB):
-    pulse.inject_into(B[i:i+1])
-    \end{lstlisting}
-  \end{minipage}
-  \begin{minipage}{.46\textwidth}
-    \begin{lstlisting}[basicstyle={\footnotesize\ttfamily}]
-weights = []
-for i in xrange(30):
-    for j in xrange(30):
-        weights.append((i, j,
-            np.dot(A_decoder[i], B_encoder[j]), 1.0))
-connection = pyNN.Projection(
-    A, B, pyNN.FromListConnector(weights))
-A.record('spikes')
-B.record('spikes')
-pyNN.run(1000)
-
-# Decoded value of B
-B_output = numpy.zeros(1000)
-for i in xrange(1000):
-    spikes = B[i:i+1].getSpikes()[:,1]
-    spikes = (spikes / dt).astype('int')
-    B_output[spikes] += B_decoder[i]
-# Filter with tau = 200 ms
-decay = np.exp(-dt / 200)
-B_output[0, :] *= (1 - decay)
-for i in xrange(1, 1000):
-    B_output[i,:] = decay * B_output[i-1,:]
-                     + (1-decay) * B_output[i,:]
-    \end{lstlisting}
-  \end{minipage}
+  \includegraphics[width=\textwidth]{pynn}
 \end{center}
@@
 \end{figure}
-
-\clearpage
 
@@
 too long to be instructive.
-
-\clearpage
 
@@
 While NEST and NEURON were not
-run with MPI,
+run on multiple cores using message passing,
 the reference simulator of PyNengo
@@
 PyNengo's OpenCL simulator is significantly faster.
+
+\clearpage
 
@@
 
-\section*{Acknowledgement}
-
-We would like to thank
-everyone who has contributed
-to this reimplementation of Nengo
-by providing examples,
-unit tests, and bugfixes:
-Peter Blouw, Brent Komer, and Bryan Tripp.
-
-% Also acknowledge Bryan for making JavaNengo, and edits
+\section*{Acknowledgments}
+
+We thank Bryan Tripp
+for editing this paper,
+contributing to PyNengo
+and the PyNengo OpenCL simulator
+and for creating the JavaNengo simulator
+that has been useful
+for the past six years,
+and will continue
+to be used for many years to come.
+We thank Peter Blouw and Brent Komer,
+who have contributed
+to PyNengo by providing examples,
+unit tests, and bugfixes.
 

### 2013-10-04 scripts/bench_lorenz.py

@@ -89,7 +89,7 @@ for ix, name in enumerate(data.keys()):
     if name == "Brian":
         plt.text(xval[2] + 60, runtimes[2], name, color=colors[ix],
                  ha='left', va='center', fontsize=14)
-    elif name in ("NEURON", "NEST"):
+    elif name == "NEST":
         plt.text(xval[-1] + 60, runtimes[-1], name, color=colors[ix],
                  ha='left', va='center', fontsize=14)
 

### 2013-10-04 scripts/nef_summary.py

@@ -262,8 +262,8 @@ plt.gca().yaxis.set_ticks_position('left')
 
 plt.subplot(2,1,2)
 plt.plot(sim.data('Oscillator')[:,0], sim.data('Oscillator')[:,1])
-plt.xlabel('$x_0$')
-plt.ylabel('$x_1$')
+plt.xlabel('$x_1$')
+plt.ylabel('$x_2$')
 plt.axis([-1.2, 1.2, -1.2, 1.2])
 plt.grid()
 plt.axvline(color='k')

### 2013-10-04 nengo.tex
Edits from Chris's feedback

@@
   a comprehensive theory of
-  how the brain implements cognitive processes
-  with biological parts.
+  how cognitive processes can be
+  implemented in a biological substrate.
   The Neural Engineering Framework (NEF)
@@
   large-scale models with the NEF.
-  Nengo is a tool that builds and simulates
-  large-scale models with the NEF;
-  it is the primary resource
-  for teaching how the NEF
-  is used to model neural systems,
+  Nengo is a software tool
+  that can be used to build and simulate
+  large-scale models based on the NEF;
+  currently, it is the primary resource
+  for both teaching how the NEF
+  is used,
   and for doing research
-  comparing NEF models
-  to experimental data.
+  that generates specific
+  NEF models
+  to explain experimental data.
   A previous version of Nengo
@@
   to create Spaun,
-  the world's largest functional brain model.
+  the world's largest functional
+  brain model \citep{eliasmith2012}.
   Simulating Spaun
-  identified limitations
+  highlighted limitations
   in Nengo's ability to
-  construct models with simple syntax,
-  simulate them quickly,
-  and collect data for subsequent analysis.
+  support model construction with simple syntax,
+  to simulate large models quickly,
+  and to collect large amounts of data
+  for subsequent analysis.
   This paper describes a new version of Nengo
@@
   simulates a benchmark model on the scale of Spaun
-  200 times faster than the previous implementation,
+  50 times faster than the previous implementation,
   and has a flexible mechanism
@@
   \tiny \keyFont{\section{Keywords:} Python, Neural Engineering
-    Framework, theoretical neuroscience, computational neuroscience,
+    Framework, Nengo, theoretical neuroscience, computational neuroscience,
     control theory, software, simulation}
@@
 Computational neuroscience
-has made significant advancements;
+has made significant advancements
+from simulating low-level biological parts in great detail,
+to solving high-level problems that humans find difficult;
 however, we still lack a mathematical account of
-how biological and electrical components
+how biological components
 implement cognitive functions such as sensory processing,
@@
 to the statistics of biological synapses \citep{markram2006}.
-Cognition has not yet emerged
-from data-driven large scale models.
-On the other end of the spectrum,
+However, cognition has not yet emerged
+from data-driven large scale models,
+and there are good reasons to think
+that cognition may never emerge \citep{trujillo-inpress}.
+At the other end of the spectrum,
 cognitive architectures \citep{anderson2004}
-and theoretical frameworks \citep{hawkins2006}
+and machine learning approaches \citep{hinton2006}
 have solved high-dimensional statistical problems,
-but do so without respecting biological constraints,
-and with worse performance than machine learning approaches.
+but do so without respecting biological constraints.
 
@@
 The NEF is a large-scale modeling approach
-that can leverage any neuron model
+that can leverage single neuron models
 to build neural networks with
@@
 Spaun is a network of 2.5 million spiking neurons
-that can perform eight high-level cognitive tasks
+that can perform eight cognitive tasks
 including memorizing lists and inductive reasoning.
@@
 the appropriate series of images
-representing the task to be performed
-and the details of that task;
+representing the task to be performed;
 for example, sequentially presenting images
@@
 If asked to recall the memorized list,
-it sends motor commands to a simulated arm model,
-tracing out the digits $1234$.
-While the tasks Spaun performs are diverse,
-all tasks use a common set of
-functional modules.
-Each functional module corresponds
+Spaun generates motor commands for a simulated arm,
+writing out the digits $1234$.
+While the tasks that Spaun performs are diverse,
+all of the tasks use a common set of
+functional cortical and subcortical components.
+Each functional component corresponds
 to a brain area that has been hypothesized
 to perform those functions
-in neuroscientific literature.
-
-The NEF shows how to
-construct a neural model
-from anatomical constraints,
-cognitive objectives, and
-dynamical systems theory.
-Perhaps surprisingly,
-constructing models from this starting point,
+in the neuroscientific literature.
+
+The NEF provides principles
+to guide the construction of
+a neural model
+that incorporates anatomical constraints,
+functional objectives, and
+dynamical systems or control theory.
+Constructing models from this starting point,
 rather than from single cell electrophysiology
-and connectivity statistics,
-still produces simulated data that explains
-and predicts experimental results.
-Single cell activity,
-response timing, and behavioral errors
+and connectivity statistics alone,
+produces simulated data that explains
+and predicts a wide variety of experimental results.
+Single cell activity \citep{stewart2012},
+response timing \citep{stewart2009a},
+and behavioral errors \citep{choo2010}
 of NEF-designed models match
@@
 without being built specifically into the design.
-These quirks seem to fall directly out
-of the need to satisfy competing objectives
+These results are a consequence of
+the need to satisfy functional objectives
 within anatomical and neurobiological constraints.
-The opposite phenomenon, in which
-cognitively useful behaviors emerge
-from networks of detailed neurons
-connected with plausible connectivity statistics,
-has not yet been observed.
 
 It has often been said that neuroscience
-is data-rich and theory-poor.
-The modeling success of
+is data-rich and theory-poor \citep{churchland1992}.
+We believe that the modeling success of
 large functional networks such as Spaun
-indicates that the NEF
-is beginning to fill the theory void.
+suggests that the NEF
+may be able to fill the theoretical void.
 However, in simulating this large model,
-we pushed our Java-powered simulator
-to its limits in terms of memory consumption,
+the Java-powered simulator
+was limited in terms of memory consumption,
 simulation speed, and extensibility.
@@
 we have begun the next generation
-of Nengo with a focus on speed,
+of Nengo development with a focus on speed,
 extensibility, and simplicity.
@@
 The JavaNengo GUI provides
-a drag-and-drop interface in which to create NEF models.
+a drag-and-drop interface in which
+NEF models can be created.
 However, as it was developed
@@
 in the Java codebase would require
-so many extensive changes
+enough extensive changes
 that we instead opted
@@
 
-PyNengo does not currently
+PyNengo does not yet
 replace all of JavaNengo's functionality.
@@
 models that JavaNengo can simulate,
-there is no graphical interface to PyNengo.
+there is currently no graphical interface for PyNengo.
 Model components must be created
@@
 like JavaNengo's GUI on top of PyNengo's object model,
-but that remains future work.
+but that remains as future work.
 
@@
 using a well-tested codebase of
-TODO lines of code,
-compared to JavaNengo's TODO
+$\sim$5000 lines of code,
+compared to JavaNengo's $\sim$45,000
 (which does not include the JavaNengo GUI).
@@
 and can therefore further test the NEF
-as a theory of the brain.
+as a theory of neural computation.
 
@@
 The Neural Engineering Framework (NEF; \citealp{eliasmith2003})
-proposes three principles
+proposes three quantitatively specified principles
 that enable the construction
 of large-scale neural models.
+Briefly, this mathematical theory defines
 
@@
   \item \textbf{Representation:} A population of neurons
-    collectively represents a time-varying vector of real numbers.
-  \item \textbf{Transformation:} Functions on those vectors
-    are computed by the connections between populations of neurons.
+    collectively represents a time-varying vector of real numbers
+    through nonlinear encoding and linear decoding.
+  \item \textbf{Transformation:} Linear and nonlinear
+    functions on those vectors
+    are computed by linear decodings
+    that are used to analytically compute
+    the connections between populations of neurons.
   \item \textbf{Dynamics:} The vectors represented
     by neural populations can be considered state variables
-    in a dynamical system, and can be analyzed using control theory.
+    in a (linear or nonlinear) dynamical system,
+    and recurrent connections can be computed
+    using principle 2.
 \end{enumerate}
@@
 Information is encoded by populations of neurons.
-We represent information
-with time-varying vectors of real numbers
-in order to manipulate that information
+The NEF represents information
+with time-varying vectors of real numbers,
+allowing theorists to propose possible
+neural computations by
+manipulating that information
 using conventional mathematics.
-By doing this, we can \textit{encode}
+The NEF suggests that we can characterize
+the \textit{encoding} of
 those vectors by injecting
@@
 With enough neurons,
-we can estimate the originally encoded vector
+the originally encoded vector
+can be estimated
 through a \textit{decoding} process.
-This idea is similar to population coding
-\citep{georgopoulos1986}, but generalized
-to vectors of arbitrary length.
+This idea is a kind of population coding
+\citep{georgopoulos1986, salinas1994}, but generalized
+to vectors of arbitrary dimensionality.
 
@@
 the spike trains are first filtered
-with an exponentially decaying filter.
+with an exponentially decaying filter
+accounting for the process
+of a spike generating a postsynaptic current.
 Those filtered spike trains are summed together
@@
 this minimization on points
-randomly sampled from
+sampled from
 the vector space
@@
 which typically causes some amount of current
-to be imparted in the downstream neuron.
+to be imparted in the postsynaptic (downstream) neuron.
 Many factors affect the
@@
 the decoding weights for that function
-in the first population
-and the encoding weights
-for the downstream population.
-
-This implies the NEF hypothesizes
+in the first population,
+the encoding weights
+for the downstream population,
+and any linear transform.
+
+This implies that the NEF hypothesizes
 that synaptic weight matrices
-have low rank, and can be factored.
+have low rank, and can be factored
+into encoders, decoders, and a linear transform.
 Note that, in practice, we rarely use
@@
 Large models can be built
-by using of the principles of the NEF
+by using the principles of the NEF
 as connectable components
@@
 
-To describe the high-level functional model,
+To describe an NEF model,
 PyNengo defines six core objects.
@@
 
+\clearpage
+
 The reference simulator
@@
 and operators used in a simple model.
-
-\clearpage
 

### 2013-10-04 nengo.tex
Edits from Chris's feedback, round 2

@@
   \item The \texttt{Node} represents non-neural information,
-    such as sensory inputs.
+    such as sensory inputs and motor outputs.
   \item The \texttt{Connection} describes how
     nodes and ensembles are connected.
-  \item The \texttt{Network} is a group of interconnected
-    nodes and ensembles.
+  \item The \texttt{Network} encapsulates a functionally related
+    group of interconnected nodes and ensembles.
   \item The \texttt{Probe} gathers data during a simulation
@@
 \end{enumerate}
+These six objects contain symbolic information
+about a Nengo model,
+enabling a strict division between
+model construction and simulation.
+This allows a Nengo model
+to be run on multiple simulators.
 
@@
 When creating an ensemble,
-the user must provide the dimensionality
-(i.e., the length of the vector it represents),
-and an object that describes
-a population of neurons.
+the user must provide a name,
+an object that describes
+a population of neurons,
+and the dimensionality
+(i.e., the length of the vector it represents).
 For example,
@@
 \end{quote}
-describes an ensemble representing a one-dimensional vector
-with 50 leaky integrate-and-fire neurons \citep{lapicque1907}
-that have a refractory period of 2 milliseconds.
+describes an ensemble, `A',
+that uses 50 leaky integrate-and-fire neurons \citep{lapicque1907}
+with a 2 millisecond refractory period
+to represent a one-dimensional vector.
 The \texttt{nengo.LIF} class defines
@@
 and responds to the neural model's output.
-Nodes allow many of the special-purpose
-routines that are necessary in real-world
-models to be integrated
-as core components of a Nengo model.
+Nodes allow Nengo to represent
+neural components,
+the body that those components drive,
+and the environment that body interacts with
+in a single unified model.
 This makes Nengo models more explicit,
@@
 and can be composed of other networks.
-The \texttt{Integrator} network, for example,
+As a simple example, the \texttt{Integrator} network
 is composed of only one recurrently connected ensemble.
@@
 the purpose of that ensemble is made explicit.
-On the other end of the spectrum,
+As a complex example,
 the \texttt{BasalGanglia} network
@@
 
-A \texttt{Probe} denotes that
+A \texttt{Probe} monitors
 a particular part of another object
-will be recorded throughout a simulation.
+that will be recorded throughout a simulation.
 Nengo models contain many variables
@@
 in files or to stream data
-directly to sockets will be added.
+directly to sockets is forthcoming.
 
@@
    (E) The signals tracked at the high level
-   can be interpreted as a dynamical system
-   that can be analyzed with control theory.
+   can be interpreted as a dynamical system.
    State variable A simply represents its input,
@@
    directly, without the use of spiking neurons,
-   in order to quickly analyze system behavior.}
+   in order to quickly analyze system behavior,
+   if desired.}
 \end{figure}
@@
 in PyNN for benchmarking purposes
-(code not shown).
+(code not shown; the PyNN script is 100 lines long,
+while the Nengo script [see Figure~\ref{fig:lorenz}
+is 20 lines long).
 
@@
    Circular Convolution network through
-   complicated transformation matrices.
+   complicated transformation matrices
+   that implement a discrete Fourier transform.
    Each ensemble within the network represents a
@@
    is projected through another complicated transformation matrix
-   to compute the final four-dimensional result.
+   that implements the inverse discrete Fourier transform,
+   computing the final four-dimensional result.
    Note that the complicated parts of the model
@@
 is circular convolution.
-Circular convolution is how
-Spaun binds two concepts together.
-Binding is central to solving many
-of the high level cognitive tasks
-that make Spaun unique.
+This function is a central part
+of the Semantic Pointer Architecture \citep{eliasmith2013}
+that uses the NEF to build
+a variety of cognitive models including Spaun.
 Circular convolution
@@
 rather than in a single connection.
-This two-layer organization is simplified
+This two-layer organization is simplified in
 the \texttt{CircularConvolution} network subclass.
@@
      except that NEURON is notably faster.
-     However, when analyzed, the NEURON model
-     emits almost no spikes after the first
-     simulated second, which may be
-     why the simulation is relatively fast.
      The full results for Brian and NEST are not shown;
@@
 NEST \citep{eppler2008}, and PCSim \citep{pecevski2009}.
-We run each parameter set five times
+We ran each parameter set five times
 on the same machine,
@@
 outliers with coefficients of 0.18 and 0.22,
-indicating that these means are robust.
+overall indicating that the reported means are robust.
 The results, shown in Figure~\ref{fig:benchmarks}A and B,
@@
 through PyNN, especially
-as the size of models increase.
+as the size of models increases.
 This is likely due to Nengo's
@@
 
-As a real-world example,
+As a larger-scale example,
 we have also benchmarked
@@
 performs much faster than JavaNengo;
-in particular, a Radeon 7970 GPU performs 500-dimensional
-circular convolution faster than real time,
-and 200 times faster than JavaNengo.
+in particular, a Radeon 7970 GPU performs
+500-dimensional circular convolution
+with about half a million neurons
+faster than real time,
+and 50 times faster than JavaNengo.
 Additionally, although both JavaNengo
@@
 that has been validated
-through the Spaun model.
+through the Spaun model and other past work.
 
@@
 accessible simulators.
-PyNN is therefore better suited to
+Because Nengo is in an earlier development stage,
+many of these detailed neuron models
+and learning rules remain to be added.
+PyNN is therefore currently better suited to
 simulate detailed neural models.
@@
 and simulators to Nengo.
-Internally, we will
+Internally, we will be
 implementing the JavaNengo use cases
@@
 the OpenCL simulator is extremely fast;
-it can simulate a large circular convolution model
-200 times faster than JavaNengo.
+it can simulate circular convolution models
+50-200 times faster than JavaNengo,
+which is faster than alternative simulators
+on simpler models.
 This makes the creation and simulation
@@
 Terrence C. Stewart contributed
-to PyNengo and helped implement
+to PyNengo, helped implement
 the PyNN scripts used in
-sections~\ref{sec:examples} and \ref{sec:benchmark}.
+sections~\ref{sec:examples} and \ref{sec:benchmark},
+and implemented the JavaNengo scripting interface
+on which PyNengo is based.
 Daniel Rasmussen contributed
@@
 \paragraph{Funding\textcolon}
-NSERC CRC, CFI, and OIT.
+NSERC Discovery, NSERC Graduate Fellowships,
+NSERC Banting Fellowship, ONR (N000141310419)
+and AFOSR (FA8655-13-1-3084).
 

### 2013-10-06 nengo.tex
Updated to conform to Frontiers guidelines
- This is the submitted version!

@@
 \usepackage{inconsolata}
-
-\definecolor{dkgreen}{rgb}{0,0.6,0}
-\definecolor{gray}{rgb}{0.5,0.5,0.5}
-\definecolor{mauve}{rgb}{0.58,0,0.82}
-
-\lstset{
-  language=Python,
-  aboveskip=3mm,
-  belowskip=3mm,
-  showstringspaces=false,
-  columns=flexible,
-  basicstyle={\small\ttfamily},
-  numbers=none,
-  numberstyle=\tiny\color{gray},
-  keywordstyle=\color{blue},
-  commentstyle=\color{dkgreen},
-  stringstyle=\color{mauve},
-  breaklines=true,
-  breakatwhitespace=true,
-  deletendkeywords={filter},
-}
-
 \graphicspath{{figures/}}
@@
 \usepackage{url,lineno}
-% \linenumbers  % TODO uncomment!
+\linenumbers
 
@@
   and Chris Eliasmith\,$^{1}$}
-\def\Address{$^{1}$Computational Neuroscience Research Group,
-  Center for Theoretical Neuroscience,
+\def\Address{$^{1}$Center for Theoretical Neuroscience,
   University of Waterloo, Waterloo, ON, Canada}
@@
 \extraAuth{}
-%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2,
-%  Organization X2, Street X2, City X2 , State XX2
-%  (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}
 \topic{Python in Neuroscience II}
@@
 
-  \tiny \keyFont{\section{Keywords:} Python, Neural Engineering
-    Framework, Nengo, theoretical neuroscience, computational neuroscience,
-    control theory, software, simulation}
+  \tiny \keyFont{\section{Keywords:} Neural Engineering
+    Framework, Nengo, Python, neuroscience, theoretical neuroscience,
+    control theory, simulation}
 \end{abstract}
-
-%max figures+tables: 15
-%max legnth: 12000 words
-%max pdf length: 12 pages
 
@@
 However, in simulating this large model,
-the Java-powered simulator
+the Java-powered Nengo simulator
 was limited in terms of memory consumption,
@@
 In order to continue to evaluate the NEF
-as a theory of the brain
+as a theory of neural computation
 by building larger and more complicated models,
@@
 enough extensive changes
-that we instead opted
+that we opted instead
 to write a new version of Nengo in Python.
@@
 to other simulation packages
-(e.g., those accessible by PyNN)
+(e.g., those accessible by PyNN [\citealp{davison2008}])
 or for faster simulators to be used
@@
 of large-scale neural models.
-Briefly, this mathematical theory defines
-
+Briefly, this mathematical theory defines:
 \begin{enumerate}
@@
 \end{enumerate}
-
-\begin{figure}
-\begin{center}
-  \includegraphics[width=0.245\textwidth]{nef_summary_enc}
-  \includegraphics[width=0.245\textwidth]{nef_summary_dec}
-  \includegraphics[width=0.245\textwidth]{nef_summary_trans}
-  \includegraphics[width=0.245\textwidth]{nef_summary_dyn}
-\end{center}
- \textbf{\refstepcounter{figure}\label{fig:nef} Figure \arabic{figure}.}{
-   Summary of the three principles of the Neural Engineering Framework
-   (NEF). (A) By the representation principle, signals are encoded
-   in neural populations based on the \textit{tuning curve}
-   of each neuron (top). The tuning curve describes
-   how active a neuron is given some input signal.
-   If we drive the eight neurons in the top panel
-   with the input signal in the middle panel,
-   we see the spike trains in the bottom panel.
-   (B) By the representation principle,
-   the spiking activity of a neural population
-   can be decoded to recover the original input signal,
-   or some transformation of that input signal.
-   First, the firing pattern shown in the top panel
-   is filtered with a decaying exponential filter (middle panel).
-   The filtered activity is then summed together
-   with a set of weights that approximates
-   the input signal (bottom panel, green)
-   and the cosine of the input signal (bottom panel, purple).
-   (C) A sine wave is encoded by population A (top panel);
-   the negative of that signal is projected
-   to population B (middle panel)
-   and the square of that signal is projected
-   to population C (bottom panel).
-   By the transformation principle,
-   populations of neurons can send signals
-   to another population by decoding
-   the desired function from the first population
-   and then encoding the decoded estimate
-   into the second population.
-   These two steps can be combined into a single step
-   by calculating a set of weights
-   that describe the strength of the connections
-   between each neuron in the first population
-   and each neuron in the second population.
-   (D) A neurally implemented dynamical system
-   has negative feedback across its two dimensions,
-   resulting in a harmonic oscillator.
-   The oscillator is plotted across time (top)
-   and in state space (bottom).
-   By the dynamcs principle,
-   signals being represented by population of neurons
-   can be thought of as state variables in a dynamical system.
-   }
-\end{figure}
+Figure~\ref{fig:nef} provides a graphical
+summary of these three principles.
 
@@
 
-Figure~\ref{fig:nef} depicts
-the encoding (A) and decoding (B) processes.
 In the encoding process, the input signal drives
@@
 that causes the neuron to be the most active).
-% This results in a predictable pattern of firing
-% given the input signal.
 
@@
 
-This implies that the NEF hypothesizes
-that synaptic weight matrices
+This implies that the NEF makes
+a hypothesis about synaptic weight matrices;
+specifically, that they
 have low rank, and can be factored
@@
 and instead store
-the encoding and decoding weights separately
-(i.e., the two factors of the connection weight matrix).
+the encoders, decoders, and linear transform separately
+(i.e., the three factors of the connection weight matrix).
 This provides significant
@@
 the maximum firing rate of its neurons,
-and so on, can also be specified
+and so on, can be specified
 either as keyword arguments
@@
 or includes enough information
-in order to generate weights during simulation.
+to generate weights during simulation.
 
@@
 A network is a collection of interconnected ensembles and nodes.
-It is primarily a way of grouping
+It is primarily a way of grouping together
 a set of connected objects
-that work together to perform a complex function.
+that collectively perform a complex function.
 Encapsulating them in a network
@@
 a particular part of another object
-that will be recorded throughout a simulation.
+in order to record its value throughout a simulation.
 Nengo models contain many variables
@@
 
-\begin{figure}
-\begin{center}
-  \includegraphics[width=\textwidth]{sim}
-\end{center}
- \textbf{\refstepcounter{figure}\label{fig:sim} Figure \arabic{figure}.}{
-   Detailed breakdown of the PyNengo reference simulator
-   running a simple model for two timesteps.
-   (A) Code describing the model being simulated.
-   It consists of ensemble A projecting the sine of its
-   encoded signal to ensemble B,
-   which is recurrently connected.
-   (B) Diagram depicting the model being simulated.
-   (C) A detailed diagram of how the reference simulator
-   organizes this model. Signals (blue) represent the values
-   tracked in the simulation.
-   Operators (red) represent the computations done on signals.
-   Signals can be grouped as low-level neural signals
-   that are used to compute the nonlinear functions
-   underlying neuron models,
-   and high-level NEF signals that are used to
-   drive neurons and track the signals
-   that the neurons are representing.
-   The operators that implement the decoding
-   and encoding steps map between
-   the low-level neural signals
-   and the high-level NEF signals.
-   (D) The signals tracked at the low level
-   can be interpreted as a model
-   commonly seen in computational neuroscience literature;
-   a population of leaky integrate-and-fire neurons
-   is driven by some time-varying input current, $J(t)$.
-   These neurons project to a population
-   of recurrently connected neurons.
-   The connection weights between the two populations,
-   and from the second population to itself,
-   can be computed by the NEF's transformation
-   principle, bypassing the need for
-   the high-level NEF signals
-   used by the reference simulator
-   for speed and data collection purposes.
-   (E) The signals tracked at the high level
-   can be interpreted as a dynamical system.
-   State variable A simply represents its input,
-   and passes its state to a sine function
-   which becomes the input to B.
-   State variable B is a simple linear system
-   that can be described with the typical
-   $\dot{x}(t) = A x(t) + B u(t)$ equation.
-   These dynamical systems can be simulated
-   directly, without the use of spiking neurons,
-   in order to quickly analyze system behavior,
-   if desired.}
-\end{figure}
-
 \section{PyNengo simulators} \label{sec:simulators}
@@
 Nengo follows this programming model by
-decoupling model creation and simulation,
+decoupling neural model creation and simulation,
 which enables PyNengo simulators
@@
 
-\clearpage
-
 The reference simulator
@@
 
-\paragraph{Reference simulator description}
+\paragraph{Reference simulator}
 
@@
 
-\begin{figure}
-\begin{center}
-  \includegraphics[width=.73\textwidth]{comm_channel}
-  \begin{minipage}{0.37\textwidth}
-    \begin{lstlisting}[basicstyle={\footnotesize\ttfamily}]
-import nengo
-from nengo.helpers import white_noise
-model = nengo.Model("Communication Channel")
-model.make_node("Input", white_noise(1, 5))
-model.make_ensemble("A", nengo.LIF(30), 1)
-model.make_ensemble("B", nengo.LIF(30), 1)
-model.connect("Input", "A")
-model.connect("A", "B")
-model.probe("Input")
-model.probe("A", filter=0.01)
-model.probe("A.spikes")
-model.probe("B", filter=0.01)
-model.probe("B.spikes")
-sim = model.simulator()
-sim.run(1)
-A_data = sim.data("A")
-A_spikes = sim.data("A.spikes")
-...
-    \end{lstlisting}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-    \includegraphics[width=\textwidth]{comm_channel_res}
-  \end{minipage}
-\end{center}
- \textbf{\refstepcounter{figure}\label{fig:comm-channel}
-   Figure \arabic{figure}.}{
-   A communication channel implemented with Nengo.
-   (A) Diagram depicting the model. Ensemble A
-   projects its encoded signal to ensemble B unchanged.
-   (B) Nengo code to build and simulate the model
-   for 1 second.
-   (C) The results of the simulation.
-   The input signal (top panel) is white noise limited to 0 to 5 Hz.
-   The signal is well represented by both ensemble A (middle panel)
-   and ensemble B (bottom panel) despite the neural firing patterns
-   (underlaid in middle and bottom panels) being different.}
-\end{figure}
-
 A communication channel
@@
 that it can be readily implemented in PyNN.
-Figure~\ref{fig:pynn} contains the code
-for implementing a communication channel in PyNN.
+Figure~\ref{fig:pynn} compares the code
+for implementing a communication channel in Nengo and PyNN.
 This highlights many of the differences
 between Nengo models and conventional neural models;
-we also use the script for benchmarking
+we also use these script for benchmarking
 (see section \ref{sec:benchmark}).
 
-\begin{figure}
-\begin{center}
-  \includegraphics[width=\textwidth]{pynn}
-\end{center}
- \textbf{\refstepcounter{figure}\label{fig:pynn} Figure \arabic{figure}.}{
-   Implementation of the communication channel in PyNN.
-   Note that several quantities are generated from Nengo;
-   the functions that generate these quantities in Nengo
-   can be decoupled, but are not included here for brevity.}
-\end{figure}
-
 \subsection{Lorenz attractor network} \label{sec:lorenz}
-
-\begin{figure}
-\begin{center}
-  \begin{minipage}{0.43\textwidth}
-    \begin{lstlisting}[basicstyle={\footnotesize\ttfamily}]
-import nengo
-tau = 0.1; sigma = 10; beta = 8.0/3; rho = 28
-model = nengo.Model('Lorenz attractor')
-model.make_ensemble('State', nengo.LIF(2000),
-                      dimensions=3, radius=60)
-
-def feedback(x):
-    dx0 = -sigma * x[0] + sigma * x[1]
-    dx1 = -x[0] * x[2] - x[1]
-    dx2 = x[0] * x[1] - beta * (x[2] + rho) - rho
-
-    return [dx0 * tau + x[0],
-            dx1 * tau + x[1],
-            dx2 * tau + x[2]]
-
-model.connect('State', 'State',
-                function=feedback, filter=tau)
-model.probe('State', filter=tau)
-model.probe('State.spikes')
-sim = model.simulator()
-sim.run(6)
-    \end{lstlisting}
-  \end{minipage}
-  \begin{minipage}{0.55\textwidth}
-    \includegraphics[width=0.95\textwidth]{lorenz}
-    \includegraphics[width=\textwidth]{lorenz_res}
-  \end{minipage}
-\end{center}
- \textbf{\refstepcounter{figure}\label{fig:lorenz} Figure \arabic{figure}.}{
-   A Lorenz attractor implemented with Nengo.
-   (A) Nengo code to build and simulate the model
-   for 6 seconds.
-   (B) Diagram depicting the model. The state ensemble
-   is recurrently connected with a complex function
-   implementing the dynamics of the Lorenz attractor.
-   Note that this populations does not receive
-   any input that might drive its initial value;
-   instead, the initial value is determined by
-   the baseline firing of the 2000 leaky integrate-and-fire
-   neurons that make up the state ensemble.
-   (C) The trajectory that the state ensemble takes
-   in its three-dimensional state space.
-   For the parameters chosen, the trajectory takes
-   the well-known butterfly shape.
-   (D) The state vector plotted over time.
-   (E) The spikes emitted by a random sample of 25
-   neurons from the state ensemble.
-   Some neurons fire uniformly across the 6 second simulation,
-   but most change depending on the state being tracked
-   due to the recurrent connection.}
-\end{figure}
 
@@
 in PyNN for benchmarking purposes
-(code not shown; the PyNN script is 100 lines long,
-while the Nengo script [see Figure~\ref{fig:lorenz}
+(code not shown; the PyNN script is $\sim$100 lines long,
+while the Nengo script in Figure~\ref{fig:lorenz}
 is 20 lines long).
@@
 \subsection{Circular convolution} \label{sec:cconv}
-
-\begin{figure}
-\begin{center}
-  \includegraphics[width=0.8\textwidth]{cconv}
-  \begin{minipage}{0.46\textwidth}
-    \begin{lstlisting}[basicstyle={\footnotesize\ttfamily}]
-import nengo
-from nengo.networks import CircularConvolution
-model = nengo.Model("Circular convolution")
-model.make_node("Input A", [-0.21, 0.5, 0.12, 0.06])
-model.make_ensemble("A", nengo.LIF(512), 4)
-model.connect("Input A", "A")
-model.make_node("Input B", [-0.18, 0.28, 0.18, -0.52])
-model.make_ensemble("B", nengo.LIF(512), 4)
-model.connect("Input B", "B")
-cconv = model.add(CircularConvolution("Circular Convolution",
-    neurons=nengo.LIF(1032), dimensions=4))
-model.connect("A", cconv.A)
-model.connect("B", cconv.B)
-model.make_ensemble("Result", nengo.LIF(512), 4)
-model.connect(cconv, "Result")
-model.probe("A", filter=0.02)
-model.probe("B", filter=0.02)
-model.probe("Result", filter=0.02)
-sim = model.simulator()
-sim.run(0.5)
-    \end{lstlisting}
-  \end{minipage}
-  \begin{minipage}{0.46\textwidth}
-    \includegraphics[width=\textwidth]{cconv_res}
-  \end{minipage}
-\end{center}
- \textbf{\refstepcounter{figure}\label{fig:cconv} Figure
-   \arabic{figure}.}{
-   Circular convolution implemented with Nengo.
-   (A) Diagram depicting the model.
-   The input vectors, A and B, represent four-dimensional vectors
-   which are mapped onto six ensembles within the
-   Circular Convolution network through
-   complicated transformation matrices
-   that implement a discrete Fourier transform.
-   Each ensemble within the network represents a
-   two-dimensional vector. The product of the two dimensions
-   is projected through another complicated transformation matrix
-   that implements the inverse discrete Fourier transform,
-   computing the final four-dimensional result.
-   Note that the complicated parts of the model
-   are contained within the network;
-   the number of ensembles and the transform matrices shown
-   are automatically generated by the network depending on
-   the dimensionality of the input vectors.
-   (B) Nengo code to build and simulate the model
-   for 0.5 seconds.
-   (C) The result of the simulation.
-   Straight horizontal lines represent
-   the target values that each ensemble
-   should represent.
-   Wavy lines represent the decoded values
-   for each dimension represented by the
-   A, B, and Result ensembles
-   (top, middle and bottom panels, respectively).
-   The ensembles represent the correct values,
-   after a startup transient of less than 0.1 seconds.}
-\end{figure}
 
@@
 
-\begin{figure}
-\begin{center}
-  \includegraphics[width=0.32\textwidth]{bench_cchannel}
-  \includegraphics[width=0.32\textwidth]{bench_lorenz}
-  \includegraphics[width=0.32\textwidth]{bench_cconv}
-\end{center}
- \textbf{
-     \refstepcounter{figure}\label{fig:benchmarks}
-     Figure \arabic{figure}.}{
-     Benchmark results for several simulators
-     on the example models described
-     in section~\ref{sec:examples}.
-     In (A) and (B), all of the simulators
-     except the PyNengo OpenCL simulator
-     were run on an Intel Core i7-965.
-     JavaNengo used all 4 cores of this processor;
-     all other simulator used only 1 core.
-     The PyNengo OpenCL simulator
-     was run on an NVidia GTX280 GPU.
-     (A) Benchmark results from simulating
-     the communication channel for 10 simulated seconds
-     at a 1 millisecond timestep.
-     For all model sizes, PyNengo simulators
-     are faster than JavaNengo,
-     which is significantly faster than NEURON and NEST,
-     which are significantly faster than Brian.
-     The full Brian results are not shown;
-     for the largest model, Brian takes $\sim$768 seconds.
-     (B) Benchmark results from simulating
-     the Lorenz attractor for 10 simulated seconds
-     at a 1 millisecond timestep.
-     For most model sizes,
-     the results are the same as (A),
-     except that NEURON is notably faster.
-     The full results for Brian and NEST are not shown;
-     for the largest model, they take $\sim$1467
-     and $\sim$601 seconds respectively.
-     (C) Benchmark results from simulating
-     circular convolution for 1 simulated second
-     at a 1 millisecond timestep.
-     For the blue lines, the simulator used
-     was the PyNengo OpenCL simulator.
-     The CPU used for JavaNengo and
-     the PyNengo reference simulator
-     was an Intel Core i7-3770;
-     all 4 cores were used by JavaNengo,
-     while PyNengo's reference simulator
-     only used one core.
-     For all model sizes,
-     the OpenCL simulator is faster
-     than the JavaNengo simulator,
-     which is faster than the PyNengo
-     reference simulator.
-     The reference simulator was only run
-     up to 50 dimensions.
-     The full results for JavaNengo
-     are not shown; for the largest model,
-     it takes $\sim$45 seconds.
-     }
-\end{figure}
-
 While benchmark models are not indicative
@@
 with Brian \citep{goodman2009}, NEURON \citep{hines2009},
-NEST \citep{eppler2008}, and PCSim \citep{pecevski2009}.
+and NEST \citep{eppler2008}.\footnote{We
+  were unable to compile PCSim \citep{pecevski2009} on the
+  machine on which we ran benchmarks.}
 We ran each parameter set five times
 on the same machine,
-and plot the mean.
+and plot the mean time elapsed.
 In most cases, the coefficient of variation
+for the five sample times
 is well below 0.1, except for two
@@
 Instead of running each simulation multiple times,
-we instead run the simulator for 10 timesteps
+we instead ran the simulator for 10 timesteps
 in order to fill various levels of CPU or GPU cache,
-and then run the simulator for 1000 more timesteps;
+and then ran the simulator for 1000 more timesteps;
 there is very little variance using this method.
@@
 and 50 times faster than JavaNengo.
+In the 50-dimensional case,
+the Radeon 7970 GPU is 200 times faster
+than JavaNengo.
 Additionally, although both JavaNengo
@@
 PyNengo's OpenCL simulator is significantly faster.
-
-\clearpage
 
@@
 while creating a new simulator is complex, but possible.
-
 By enabling outside contributions,
@@
 and simulators to Nengo.
-Internally, we will be
-implementing the JavaNengo use cases
-not currently covered by PyNengo,
-and create a graphical interface
+Internally, our short-term goal
+is to implement the JavaNengo use cases
+not currently covered by PyNengo.
+Our long-term goal is to
+create a graphical interface
 to build models and
@@
 50-200 times faster than JavaNengo,
-which is faster than alternative simulators
+which itself is faster than alternative simulators
 on simpler models.
@@
 These models will further test
-the NEF as a theory of the brain;
-PyNengo will make those models
-accessible to anyone with a modern computer.
+the NEF as a theory of neural computation;
+PyNengo makes those models
+accessible to anyone with
+a modern computer.
 
@@
 
-Trevor Bekolay led development of the
+TB led development of the
 PyNengo object model and scripting interface,
@@
 and prepared all of the figures.
-James Bergstra led development of
+JB led development of
 the PyNengo reference simulator and OpenCL simulator,
@@
 Figure~\ref{fig:benchmarks}C.
-Eric Hunsberger contributed significantly
+EH contributed significantly
 to PyNengo and both of its simulators, and edited text.
-Travis DeWolf led development
+TD led development
 of a Theano-backed
@@
 the version of PyNengo described in this paper.
-Terrence C. Stewart contributed
+TCS contributed
 to PyNengo, helped implement
@@
 on which PyNengo is based.
-Daniel Rasmussen contributed
+DR contributed
 to PyNengo and the reference simulator,
 and edited text.
-Xuan Choo contributed to PyNengo
+XC contributed to PyNengo
 and the reference simulator.
-Aaron Voelker ran the benchmarks
+AV ran the benchmarks
 shown in Figure~\ref{fig:benchmarks}A and B.
-Chris Eliasmith oversaw all development,
+CE oversaw all development,
 contributed to PyNengo,
@@
 
+\clearpage
+
+\begin{figure}
+ \textbf{\refstepcounter{figure}\label{fig:nef} Figure \arabic{figure}.}{
+   Summary of the three principles of the Neural Engineering Framework
+   (NEF). \textbf{(A)} By the representation principle, signals are encoded
+   in neural populations based on the \textit{tuning curve}
+   of each neuron (top). The tuning curve describes
+   how active a neuron is given some input signal.
+   If we drive the eight neurons in the top panel
+   with the input signal in the middle panel,
+   we see the spike trains in the bottom panel.
+   \textbf{(B)} By the representation principle,
+   the spiking activity of a neural population
+   can be decoded to recover the original input signal,
+   or some transformation of that input signal.
+   First, the firing pattern shown in the top panel
+   is filtered with a decaying exponential filter (middle panel).
+   The filtered activity is then summed together
+   with a set of weights that approximates
+   the input signal (bottom panel, green)
+   and the cosine of the input signal (bottom panel, purple).
+   \textbf{(C)} A sine wave is encoded by population A (top panel);
+   the negative of that signal is projected
+   to population B (middle panel)
+   and the square of that signal is projected
+   to population C (bottom panel).
+   By the transformation principle,
+   populations of neurons can send signals
+   to another population by decoding
+   the desired function from the first population
+   and then encoding the decoded estimate
+   into the second population.
+   These two steps can be combined into a single step
+   by calculating a set of weights
+   that describe the strength of the connections
+   between each neuron in the first population
+   and each neuron in the second population.
+   \textbf{(D)} A neurally implemented dynamical system
+   has negative feedback across its two dimensions,
+   resulting in a harmonic oscillator.
+   The oscillator is plotted across time (top)
+   and in state space (bottom).
+   By the dynamcs principle,
+   signals being represented by population of neurons
+   can be thought of as state variables in a dynamical system.
+   }
+\end{figure}
+
+\begin{figure}
+ \textbf{\refstepcounter{figure}\label{fig:sim} Figure \arabic{figure}.}{
+   Detailed breakdown of the PyNengo reference simulator
+   running a simple model for two timesteps.
+   \textbf{(A)} Code describing the model being simulated.
+   It consists of ensemble A projecting the sine of its
+   encoded signal to ensemble B,
+   which is recurrently connected.
+   \textbf{(B)} Diagram depicting the model being simulated.
+   \textbf{(C)} A detailed diagram of how the reference simulator
+   organizes this model. Signals (blue) represent the values
+   tracked in the simulation.
+   Operators (red) represent the computations done on signals.
+   Signals can be grouped as low-level neural signals
+   that are used to compute the nonlinear functions
+   underlying neuron models,
+   and high-level NEF signals that are used to
+   drive neurons and track the signals
+   that the neurons are representing.
+   The operators that implement the decoding
+   and encoding steps map between
+   the low-level neural signals
+   and the high-level NEF signals.
+   \textbf{(D)} The signals tracked at the low level
+   can be interpreted as a model
+   commonly seen in computational neuroscience literature;
+   a population of leaky integrate-and-fire neurons
+   is driven by some time-varying input current, $J(t)$.
+   These neurons project to a population
+   of recurrently connected neurons.
+   The connection weights between the two populations,
+   and from the second population to itself,
+   can be computed by the NEF's transformation
+   principle, bypassing the need for
+   the high-level NEF signals
+   used by the reference simulator
+   for speed and data collection purposes.
+   \textbf{(E)} The signals tracked at the high level
+   can be interpreted as a dynamical system.
+   State variable A simply represents its input,
+   and passes its state to a sine function
+   which becomes the input to B.
+   State variable B is a simple linear system
+   that can be described with the typical
+   $\dot{x}(t) = A x(t) + B u(t)$ equation.
+   These dynamical systems can be simulated
+   directly, without the use of spiking neurons,
+   in order to quickly analyze system behavior,
+   if desired.}
+\end{figure}
+
+\begin{figure}
+ \textbf{\refstepcounter{figure}\label{fig:comm-channel}
+   Figure \arabic{figure}.}{
+   A communication channel implemented with Nengo.
+   \textbf{(A)} Diagram depicting the model. Ensemble A
+   projects its encoded signal to ensemble B unchanged.
+   \textbf{(B)} Nengo code to build and simulate the model
+   for 1 second.
+   \textbf{(C)} The results of the simulation.
+   The input signal (top panel) is white noise limited to 0 to 5 Hz.
+   The signal is well represented by both ensemble A (middle panel)
+   and ensemble B (bottom panel) despite the neural firing patterns
+   (underlaid in middle and bottom panels) being different.}
+\end{figure}
+
+\begin{figure}
+ \textbf{\refstepcounter{figure}\label{fig:pynn} Figure
+   \arabic{figure}.}{
+   Implementation of the communication channel (left) in Nengo
+   and (right) in PyNN. Solving for decoding weights
+   takes approximately 40 lines of code,
+   which are not included in this figure.}
+\end{figure}
+
+\begin{figure}
+ \textbf{\refstepcounter{figure}\label{fig:lorenz} Figure \arabic{figure}.}{
+   A Lorenz attractor implemented with Nengo.
+   \textbf{(A)} Nengo code to build and simulate the model
+   for 6 seconds.
+   \textbf{(B)} Diagram depicting the model. The state ensemble
+   is recurrently connected with a complex function
+   implementing the dynamics of the Lorenz attractor.
+   Note that this populations does not receive
+   any input that might drive its initial value;
+   instead, the initial value is determined by
+   the baseline firing of the 2000 leaky integrate-and-fire
+   neurons that make up the state ensemble.
+   \textbf{(C)} The trajectory that the state ensemble takes
+   in its three-dimensional state space.
+   For the parameters chosen, the trajectory takes
+   the well-known butterfly shape.
+   \textbf{(D)} The state vector plotted over time.
+   \textbf{(E)} The spikes emitted by a random sample of 25
+   neurons from the state ensemble.
+   Some neurons fire uniformly across the 6 second simulation,
+   but most change depending on the state being tracked
+   due to the recurrent connection.}
+\end{figure}
+
+\begin{figure}
+ \textbf{\refstepcounter{figure}\label{fig:cconv} Figure
+   \arabic{figure}.}{
+   Circular convolution implemented with Nengo.
+   \textbf{(A)} Nengo code to build and simulate the model
+   for 0.2 seconds.
+   \textbf{(B)} Diagram depicting the model.
+   The input vectors, A and B, represent four-dimensional vectors
+   which are mapped onto six ensembles within the
+   Circular Convolution network through
+   complicated transformation matrices
+   that implement a discrete Fourier transform.
+   Each ensemble within the network represents a
+   two-dimensional vector. The product of the two dimensions
+   is projected through another complicated transformation matrix
+   that implements the inverse discrete Fourier transform,
+   computing the final four-dimensional result.
+   Note that the complicated parts of the model
+   are contained within the network;
+   the number of ensembles and the transform matrices shown
+   are automatically generated by the network depending on
+   the dimensionality of the input vectors.
+   \textbf{(C)} The result of the simulation.
+   Straight horizontal lines represent
+   the target values that each ensemble
+   should represent.
+   Wavy lines represent the decoded values
+   for each dimension represented by the
+   A, B, and Result ensembles
+   (top, middle and bottom panels, respectively).
+   The ensembles represent the correct values,
+   after a startup transient of less than 0.1 seconds.}
+\end{figure}
+
+\begin{figure}
+ \textbf{
+   \refstepcounter{figure}\label{fig:benchmarks}
+   Figure \arabic{figure}.}{
+   Benchmark results for several simulators
+   on the example models described
+   in section~\ref{sec:examples}.
+   In (A) and (B), all of the simulators
+   except the PyNengo OpenCL simulator
+   were run on an Intel Core i7-965.
+   JavaNengo used all 4 cores of this processor;
+   all other simulator used only 1 core.
+   The PyNengo OpenCL simulator
+   was run on an NVidia GTX280 GPU.
+   \textbf{(A)} Benchmark results from simulating
+   the communication channel for 10 simulated seconds
+   at a 1 millisecond timestep.
+   For all model sizes, PyNengo simulators
+   are faster than JavaNengo,
+   which is significantly faster than NEURON and NEST,
+   which are significantly faster than Brian.
+   The full Brian results are not shown;
+   for the largest model, Brian takes $\sim$768 seconds.
+   \textbf{(B)} Benchmark results from simulating
+   the Lorenz attractor for 10 simulated seconds
+   at a 1 millisecond timestep.
+   For most model sizes,
+   the results are the same as (A),
+   except that NEURON is notably faster.
+   The full results for Brian and NEST are not shown;
+   for the largest model, they take $\sim$1467
+   and $\sim$601 seconds respectively.
+   \textbf{(C)} Benchmark results from simulating
+   circular convolution for 1 simulated second
+   at a 1 millisecond timestep.
+   For the blue lines, the simulator used
+   was the PyNengo OpenCL simulator.
+   The CPU used for JavaNengo and
+   the PyNengo reference simulator
+   was an Intel Core i7-3770;
+   all 4 cores were used by JavaNengo,
+   while PyNengo's reference simulator
+   only used one core.
+   For all model sizes,
+   the OpenCL simulator is faster
+   than the JavaNengo simulator,
+   which is faster than the PyNengo
+   reference simulator.
+   The reference simulator was only run
+   up to 50 dimensions.
+   The full results for JavaNengo
+   are not shown; for the largest model,
+   it takes $\sim$45 seconds.
+ }
+\end{figure}
+
 \end{document}

### 2013-10-06 scripts/combine.py

@@ -1,0 +1,181 @@
+import os.path
+import subprocess
+
+import svgutils.transform as sg
+
+
+pt_to_px = 1.25
+px_to_pt = 1.0 / pt_to_px
+mm_to_px = 3.54
+
+
+def el(char, path, x, y, w=None, h=None, scale=1, offset=(0, 0)):
+    svg, textel = [], []
+
+    if char is not None:
+        textel = [sg.TextElement(x + offset[0] + 0.5, y + offset[1] + 11, char,
+                                 size=14, weight='bold', font='Arial')]
+
+    if path is not None:
+        svg = sg.fromfile(path)
+        if w is not None and h is not None:
+            svg.set_size((w, h))
+        svg = svg.getroot()
+        svg.moveto(str(x), str(y), scale)
+        svg = [svg]
+    return svg + textel
+
+
+def svgfig(w, h):
+    w = str(w) + 'px'
+    h = str(h) + 'px'
+    return sg.SVGFigure(w, h)
+
+
+def savefig(fig, out):
+    fig.save('%s.svg' % out)
+    print "Saved %s.svg" % out
+    subprocess.call(['inkscape',
+                     '--export-pdf=%s.pdf' % out,
+                     '%s.svg' % out])
+    print "Saved %s.pdf" % out
+    subprocess.call(['inkscape', '--export-text-to-path',
+                     '--export-eps=%s.eps' % out,
+                     '%s.svg' % out])
+    print "Saved %s.eps" % out
+    subprocess.call(['inkscape',
+                     '--export-dpi=1200',
+                     '--export-png=%s.png' % out,
+                     '%s.svg' % out])
+    print "Saved %s.png" % out
+    subprocess.call(['convert',
+                     '%s.png' % out,
+                     # '-compress', 'LZW',
+                     '%s.tiff' % out])
+    print "Saved %s.tiff" % out
+
+def check_fig(in_svg, out_svg):
+    for svg in in_svg:
+        if not os.path.exists(svg):
+            return False
+
+    if os.path.exists(out_svg + ".svg"):
+        return False
+
+    print "Generating " + out_svg + ".svg"
+    return True
+
+
+def fig1():
[…]

### 2013-10-06 scripts/plot_bench_cchannel.py

@@ -74,7 +74,9 @@ data['NEST'][(1000, 50)] = [
 
 cvs = []
 
-plt.figure(figsize=(4,5))
+mm_to_inches = 0.0393701
+figsize = (180. * mm_to_inches / 3., 180. * mm_to_inches / 2.5)
+plt.figure(figsize=figsize)
 ax = plt.subplot(1,1,1)
 
 x = data.values()[0].keys()
@@ -86,13 +88,13 @@ for ix, name in enumerate(data.keys()):
     for (n, d), times in data[name].items():
         cvs.append(np.std(times) / np.mean(times))
         runtimes.append(np.mean(times))
-    plt.plot(xval, runtimes, '.-', ms=15, color=colors[ix], label=name)
+    plt.plot(xval, runtimes, '.-', ms=8, color=colors[ix], label=name)
     if name == "Brian":
         plt.text(xval[-3] + 40, runtimes[-3], name, color=colors[ix],
-                 ha='left', va='center', fontsize=14)
+                 ha='left', va='center', fontsize='large')
     elif name in ("NEURON", "NEST"):
         plt.text(xval[-1] + 40, runtimes[-1], name, color=colors[ix],
-                 ha='left', va='center', fontsize=14)
+                 ha='left', va='center', fontsize='large')
 
 del xval[1]
 del xlabels[1]
@@ -118,7 +120,7 @@ for ix, name in enumerate(data.keys()):
     for (n, d), times in data[name].items():
         cvs.append(np.std(times) / np.mean(times))
         runtimes.append(np.mean(times))
-    inset.plot(xval, runtimes, '.-', ms=15, color=colors[ix], label=name)
+    inset.plot(xval, runtimes, '.-', ms=8, color=colors[ix], label=name)
     yoffset = 0
     if name == "PyNengo\nOpenCL":
         yoffset = 0.85
@@ -126,7 +128,7 @@ for ix, name in enumerate(data.keys()):
         yoffset = -0.85
     if 'Nengo' in name:
         plt.text(xval[-1] + 7, runtimes[-1] + yoffset, name, color=colors[ix],
-                 ha='left', va='center', fontsize=14)
+                 ha='left', va='center', fontsize='large')
 
 
 inset.set_xlim(980, 1065)
@@ -140,7 +142,7 @@ inset.set_xticks(())
[…]

### 2013-10-06 scripts/plot_bench_cconv.py

@@ -71,12 +71,14 @@ data['Intel Xeon\nE5-2620'][100] = 1.68300008774
 data['Intel Xeon\nE5-2620'][200] = 3.46999979019
 data['Intel Xeon\nE5-2620'][500] = 12.3960001469
 
-plt.figure(figsize=(4,5))
+mm_to_inches = 0.0393701
+figsize = (180. * mm_to_inches / 3., 180. * mm_to_inches / 2.5)
+plt.figure(figsize=figsize)
 ax = plt.subplot(1,1,1)
 
 for ix, name in enumerate(data.keys()):
     n = [neurons[d] for d in data[name].keys()]
-    plt.plot(n, data[name].values(), '.-', ms=15, color=colors[ix])
+    plt.plot(n, data[name].values(), '.-', ms=8, color=colors[ix])
     yoffset = 0
     if "3770" in name:
         yoffset = 0.68
@@ -93,10 +95,10 @@ for ix, name in enumerate(data.keys()):
 
     if name == "JavaNengo":
         plt.text(n[1] + 12000, data[name].values()[1], name, color=colors[ix],
-                 ha='left', va='center', fontsize=12)
+                 ha='left', va='center', fontsize='large')
     else:
         plt.text(n[-1] + 12000, data[name].values()[-1] + yoffset, name,
-                 color=colors[ix], ha='left', va='center', fontsize=12)
+                 color=colors[ix], ha='left', va='center', fontsize='large')
 
 dims = [5, 200, 500]  # Ignoring 10, 20, 50, 100
 xval = [neurons[d] for d in dims]
@@ -106,7 +108,7 @@ plt.xlim(0, 500000)
 plt.ylim(0, 27.5)
 plt.text(280000, 28.5, 'Circular convolution benchmarks',
          ha='center', va='center', fontsize='large')
-ax.tick_params(axis='x', which='major', labelsize=11)
+ax.tick_params(axis='x', which='major', labelsize='small')
 plt.ylabel('Simulation time (s)')
 ax.spines['top'].set_visible(False)
 ax.spines['right'].set_visible(False)
@@ -114,4 +116,4 @@ ax.xaxis.set_ticks_position('none')
 ax.yaxis.set_ticks_position('left')
 
 plt.subplots_adjust(left=0.125, bottom=0.085, top=0.93, right=0.82)
-plt.savefig("../figures/bench_cconv.pdf")
+plt.savefig("../figures/bench_cconv.svg")

### 2013-10-06 scripts/plot_bench_lorenz.py

@@ -74,7 +74,9 @@ data['NEST'][2000] = [
 
 cvs = []
 
-plt.figure(figsize=(4,5))
+mm_to_inches = 0.0393701
+figsize = (180. * mm_to_inches / 3., 180. * mm_to_inches / 2.5)
+plt.figure(figsize=figsize)
 ax = plt.subplot(1,1,1)
 
 xval = data.values()[0].keys()
@@ -85,13 +87,13 @@ for ix, name in enumerate(data.keys()):
     for n, times in data[name].items():
         cvs.append(np.std(times) / np.mean(times))
         runtimes.append(np.mean(times))
-    plt.plot(xval, runtimes, '.-', ms=15, color=colors[ix])
+    plt.plot(xval, runtimes, '.-', ms=8, color=colors[ix])
     if name == "Brian":
         plt.text(xval[2] + 60, runtimes[2], name, color=colors[ix],
-                 ha='left', va='center', fontsize=14)
+                 ha='left', va='center', fontsize='large')
     elif name == "NEST":
         plt.text(xval[-1] + 60, runtimes[-1], name, color=colors[ix],
-                 ha='left', va='center', fontsize=14)
+                 ha='left', va='center', fontsize='large')
 
 plt.xticks(xval, xlabels)
 plt.xlim(50, 2100)
@@ -113,7 +115,7 @@ for ix, name in enumerate(data.keys()):
     for n, times in data[name].items():
         cvs.append(np.std(times) / np.mean(times))
         runtimes.append(np.mean(times))
-    plt.plot(xval, runtimes, '.-', ms=15, color=colors[ix])
+    plt.plot(xval, runtimes, '.-', ms=8, color=colors[ix])
     yoffset = 0
     if name == "PyNengo\nreference":
         yoffset = 1.5
@@ -121,7 +123,7 @@ for ix, name in enumerate(data.keys()):
         yoffset = -1.5
     if 'Nengo' in name or name == 'NEURON':
         plt.text(xval[-1] + 4, runtimes[-1] + yoffset, name, color=colors[ix],
-                 ha='left', va='center', fontsize=14)
+                 ha='left', va='center', fontsize='large')
 inset.set_xlim(1995, 2042)
 inset.set_ylim(-2, 21.5)
 inset.xaxis.set_ticks_position('bottom')
@@ -133,7 +135,7 @@ inset.set_xticks(())
[…]

### 2013-10-06 scripts/plot_cconv.py

@@ -50,9 +50,11 @@ def plot(actual, ens):
     plt.plot(sim.data(model.t), sim.data(ens))
     for color, l in zip(colors, actual):
         plt.axhline(l, color=color)
-    plt.xlim(0, 0.5)
+    plt.xlim(0, 0.2)
 
-plt.figure(figsize=(5,5))
+mm_to_inches = 0.0393701
+figsize = (40 * mm_to_inches, 49.16 * mm_to_inches)
+plt.figure(figsize=figsize)
 plt.subplot(3, 1, 1)
 plot(a_in, "A")
 plt.gca().spines['bottom'].set_visible(False)
@@ -70,5 +72,5 @@ plot(actual_out, "Result")
 plt.xlabel("Time (s)")
 plt.ylabel("Result")
 
-plt.tight_layout(h_pad=0.12)
-plt.savefig("../figures/cconv_res.pdf")
+plt.tight_layout(pad=0.1, h_pad=0.12)
+plt.savefig("../figures/cconv_res.svg")

### 2013-10-06 scripts/plot_comm_channel.py

@@ -36,16 +36,18 @@ def adjust(ax):
     ax.set_yticks((-1, 0, 1))
     ax.axis([0, 1, -1.25, 1.25])
 
-plt.figure(figsize=(4,5))
+mm_to_inches = 0.0393701
+figsize = (85. * mm_to_inches / 2., (39.31 + 4) * mm_to_inches)
+plt.figure(figsize=figsize)
 ax = plt.subplot(3,1,1)
 ax.plot(sim.data(model.t), sim.data("Input"), color='k')
-ax.text(0.5, 1.0, "Input", ha='center', va='center', fontsize=20)
+ax.text(0.5, 1.0, "Input", ha='center', va='center', fontsize='large')
 adjust(ax)
 ax.spines['bottom'].set_visible(False)
 ax.set_xticks(())
 
 ax = plt.subplot(3,1,2)
-ax.eventplot(a_spikes, colors=[(0.75,0.75,0.75)], rasterized=True)
+ax.eventplot(a_spikes, colors=[(0,0,0)], rasterized=True, linewidths=0.1)
 ax.spines['bottom'].set_visible(False)
 ax.spines['top'].set_visible(False)
 ax.spines['left'].set_visible(False)
@@ -54,14 +56,14 @@ ax.axis([0, 1, -0.5, len(a_spikes)-0.5])
 
 ax = ax.twinx()
 ax.plot(sim.data(model.t), sim.data("A"), color='k')
-ax.text(0.5, 1.0, "A", ha='center', va='center', fontsize=20,
+ax.text(0.5, 1.0, "A", ha='center', va='center', fontsize='large',
         bbox=dict(ec='none', fc='w', alpha=0.8))
 ax.spines['bottom'].set_visible(False)
 adjust(ax)
 ax.set_xticks(())
 
 ax = plt.subplot(3,1,3)
-ax.eventplot(b_spikes, colors=[(0.75,0.75,0.75)], rasterized=True)
+ax.eventplot(b_spikes, colors=[(0,0,0)], rasterized=True, linewidths=0.1)
 ax.spines['top'].set_visible(False)
 ax.spines['left'].set_visible(False)
 ax.xaxis.set_ticks_position('bottom')
@@ -71,9 +73,9 @@ ax.set_xlabel("Time (s)")
 
 ax = ax.twinx()
 ax.plot(sim.data(model.t), sim.data("B"), color='k')
[…]

### 2013-10-06 scripts/plot_lorenz.py

@@ -32,19 +32,18 @@ sim = model.simulator()
 sim.run(6)
 
 state_spikes = sorted_spikes(sim, "State", iterations=250, every=80)
-#state_spikes = [np.linspace(0, 6, 6 * 100)]
 
-plt.figure(figsize=(7,4.55))
+mm_to_inches = 0.0393701
+figsize = (100. * mm_to_inches, 80. * mm_to_inches)
+plt.figure(figsize=figsize)
 ax = plt.subplot2grid((2,3), (0,0), colspan=2, rowspan=2, projection='3d')
 ax.plot(sim.data('State')[:,0], sim.data('State')[:,1], sim.data('State')[:,2],
         color='k')
-# ax.set_xticklabels(())
-# ax.set_yticklabels(())
-# ax.set_zticklabels(())
-ax.dist = 8.5
+ax.dist = 9
 
 ax = plt.subplot2grid((2,3), (0,2))
 ax.plot(sim.data(model.t), sim.data('State'))
+ax.set_xlim(0, 6)
 ax.set_ylabel("Amplitude")
 ax.spines['top'].set_visible(False)
 ax.spines['left'].set_visible(False)
@@ -54,8 +53,7 @@ ax.yaxis.tick_right()
 ax.yaxis.set_label_position("right")
 
 ax = plt.subplot2grid((2,3), (1,2))
-ax.eventplot(state_spikes, colors=[(0.1,0.1,0.1)], linewidths=0.05,
-             rasterized=True)
+ax.eventplot(state_spikes, colors=[(0,0,0)], rasterized=True, linewidths=0.1)
 ax.axis([0, 6, len(state_spikes)-0.5, -0.5])
 ax.set_xlabel("Time (s)")
 ax.set_ylabel("Neuron")
@@ -65,6 +63,6 @@ ax.xaxis.set_ticks_position('bottom')
 ax.yaxis.tick_right()
 ax.yaxis.set_label_position("right")
 
-plt.tight_layout(h_pad=0.12)
+plt.tight_layout(pad=0.1, h_pad=0.12)
 plt.subplots_adjust(left=0)
-plt.savefig("../figures/lorenz_res.pdf")
+plt.savefig("../figures/lorenz_res.svg")

### 2013-10-06 scripts/plot_nef_summary.py

@@ -27,7 +27,8 @@ cos = model.connect("A", "Dummy", function=np.cos, filter=0.03)
 model.probe(cos)
 
 ## Encoding plot
-figsize = (3.5, 7)
+mm_to_inches = 0.0393701
+figsize = (180. * mm_to_inches / 4.0, 180. * mm_to_inches / 2.0)
 
 sim = model.simulator()
 sim.run(1)
@@ -74,7 +75,7 @@ plt.xlabel("Time (s)")
 plt.ylabel("Neuron")
 
 plt.tight_layout(pad=0.1, h_pad=0.1)
-plt.savefig("../figures/nef_summary_enc.pdf")
+plt.savefig("../figures/nef_summary_enc.svg")
 
 ## Decoding
 # NB: We don't make a new network so that the plots match
@@ -135,8 +136,7 @@ plt.xlabel("Time (s)")
 plt.axis([0, 1, -1.25, 1.25])
 
 plt.tight_layout(pad=0.1, h_pad=0.5)
-plt.savefig("../figures/nef_summary_dec.pdf")
-
+plt.savefig("../figures/nef_summary_dec.svg")
 
 ## Transformation
 
@@ -162,7 +162,7 @@ model.probe("C", filter=0.01)
 sim = model.simulator()
 sim.run(2)
 t = sim.data(model.t)
-grey = [(0.75, 0.75, 0.75)]
+grey = [(0, 0, 0)]
 
 plt.figure(figsize=figsize)
 
@@ -178,7 +178,7 @@ plt.yticks(())
 
 plt.gca().twinx()
 plt.plot(sim.data(model.t), sim.data("A"))
-plt.text(1, 1, "A", ha='center', va='center', fontsize=16,
+plt.text(1, 1, "A", ha='center', va='center', fontsize='large',
          bbox=dict(ec='none', fc='w', alpha=0.8))
 plt.axhline(0, color='k')
 plt.xticks(())
@@ -199,7 +199,7 @@ plt.yticks(())
 
 plt.gca().twinx()
 plt.plot(sim.data(model.t), sim.data("B"))
-plt.text(1, 1, "B=-A", ha='center', va='center', fontsize=16,
+plt.text(1, 1, "B=-A", ha='center', va='center', fontsize='large',
          bbox=dict(ec='none', fc='w', alpha=0.8))
 plt.axhline(0, color='k')
 plt.xticks(())
@@ -221,7 +221,7 @@ plt.xlabel("Time (s)")
 
[…]

### 2013-12-04 rebuttal.txt
Manuscript revised based on reviewer comments
- Tried to address all reviewer concerns
- Also included responses to reviewer comments in rebuttal.txt

Dear Editor,

Please find a revised version of our Methods paper "Nengo: A Python
tool for building large-scale functional brain models" attached for
your consideration for publication in Frontiers in Neuroinformatics.

Sincerely,

Trevor Bekolay, James Bergstra, Eric Hunsberger, Travis DeWolf,
Terrence C. Stewart, Daniel Rasmussen, Xuan Choo, Aaron Voelker, and
Chris Eliasmith

Reviewer 1
==========

Expert Authority
----------------

>> Are relevant methods of other authors mentioned? If not, please
>> specify the methods the author should incorporate.

> Emergent should be mentioned. It is not based on spiking simulators,
> but has a track record in cognitive modelling.

We have now cited Emergent in the introduction alongside ACT-R in the
introduction.

> MIIND could be mentioned (Disclaimer: the reviewer is involved with
> MIIND) It models population dynamics, admittedly not based on
> spiking neurons, but again its aim is large-scale cognitive models
> and shares some design ideas (see paper in Neural Networks). Please
> judge for yourself.

We have revised Section 7.1, which now includes a list of tools
that simulate cognitive phenomena, including MIIND.

The revised text reads:

"There are many other neural simulators dedicated to building
large-scale neural models (Eppler et al., 2008; Goodman and Brette,
2009; Hines et al., 2009), and many tools for simulating cognitive
phenomena with various levels of biologically plausibility (Cooper and
Fox, 1998; Sun, 2001; Anderson et al., 2004; Franklin et al., 2007;
Aisa et al., 2008; de Kamps et al., 2008; Laird, 2012). However, Nengo
is unique in being built on a theoretical framework that has enabled a
cognitive architecture (the Semantic Pointer Architecture) that
maintains a high level of biological plausibility, and has been
validated through the Spaun model and other past work."

Other Comments
--------------

>> Please add any further comments you have regarding this manuscript.

> It looks like an interesting tool, and I think at this level
> existing simulators can do with a bit of competition. But the
> description of the tool leaves much to be desired. In particular it
> doesn't make clear enough what its advantages over NEST or BRIAN
> are. The discussion on scaling is not entirely convincing because in
> part, it is hard to understand how to compare with those simulators.

We have expanded Section 7.1 to compare to projects that have similar
design goals and present a similar interface (e.g., PyNN) and between
the underlying neural simulators (e.g., NEST, Brian).

The revised section on neural simulators reads:

"On the neural simulator side, we have shown that both Nengo's
reference simulator and OpenCL simulator are able to simulate two
benchmark models much faster than Brian, NEST and NEURON (see
Figure~\ref{fig:benchmarks}). This is, in part, because Nengo stores
the factors of the connection weight matrix, rather than storing the
entire matrix. However, these simulators are able to simulate many
detailed neuron models and learning rules, and have access to a wealth
of existing neuron models and learning rules. Because Nengo 2.0 is in
an earlier development stage, many of these detailed neuron models and
learning rules remain to be added. Neural simulators like Brian, NEST,
and NEURON are therefore currently better suited for simulating a
wider range of single cell models, while Nengo is designed for large
networks of simple neural models that are connected together according
to the principles of the Neural Engineering Framework, and can
simulate these types of models efficiently.

"One key difference between Nengo's simulators and traditional neural
simulators is the target platform. While NEST and NEURON can be run on
commodity hardware, networks of modest size are typically simulated in
high-performance computing environments by using the Message-Passing
Interface (MPI). Nengo enables large-scale neural simulation on
commodity hardware, allowing researchers and hobbyists without access
to large computing clusters the ability to experiment with theoretical
neuroscience as is currently practiced in cutting edge research. In
particular, GPUs are a powerful, low-cost computing resource that are
available in most modern workstations. The OpenCL simulator makes full
use of GPUs, while the previously discussed simulators currently do
not."

> Where is it available? What do I need in terms of 3rd party software?

Thanks for pointing out that we omitted these important details in the
original manuscript. We have added a link to the release and source
versions of Nengo in a footnote on Page 2 of the revised manuscript
that reads:

"Nengo is available for download through the Python Package Index at
https://pypi.python.org/pypi/nengo. The full source is available at
https://github.com/ctn-waterloo/nengo."

We have also added the following paragraph to the introduction in
order to clarify Nengo's 3rd party dependencies.

"Nengo 2.0 has been rewritten from scratch in Python, leveraging NumPy
(Oliphant, 2007) for manipulating large amounts of data. While NumPy
is its only dependency, Nengo contains optional extensions for
plotting if Matplotlib is available (Hunter, 2007) and for interactive
exploration if IPython 63 available (Perez and Granger, 2007). Since
Nengo only depends on one third-party library, it is easy to integrate
Nengo models in arbitrary CPython programs, opening up possibilities
for using neurally implemented algorithms in web services, games, and
other applications."

> The NEF should be explained in more detail. The paper should be more
> self-contained. It is not entirely clear how the simulations are
> implemented. I think an ensemble of, for example, LIF neurons is
> simulated, and the dynamics and input that they receive AS A
> POPULATION is presented. That is interesting, but I'm not entirely
> sure how this is done. What happens in the Lorenz attractor? Do all
> neurons receive feedback according to the differential equations in
> the script, i.e. as a current? Or as spike trains modulated by this
> system? What is 'input' anyway? This is not described very well at
> all. I would walk the reader through a simple simulation more
> explicitly.

In response to your direct question, all of the neurons in the Lorenz
attractor receive input in the form of current; that current is
determined by a summation of the spiking activities of afferent
neurons weighted by connection weights that are determined using the
NEF's transformation principle.

These details are now included in Appendix 1, "Neural Engineering
Framework Details", which we have added to make this paper more
self-contained. This appendix gives the mathematical details behind
the principles discussed in Section 2. (The text of the Appendix is
too long to include in the interactive review forum.)

We have also described the example scripts in more detail in
sections 5.1, 5.2, and 5.3.

In regard to walking the reader through a simple simulation, we hope
that the addition of the NEF appendix will make the existing Figure 2
a sufficiently detailed account of the first two timesteps of a simple
simulation.

> Having said that, the paper contains far too many irrelevant design
> details. The paragraph on JAVANENGO is bizarre. Do you really want
> to burden you readers with the fact that its class interfaces are
> overloaded? If you want to promote the new Python tool, there is no
> reason to blast the Java version. I recommend dropping the entire
> JAVANENGO section, if the Python version is what you want people to
> use.

Thanks for pointing this out. As the designers of the new version, we
have had many discussions about whether we should improve the old
version or start from scratch; some of that discussion appeared in the
paper, and upon a second look, we wholeheartedly agree that the
mentions of JavaNengo are out of place.

We have removed that section completely in the revised manuscript.
Where mentions of the previous version of Nengo were unavoidable, we
have referred to JavaNengo as Nengo 1.4, and the new version as Nengo
2.0 rather than PyNengo.

> I would mention design considerations only when the relate to use
> cases, i.e. explain why certain scaling properties are achieved. The
> paper contains far too many design considerations, it should focus
> on use cases. Scaling is also an issue of use.
>
> This question is important in order to judge how similar the
> simulator is from other tools such as NEST and BRIAN. Can you set up
> simulations in the two of those that are similar to NENGO and THEN
> compare them? Also bear in mind that NEST has an MPI backbone.

We wanted this paper to focus on the design of Nengo 2.0, and so our
intention in saying that the implemented simulators scale well is
primarily to indicate that the design enables the implementation of
simulators that scale well. It is difficult to come up with use cases
that can be adequately explained in this paper; in a sense, the papers
that are cited in the introduction are those use cases, as they were
all implemented with Nengo 1.4 (Conklin and Eliasmith, 2005; Singh and
Eliasmith, 2006; Choo and Eliasmith, 2010; Rasmussen and Eliasmith,
2014; DeWolf and Eliasmith, 2011; Stewart et al., 2012; Eliasmith et
al., 2012). We believe that the examples discussed in Section 5 are of
sufficient length to be instructive.

Setting up simulations first in Brain or NEST (or PyNN) and then
mapping those to Nengo models would not be very instructive. The point
of Nengo is to build neural models based on the principles of the NEF;
models built in those other tools would not use these principles, and
would result in models containing many ensembles whose underlying
neural populations would be connected directly with a defined weight
matrix. We do make this possible for certain important cases (e.g.,
inhibitory gating signals), but these cases are not the norm when
using the principles of the NEF. We hope that the focus of the paper
is clearer given our additions in the introduction and in Section 7.1.

We agree that it would be an interesting comparison to simulate two
neural populations connected directly with a defined weight matrix.
However, that comparison would be more apt in a paper that gives more
implementation details of the simulators being compared. We have not
provided those details here, and have instead focused on the novel
contributions of Nengo.

In sum, the difference between Nengo and other neural simulators is
primarily that Nengo is designed to define and simulate neural models
created with the principles of the NEF. The updated introduction and
Section 7.1 (copied above) highlights this difference.

> Can you do other neural models than LIF? Again this harks back to
> the criticism that the philosophy of NEF is well described, and
> interesting, but its implementation is not.

Currently in the main development branch of Nengo 2.0, we have only
implemented the LIF model, and a rate approximation of the LIF model.
Nengo 1.4 contained implementations of LIF, an adaptive LIF model
(Koch & Schutter, 1999), and the well-known Hodgkin-Huxley model, so
we know that it is possible to use other neuron models in NEF-based
networks. Additionally, we have already completed work done on another
simulator internal to the lab that uses a neuron model emulating a
piece of neuromorphic hardware.

The added appendix describing the NEF makes explicit how the neuron
model is used. Additionally, we have added the following text to
Section 2:

"Importantly, tuning curves can be determined for any type of neuron,
and therefore the encoding process (and the NEF as a whole) is not
dependent on any particular neuron model."

> Two million spiking neurons is not that much. Edelman and Tononi did
> that 15 years ago. Izhekevich blasts you out of the water. Moreover,
> it is a meaningless quantity, it depends on what you do. 2.5 million
> compartmental neurons is still impressive, 2.5 point model neurons is
> not. In our lab have models with hundreds of thousands of rate-based
> populations, we could model each population by a thousand neurons if
> we wanted to. Does that make us record holder? I'd be a bit more
> careful with such claims, or qualify them better.

We agree that the number of neurons is a meaningless quantity, and
believe that function is the more important metric. That is why our
focus is well-exemplified by Spaun, the largest functional brain model
(which has been previously stated as such in Eliasmith & Trujillo,
2013), where functional in this case refers to the cognitive functions
that Spaun performs.

> In my opinion, this paper cannot be read without going to the original
> paper explaining NEF. Given the brevity of this paper, I think that's
> not good enough. There is ample space for a better explanation of how
> it works and how that is different from more conventional simulators.

Thank you for pointing out these weaknesses. We hope that the addition
of an appendix describing the NEF, an expanded Section 7.1 comparing
Nengo to conventional simulators, and other improvements throughout
the revised manuscript have addressed your concerns and have made this
paper self-contained. In particular, we hope that we have made it
clear that Nengo's strength lies in the fact that it is based on the
principles of the NEF; we do not wish to claim that Nengo is a better
general-purpose neural modeling tool than PyNN, or a better
general-purpose simulator than Brain or NEST. However, we do wish to
claim that Nengo is a better modeling tool and simulator for NEF-style
networks, and that NEF-style networks are less computationally
intensive to simulate, and therefore are feasible to run on commodity
hardware.

Reviewer 2
==========

Expert Authority
----------------

>> Are relevant methods of other authors mentioned? If not, please
>> specify the methods the author should incorporate.

> Authors could perhaps mention the Topographica
> (http://topographica.org/index.html) simulation framework which is
> also implemented in python and offers a high-level model
> specification environment. It also offers a lot of analysis and
> visualization resources and a GUI, which could serve as inspiration
> for some of the future work mentioned in the manuscript.

Thank you for bringing Topographica to our attention! While we have
looked at it some time ago, the project seems to have grown much
lately, and we will investigate it more soon. For now, we have cited
it in Section 7.1, as it seems to have similar goals as Nengo and
PyNN.

Article Length
--------------

>> A Method Article should not exceed 12,000 words. Should any part of
>> the article be shortened? If yes, please specify which part should be
>> shortened.

> I found the introduction unnecessarily lengthy with rather small
> proportion actually discussing the main topic (pyNENGO). Especially
> the JavaNENGO section could in my opinion be omitted or at least
> significantly shortened. Also I found the first part of the
> introduction discussing NEF unnecessarily long - this paper should
> motivate pynengo not NEF.

Thanks for pointing this out. As the designers of the new version, we
have had many discussions about whether we should improve the old
version or start from scratch; some of that discussion appeared in the
paper, and upon a second look, we wholeheartedly agree that the
mentions of JavaNengo are out of place.

We have removed that section completely in the revised manuscript.
Where mentions of the previous version of Nengo were unavoidable, we
have referred to JavaNengo as Nengo 1.4, and the new version as Nengo
2.0 rather than PyNengo.

We have also removed some sections of the introduction motivating the
NEF, and have replaced them with motivations for Nengo. Specifically,
the paragraph:

"It has often been said that neuroscience is data-rich and theory-poor
(Churchland and Sejnowski, 1992). We believe that the modeling success
of large functional networks such as Spaun suggests that the NEF may
be able to fill the theoretical void. In order to continue to evaluate
the NEF as a theory of neural computation by building larger and more
complicated models, we have begun the next generation of Nengo
development with a focus on speed, extensibility, and simplicity."

has been removed, and replace with the following text:

"The transformation principle of the NEF proposes that the connection
weight matrix between two neural populations can compute a nonlinear
function, and can be factored into two significantly smaller matrices.
By using these factors instead of full connection weight matrices,
NEF-designed models are more computationally efficient, which allows
Nengo to run large-scale neural models on low-cost commodity
hardware."

Other Comments
--------------

> Section 1.2 alludes to the possibility of exporting model from pynengo
> to PyNN. Is this a future work, or is there already some
> infrastructure in place for this? Please clarify.

The ability to export a model to PyNN is a future work. We have
removed that confusing passage from Section 1.2 in the revised
manuscript, but that functionality is possible and we hope to
implement it soon.

We have already have two work-in-progress simulators allowing for
Nengo 2.0 models to be run on two pieces of neuromorphic hardware. We
have added a mention of these, copied below, to Section 7.2 in order
to show that multiple backend simulators are possible (including a
future PyNN backend):

"We are also developing two simulators that will take the same NEF
model description as the existing simulators, but will target two
pieces of neuromorphic hardware to achieve greater speed and power
efficiency than the OpenCL simulator."

> In section 3.2 authors discuss parameters, some of which can be set on
> ensembles and some directly on neural populations. It isn't very
> intuitive why some belong to one or the other (i.e. why maximum firing
> rate is a parameter of ensemble rather than the population). I feel it
> could be instructive to discuss a bit more in detail the separation of
> the high level NEF concepts from the low-level spiking neuron concepts
> in pynengo.

The intuition as to why some parameters are ensemble-level and some
are neural population-level relies on a mathematical description of
how we compute the current that we inject into each of the neuron in
the neural population. Any parameter that is part of that equation is
ensemble-level, and any parameter that is not is stored on the neural
population. We have now included these equations in an appendix that
gives a mathematical description of the NEF; specifically, Equation
(1) describes the current that is injected into each neuron model. The
parameters of this equation are stored on the ensemble, whereas the
parameters that are used in the neural nonlinearity would be stored in
the neural population object.

In the revised manuscript we have also removed the mention of the
maximum firing rate in Section 3.1 (formerly 3.2), as the encoding
weight is more intuitively a part of the NEF's representation scheme
specifically and not a property of a neuron.

> I was intrigued by the mention of automatic parameter setting based on
> neurobiological constraints. Could authors discuss this issue further
> in the manuscript?

The description "automatic parameter setting" may be too generous due
to our choice of wording. Specifically, in our original manuscript we
wrote in Section 3.2 (now 3.1):

"If these attributes are not set, Nengo attempts to maintain
neurobiological constraints by selecting neural parameters from
distributions consistent with neocortical pyramidal cells."

However, it is more accurate to say that, if parameters are unset,
Nengo will use defaults that correspond to neocortical pyramidal
cells. The text in the updated manuscript now reads:

"If these attributes are not set, Nengo attempts to maintain
neurobiological constraints by using default parameters consistent
with neocortical pyramidal cells."

We do wish to do some kind of automatic parameters setting in the
future, both in terms of neural parameters and connection parameters.
In the newly added NEF appendix, we also mention how connection
parameters can correspond to experimental data:

"The decoding process is depicted in Figure 1B, where the optimal
linear decoders have been found and used for eight neurons. Notably,
this kind of temporal decoding requires an assumption about the nature
of the temporal filter being used. Here we assume that post-synaptic
currents are such filters, and set the time constants to reflect the
kind of neurotransmitter receptors in the connection (e.g., AMPA
receptors have short time constants, ~10 ms, and NMDA receptors have
longer time constants, ~50 ms)."

> It isn't very clear from the manuscript what Network is. Is it really
> just a simplest form of a container as would appear from the
> description or does it actually offer some distinct functionality?

A Network really is just a simple container, which is used for
organizing large models. Our experience building large models has
shown that this kind of hierarchical organization becomes increasingly
useful as the size of the model increases, and we include Networks for
that reason. The text in Section 3.5 has been revised to explicitly
state that networks are only containers, and can be used for
hierarchical grouping.

> The first paragraph of section 4 implies a simulator dependent
> results. If this is the case, I think it warrants a more detailed
> discussion.

Thanks for pointing this out. We have revised this paragraph to remove
the simulator dependent language, and have moved that information to
Section 4.1, which collects together information specific to the
reference simulator.

The revised first paragraph of Section 4 is:

"Decoupling model creation and simulation has been done previously by
PyNN (Davison et al., 2008). In PyNN, the same Python script can be
used to run a model on four different simulators. Nengo follows this
programming model by decoupling neural model creation and simulation,
which enables Nengo simulators to allocate memory and schedule
computations in the most efficient ways possible. Simulators are given
a Model as an argument; this Model is a static symbolic description.
The simulator can take the model description and build whatever data
structures are best suited to implement the simulation."

> The benchmark results are impressive, but it should be noted that
> most spiking simulators these days are used in HPC environments, and
> thus parallelization comparisons in multi-processor environments
> would be more relevant. If pynengo does not offer this
> functionality, this should be acknowledged and perhaps future plans
> in this direction should be discussed.

This is a good point. While we are currently implementing a simulator
that uses MPI, in this paper we want to highlight that Nengo is able
to simulate large networks that are typically only accessible in HPC
environments on commodity hardware. The paragraph in the introducton,
listed above, makes this explicit early on. We have also included the
following paragraph in Section 7.1:

"One key difference between Nengo's simulators and traditional neural
simulators is the target platform. While NEST and NEURON can be run on
commodity hardware, networks of modest size are typically simulated in
high-performance computing environments by using the Message-Passing
Interface (MPI). Nengo enables large-scale neural simulation on
commodity hardware, allowing researchers and hobbyists without access
to large computing clusters the ability to experiment with theoretical
neuroscience as is currently practiced in cutting edge research. In
particular, GPUs are a rich, low-cost computing resource that are
available in most modern workstations. The OpenCL simulator makes full
use of GPUs, while other simulators currently do not."

> Googling pynengo doesn't return anything, and I thought I can't find
> the package at all until I noticed that in table 1 it says that
> pynengo is the 'nengo' repository. I would advise authors making this
> more clear, to facilitate the promotion of pynengo. Perhaps end the
> introduction or discussion with clear instructions how to access the
> software.
>
> Did authors considered releasing pynengo as pypi package?

Thanks for bringing this up. As mentioned above, we have changed all
references to PyNengo to Nengo 2.0. Additionally, we have pointed to
Nengo on PyPI and Github in a footnote on page 2:

"Nengo is available for download through the Python Package Index at
https://pypi.python.org/pypi/nengo. The full source is available at
https://github.com/ctn-waterloo/nengo."

Currently, we have only registered the nengo namespace on PyPI, but we
are hoping to make an initial 2.0 release before the publication of
this paper.

> Section 7.2 about future talks about future work only in the last
> sentence.

Thank you for noticing this. We have completely rewritten Section 7.2
to more appropriately focus on future work.

### 2013-12-04 nengo.tex
Manuscript revised based on reviewer comments
- Tried to address all reviewer concerns
- Also included responses to reviewer comments in rebuttal.txt

@@
   to explain experimental data.
-  A previous version of Nengo
-  implemented in Java was used
-  to create Spaun,
+  Nengo 1.4, which was implemented in Java,
+  was used to create Spaun,
   the world's largest functional
@@
   highlighted limitations
-  in Nengo's ability to
+  in Nengo 1.4's ability to
   support model construction with simple syntax,
@@
   for subsequent analysis.
-  This paper describes a new version of Nengo
-  implemented in Python that overcomes these limitations.
+  This paper describes Nengo 2.0,
+  which is implemented in Python
+  and overcomes these limitations.
   It uses simple and extendable syntax,
   simulates a benchmark model on the scale of Spaun
-  50 times faster than the previous implementation,
+  50 times faster than Nengo 1.4,
   and has a flexible mechanism
@@
 and there are good reasons to think
-that cognition may never emerge \citep{trujillo-inpress}.
+that cognition may never emerge \citep{eliasmith2013a}.
 At the other end of the spectrum,
-cognitive architectures \citep{anderson2004}
+cognitive architectures \citep{anderson2004,aisa2008}
 and machine learning approaches \citep{hinton2006}
@@
 
-Nengo is a neural simulator
+Nengo\footnote{Nengo is available for download through
+  the Python Package Index at
+  \url{https://pypi.python.org/pypi/nengo}.
+  The full source is available at
+  \url{https://github.com/ctn-waterloo/nengo}.}
+is a neural simulator
 based on a theoretical framework proposed
@@
 list memory [\citealp{choo2010}],
-inductive reasoning [\citealp{rasmussen2011}],
+inductive reasoning [\citealp{rasmussen2014}],
 motor control [\citealp{dewolf2011}],
@@
 response timing \citep{stewart2009a},
-and behavioral errors \citep{choo2010}
+behavioral errors \citep{choo2010},
+and age-related cognitive decline \citep{rasmussen2014},
 of NEF-designed models match
-psychological and physiological findings
+physiological and psychological findings
 without being built specifically into the design.
@@
 
-It has often been said that neuroscience
-is data-rich and theory-poor \citep{churchland1992}.
-We believe that the modeling success of
-large functional networks such as Spaun
-suggests that the NEF
-may be able to fill the theoretical void.
-However, in simulating this large model,
-the Java-powered Nengo simulator
-was limited in terms of memory consumption,
-simulation speed, and extensibility.
-In order to continue to evaluate the NEF
-as a theory of neural computation
-by building larger and more complicated models,
-we have begun the next generation
-of Nengo development with a focus on speed,
-extensibility, and simplicity.
-
-This paper introduces a new version
-of Nengo (PyNengo) written in Python.
-The remainder of this introduction outlines
-the design issues in the previous version (JavaNengo)
-that motivated the reimplementation.
-PyNengo is not currently meant
-to completely replace JavaNengo;
-JavaNengo remains a capable visual NEF
-simulator and a valuable teaching tool
-due to its graphical interface.
-PyNengo's focus is on speed
-and a simplified internal design
-to support ongoing research in
-large adaptive functional models.
-The remainder of the paper outlines PyNengo's design,
-scripting interface, and performance relative to
-JavaNengo and alternatives such as Brian and NEST.
-
-\subsection{JavaNengo limitations}
-
-\begin{table}[!t]
-\processtable{Tools bearing the name Nengo.\label{tab:nengo-vers}}
-{\begin{tabular}{p{2.6cm} p{2.7cm} l p{7cm}} \toprule
-\textbf{Name} & \textbf{Implementation language}
-  & \textbf{Code repository} & \textbf{Purpose} \\\midrule
-JavaNengo simulator & Java & \texttt{nengo\_java}
-  & Simulates many types of spiking neuron models.
-  Includes methods to construct and simulate models using the NEF. \\
-JavaNengo GUI & Java & \texttt{nengo\_java\_gui}
-  & Graphical interface to JavaNengo simulator.
-  Allows modellers to drag-and-drop components built with the NEF. \\
-JavaNengo scripting interface & Jython & \texttt{nengo\_jython}
-  & Scripting interface to JavaNengo simulator.
-  Provides shortcuts to classes and functions in the simulator.
-  Also accessible through and interacts with the JavaNengo GUI. \\
-PyNengo & Python & \texttt{nengo}
-  & Constructs and simulates models using the NEF. \\
-PyNengo OpenCL simulator & Python+PyOpenCL
-  & \texttt{nengo\_ocl}
-  & Simulates models constructed with PyNengo. \\\botrule
-\end{tabular}}{Code repositories avilable at \url{http://github.com/ctn-waterloo/}.}
-\end{table}
-
-JavaNengo is a suite of tools composed of
-a neural simulator and graphical user interface (GUI)
-implemented in Java,
-and a scripting interface implemented in Jython
-(as detailed in Table~\ref{tab:nengo-vers}).
-JavaNengo's simulator was created
-with the intention of being a general
-neural simulator that included
-the methods of the NEF
-as one of several
-methods for creating neural models.
-As a result, its implementation
-of important NEF components
-are nested in deep class hierarchies,
-adding unnecessary layers of complexity
-for developers wishing
-to extend those NEF components.
-For example,
-the most often used class for making
-NEF models, \texttt{NEFEnsemble},
-extends from five ancestor classes
-and implements seven interfaces.
-Adding support for synaptic plasticity
-in \texttt{NEFEnsemble}
-involved adding or significantly modifying
-25 Java classes \citep{bekolay2010}.
-
-The JavaNengo GUI provides
-a drag-and-drop interface in which
-NEF models can be created.
-However, as it was developed
-before the scripting interface,
-the GUI manipulates simulator objects directly,
-rather than constructing a script
-that manipulates those simulator objects.
-Typically, modellers start by following
-tutorials that use the GUI interface,
-and transition to the scripting interface
-in order to implement more complex models;
-since the GUI modifies simulator objects directly,
-modelling concepts learned in the GUI
-have to be relearned in the scripting interface.
-
-The JavaNengo scripting interface,
-described in \citet{stewart2009},
-provides a simpler interface
-to the NEF-specific portions of the simulator.
-This greatly enhanced productivity,
-allowing models to be described
-by short scripts.
-Modellers quickly started using
-the scripting interface
-as their primary tool,
-but still needed to access the GUI
-for interactive model inspection,
-and still needed to modify the simulator
-to add new functionality
-(e.g., to add new neuron models).
-This is because Jython makes it easy
-to inspect and create Java objects,
-but does not make it easy to
-write Java code to
-inspect and create Jython objects.
-Modellers, therefore, had to be
-proficient in Java whether
-they used the scripting interface or not.
-As there are few neural modellers
-that are proficient with both Java and Python syntax,
-the scripting interface still lacks
-convenient ways to record the results of simulations,
-and to define complex experimental setups.
-Additionally, Jython is always several versions
-behind CPython, making it impossible
-to use new language features
-like dictionary comprehensions.
-Jython also cannot access libraries
-that contain any C code,
-including all scientific libraries
-that use NumPy.
-
-Finally, perhaps the most significant limitation
-of JavaNengo is its performance.
-While Java has good support for threads,
-it lacks native high-performance numerical libraries
-for doing mathematical operations on large matrices.
-Additionally, it is difficult to interface with
-hardware such as graphical processing units (GPUs),
-which are well-suited to the computations
-required in neural simulations.
-
-Addressing all of these issues
-in the Java codebase would require
-enough extensive changes
-that we opted instead
-to write a new version of Nengo in Python.
-JavaNengo's Jython scripting interface
-already used Python syntax,
-and since Python is widely used
-for neural simulation and data analysis,
-we targeted the CPython interpreter
-as the platform for the new version of Nengo.
-
-\subsection{PyNengo}
-
-PyNengo is a Python package for
-defining and simulating
-neural models using the NEF.
-Its interfaces are based on NumPy,
-so it integrates seamlessly
-with other scientific Python tools
-for interactive exploration \citep{perez2007},
-scientific computation \citep{oliphant2007},
-and plotting \citep{hunter2007}.
-Arbitrary CPython programs
-can use Nengo models,
+The transformation principle of the NEF
+proposes that the connection weight matrix
+between two neural populations
+can compute a nonlinear function,
+and can be factored into
+two significantly smaller matrices.
+By using these factors
+instead of full connection weight matrices,
+NEF-designed models are
+more computationally efficient,
+which allows Nengo to run
+large-scale neural models
+on low-cost commodity hardware.
+
+In order to make Nengo more simple, extensible, and fast,
+we have rewritten Nengo 2.0 from scratch in Python,
+leveraging NumPy \citep{oliphant2007}
+for manipulating large amounts of data.
+While NumPy is its only dependency,
+Nengo contains optional extensions
+for plotting if Matplotlib is available
+\citep{hunter2007}
+and for interactive exploration
+if IPython is available \citep{perez2007}.
+Since Nengo only depends on one third-party library,
+it is easy to integrate Nengo models
+in arbitrary CPython programs,
 opening up possibilities
@@
 
-PyNengo has a simple object model,
+Nengo 2.0 has a simple object model,
 which makes it easy to
 document, test, and modify.
-These objects are accessible
-through a scripting interface
-similar to JavaNengo's \citep{stewart2009},
-with improvements
-for easier incorporation of user-defined networks,
-neuron models and learning rules.
-Critically, and unlike JavaNengo,
-PyNengo's scripting interface
-contains functions to record simulated data
-and access that data after simulation.
-
 Model creation and simulation are decoupled,
-allowing for models to be exported
-to other simulation packages
-(e.g., those accessible by PyNN [\citealp{davison2008}])
-or for faster simulators to be used
-as drop-in replacements for PyNengo's
+allowing for models to be run with
+other simulators as drop-in replacements for Nengo 2.0's
 platform-independent reference simulator.
@@
 of a GPU or multicore CPU.
-
-PyNengo does not yet
-replace all of JavaNengo's functionality.
-While it can simulate most
-models that JavaNengo can simulate,
-there is currently no graphical interface for PyNengo.
-Model components must be created
-using Python scripts rather than
-the drag-and-drop interface presented by JavaNengo.
-Some of the means of interacting
-with a running model in JavaNengo
-can be provided using packages
-such as IPython and Matplotlib,
-but such tools have not been our focus thus far.
-It is possible to build a graphical program
-like JavaNengo's GUI on top of PyNengo's object model,
-but that remains as future work.
-
-To summarize, PyNengo overcomes
-JavaNengo's limitations.
-It can simulate most models
-currently simulated by JavaNengo
-using a well-tested codebase of
-$\sim$5000 lines of code,
-compared to JavaNengo's $\sim$45,000
-(which does not include the JavaNengo GUI).
-A new internal representation of neural models
-allows the OpenCL simulator to simulate
+The OpenCL simulator can simulate
 large models on the scale of Spaun
-hundreds of times faster than JavaNengo
+at least 50 times faster than Nengo 1.4
 using inexpensive commodity hardware.
-Compatibility with CPython makes scientific Python software
-available for preparing and analyzing data.
-In all, PyNengo provides a platform for
+
+In all, Nengo 2.0 provides a platform for
 simulating larger and more complex models than Spaun,
@@
 
-\section{Neural Engineering Framework (NEF)}
+\section{Neural Engineering Framework (NEF)} \label{sec:nef}
 
@@
 that causes the neuron to be the most active).
+Importantly, tuning curves can be determined
+for any type of neuron,
+and therefore the encoding process
+(and the NEF as a whole)
+is not dependent on any particular neuron model.
 
@@
 
-\paragraph{Transformation}  % NEF as neural model...
+In Nengo, the representation principle
+can be seen in the \texttt{Ensemble}
+object (see Section~\ref{sec:ensemble}).
+
+\paragraph{Transformation}
 Neurons communicate through
@@
 
+In Nengo, the transformation principle
+can bee seen in the \texttt{Connection}
+object (see Section~\ref{sec:connection}).
+
 \paragraph{Dynamics}
@@
 
+In Nengo, dynamics can be seen when
+an \texttt{Ensemble} is connected to itself.
+Several of the \texttt{Network}s
+implemented in Nengo also exhibit dynamics.
+
 \paragraph{NEF and Nengo}
@@
 modellers to create and connect those components.
-Components describe
+Ensembles describe
 what information is being represented,
@@
 how that information is transformed.
-Nengo translates those descriptions
+Nengo implements those descriptions
+with its object model,
+and translates those objects
 to a network of interconnected neurons,
@@
 
-\section{PyNengo object model}
+\section{Nengo object model}
 
 To describe an NEF model,
-PyNengo defines six core objects.
+Nengo defines six core objects.
 \begin{enumerate}
-  \item The \texttt{Model} encapsulates the entire model and
-    exposes a simplified interface to other objects.
   \item The \texttt{Ensemble} contains a group of neurons
@@
     nodes and ensembles are connected.
+  \item The \texttt{Probe} gathers data during a simulation
+    for later analysis.
   \item The \texttt{Network} encapsulates a functionally related
     group of interconnected nodes and ensembles.
-  \item The \texttt{Probe} gathers data during a simulation
-    for later analysis.
+  \item The \texttt{Model} encapsulates a Nengo model.
 \end{enumerate}
@@
 
-\subsection{Model}
-
-The \texttt{Model} object is primarily a container
-for Nengo objects,
-but also doubles as a scripting interface
-that simplifies creating, connecting,
-and probing objects.
-Basic models
-can be created solely through interacting
-with the \texttt{Model} object
-(see sections \ref{sec:comm-channel} \& \ref{sec:lorenz}
-for examples).
-In advanced models,
-objects are instantiated
-and then added to the model
-(see section \ref{sec:cconv} for an example).
-
-\subsection{Ensemble}
+\subsection{Ensemble} \label{sec:ensemble}
 
@@
 \begin{quote}
-  \texttt{Ensemble('A', nengo.LIF(50, tau\_ref=0.002), 1)}
+  \texttt{nengo.Ensemble(nengo.LIF(50, tau\_ref=0.002), 1)}
 \end{quote}
-describes an ensemble, `A',
+describes an ensemble
 that uses 50 leaky integrate-and-fire neurons \citep{lapicque1907}
@@
 such as its encoding weights,
-the maximum firing rate of its neurons,
-and so on, can be specified
+can be specified
 either as keyword arguments
@@
 neurobiological constraints
-by selecting neural parameters
-from distributions consistent with
+by using default parameters
+consistent with
 neocortical pyramidal cells.
@@
 
-A node represents information
+Generally, a node represents information
 that cannot be decoded from an ensemble.
@@
 
-\subsection{Connection}
+\subsection{Connection} \label{sec:connection}
 
@@
 
-The most important type of connection
-in Nengo is the \texttt{DecodedConnection}.
-This connection implements
+When an ensemble is connected
+to another object,
+the connection implements
 the NEF's transformation principle.
-In other words, the \texttt{DecodedConnection}
+In other words, the \texttt{Connection}
 allows ensembles to project
@@
 to synaptic weights.
-By default, ensembles are connected
-with a \texttt{DecodedConnection}.
 
@@
 with synaptic connection weights
-using the \texttt{NonlinearityConnection}.
-Both connection types
-can be temporally filtered,
+by connecting an ensemble's
+neurons directly to another ensemble's neurons
+(e.g., \texttt{nengo.Connection(ens1.neurons, ens2.neurons, ...)}).
+All connections can be temporally filtered,
 and the weights involved in the connection
-(decoding weights for \texttt{DecodedConnection},
-synaptic connection weights for \texttt{NonlinearityConnection})
 can be modified over time with learning rules.
-
-\subsection{Network}
-
-A network is a collection of interconnected ensembles and nodes.
-It is primarily a way of grouping together
-a set of connected objects
-that collectively perform a complex function.
-Encapsulating them in a network
-makes its purpose explicit
-and hides the complexity of the function
-(see section \ref{sec:cconv} for an example).
-\texttt{Network} is a base class designed to be
-subclassed by modellers.
-The code that creates and connects
-several objects in a model can be
-grouped into a \texttt{Network} subclass
-with only minor changes.
-
-Networks can vary dramatically in size,
-and can be composed of other networks.
-As a simple example, the \texttt{Integrator} network
-is composed of only one recurrently connected ensemble.
-By encapsulating that logic in a network,
-the purpose of that ensemble is made explicit.
-As a complex example,
-the \texttt{BasalGanglia} network
-is composed of five groups of ensembles
-connected with several specific functions
-that together implement a ``winner-take-all'' circuit.
-Encapsulating the code to create those ensembles
-and connections in a network subclass
-makes a complicated section of code
-easy to include in many different models.
 
@@
 Like nodes, a probe could be implemented
-outside of the model.
+outside of the neural model.
 However, doing so requires detailed knowledge
@@
 
-\section{PyNengo simulators} \label{sec:simulators}
+\subsection{Network}
+
+A network is a collection of interconnected ensembles and nodes.
+Networks provide a way of grouping together
+a set of connected objects
+that collectively perform a complex function.
+Encapsulating them in a network
+makes its purpose explicit
+and hides the complexity of the function
+(see Section~\ref{sec:cconv} for an example).
+This grouping can be hierarchical,
+as networks can contain other network.
+\texttt{Network} is a base class designed to be
+subclassed by modellers.
+The code that creates and connects
+several objects in a model can be
+grouped into a \texttt{Network} subclass
+with only minor changes.
+Nengo comes with several networks already implemented
+which can be used directly,
+or can be used as a template
+for modelers wanting to implement their own networks.
+
+As a simple example, the \texttt{Integrator} network
+is composed of only one recurrently connected ensemble.
+By encapsulating that logic in a network,
+the purpose of that ensemble is made explicit.
+As a complex example,
+the \texttt{BasalGanglia} network
+is composed of five groups of ensembles
+connected with several specific functions
+that together implement a ``winner-take-all'' circuit.
+Encapsulating the code to create those ensembles
+and connections in a network
+makes a complicated section of code
+easy to include in many different models.
+
+\subsection{Model}
+
+The \texttt{Model} object is a container
+for Nengo objects.
+Conceptually, they are similar to networks,
+but are privileged in that
+simulators must have a model
+passed into their constructor,
+making the \texttt{Model} akin
+to a network that contains
+all of the objects defined for a Nengo model.
+A simulator's job is to simulate a model.
+
+\section{Nengo simulators} \label{sec:simulators}
 
@@
 decoupling neural model creation and simulation,
-which enables PyNengo simulators
+which enables Nengo simulators
 to allocate memory and schedule computations
@@
 as an argument;
-this \texttt{Model} is a symbolic description,
-which a simulator can assume will not change.
-The simulator, however,
-can modify the model as it sees fit;
-in most cases, this means that the simulator
-will fill in many of the details
-not specified at the symbolic level.
-
-We have implemented
-a reference simulator using NumPy
-for vectorized computations,
-and an OpenCL simulator
-using PyOpenCL to parallelize
+this \texttt{Model} is a static symbolic description.
+The simulator can take the model description
+and build whatever data structures
+are best suited to implement the simulation.
+
+We have implemented a platform-independent
+reference simulator as an example
+for simulator designers.
+This simulator is not a specification;
+any object that accepts a Nengo \texttt{Model}
+as an argument is considered a Nengo simulator.
+To show that model creation and simulation
+are fully decoupled,
+we have also implemented
+an OpenCL simulator
+that uses PyOpenCL to parallelize
 computations on GPUs and multicore CPUs.
-In the remainder of this section,
+However, in the remainder of this section,
 we will describe
@@
 that are beyond the scope of this paper.
-Additionally, it should be noted that
-any simulator that takes a \texttt{Model}
-as input and exposes a \texttt{step}
-function can be used in PyNengo;
+
+\subsection{Nengo reference simulator}
+
+The Nengo reference simulator
+makes a copy of the objects in the model
+and fills in many of the details
+not specified at the symbolic level.
+For example, \textit{encoders}
+are often not specified when the model is created,
+so the reference simulator randomly chooses
+them as unit vectors in the space that
+the ensemble is representing.
+After filling in these details,
 the reference simulator
-is provided as an example
-for simulator designers,
-not as a specification.
-
-The reference simulator
-uses a reduced set of objects
-that describe computations
+builds a reduced set of objects
+that describe the computations
 occurring in the model.
 Specifically, the simulator
-contains signals, which represent values,
+uses signals, which represent values,
 and operators, which represent computations
@@
 will be used by the simulator.
-Each high-level Nengo object contains
-several signals;
-for example, an ensemble contains signals
+Several signals are created
+for each high-level Nengo object;
+for example, for each ensemble,
+the simulator creates signals
 that represent the high-level input
@@
 and the encoding weights.
-It also contains a neural population,
-which contains signals that represent
+The ensemble also contains a neural population,
+for which the simulator creates signals that represent
 input currents, bias currents,
@@
 and therefore can share a common operator;
-this can be helpful for parallelizing computations.
+this is helpful for parallelizing computations.
 Nonlinear functions, however,
@@
 Each supported neuron model and learning rule
-has an associated operator;
-the simulator explicitly maps
+has an associated operator.
+The simulator explicitly maps
 from symbolic neuron objects in ensembles
 and from symbolic learning rule objects
-in connections to their associated operator.
+in connections to their associated operators.
 
@@
 \begin{enumerate}
+  \item fills in unspecified details of high-level objects,
   \item translates high-level objects to
@@
 The scripting interface provides
-a simple way to add PyNengo objects to a model,
+a simple way to add Nengo objects to a model,
 simulate that model,
@@
 that highlight the types of models that can be
-built with PyNengo.
+built with Nengo.
 We have also implemented two of these three examples
@@
 
-A communication channel
-is a simple example demonstrating
-the representation and transformation principles
-of the NEF.
-A communication channel
-represents some signal and transmits it unchanged;
-i.e., it computes the function $f(x) = x$.
+% A communication channel represents some signal
+% and transmits it unchanged to another population
+% (i.e. it implements an identify function $f(x) = x$).
+% One of the most important operations
+% in large neural models is routing with this sort of channel.
+% When connected recurrently,
+% a communication channel can also be seen as
+% a simple memory, mapping a real-valued number
+% $x$ to itself at a future time [awkward wording].
+% In this simple case,
+% we can see how the principles of the NEF
+% help to translate a mathematical description
+% of our model to neural circuitry.
+% An Ensemble embodies the idea
+% that a population of neurons
+% should represent a real-valued quantity
+% $x \in \mathbb{R}^1$.
+% Technically, Nengo does this
+% by creating a random projection matrix
+% $E$ and driving the first population
+% with current $Ex + b$,
+% where $b$ is a vector of bias values.
+% A vector $s$ of spikes emitted
+% by this population are weighted and combined
+% by a {\em decoder} matrix D
+% so that $Ds \approx f(x)$.
+% Similarly the second population
+% has encoders $E_2$ and $D_2$.
+% A full weight matrix $W$
+% of synapse strengths
+% between populations 1 and 2
+% can be obtained
+% by multiplying matrices $E_2$ and $D_1$ together:
+% $W=E_2 D_1$.
+% We ran simulations in PyNN
+% by computing this product explicitly
+% and using $W$ and the connection matrix.
+% Unfortunately it much slower to multiply
+% $s$ by $W$ than it is to compute $E_2(D_1s)$,
+% which is why Nengo
+% is a faster simulator for NEF-style models.
+
+As detailed in Section~\ref{sec:nef},
+NEF models are based on the principles
+of representation, transformation, and dynamics.
+One of the most important operations
+in a large neural model is to route
+represented information from
+one ensemble to another without any change.
+For example, in the visual system of Spaun,
+a high-dimensional representation of the visual field
+is compressed to a low-dimensional representation,
+and then sent unchanged to several areas,
+including the working memory
+and action selection networks.
+This routing is implemented
+with a transformation
+called a communication channel.
+This transform simply implements
+the identity function, $f(x) = x$.
+
 Figure~\ref{fig:comm-channel}
 depicts a scalar communication channel
-representing band-limited Gaussian white noise.
+in which band-limited Gaussian white noise
+is represented in one ensemble
+and projected to another ensemble.
 
@@
 for implementing a communication channel in Nengo and PyNN.
-This highlights many of the differences
+This figure highlights many of the differences
 between Nengo models and conventional neural models;
 we also use these script for benchmarking
-(see section \ref{sec:benchmark}).
+(see Section~\ref{sec:benchmark}).
 
@@
 
+While the communication channel exemplifies
+the representation and transformation
+principles of the NEF,
+the Lorenz attractor exemplifies the dynamics principle.
 Many models in theoretical neuroscience
@@
 to implement many different types of
-attractor networks \citep{eliasmith2005}.
+attractor networks by recurrently connecting
+ensembles with functions that implement
+dynamical systems \citep{eliasmith2005}.
 Figure~\ref{fig:lorenz} depicts
@@
 
-A common function computed by Nengo models
-is circular convolution.
-This function is a central part
-of the Semantic Pointer Architecture \citep{eliasmith2013}
-that uses the NEF to build
-a variety of cognitive models including Spaun.
+Communication channels and attractor networks
+show up in many Nengo models,
+but are still relatively simple
+to implement without Nengo,
+as can be seen with the PyNN implementation
+in Figure~\ref{fig:pynn}.
+As the NEF has been used to construct
+larger models that have the capabilities
+of non-neural cognitive models,
+a theory called the
+Semantic Pointer Architecture \citep{eliasmith2013}
+has emerged.
+This theory uses high-dimensional vectors
+as symbol-like structures
+that be combined together to form
+novel concepts.
+
+One of the functions that
+is performed on these vectors
+is to compress two $n$-dimensional vectors
+into a single $n$-dimensional vector,
+which can be decompressed
+into noisy versions of
+the two originally compressed vectors.
+We implement this compression
+using the circular convolution function.
 Circular convolution
 is best implemented in a two-layer network,
-rather than in a single connection.
-This two-layer organization is simplified in
-the \texttt{CircularConvolution} network subclass.
+rather than in a single connection,
+which we have simplified through
+the \texttt{CircularConvolution} network.
 The complexity encapsulated in that network
@@
 increasing simulation speed
-was a primary motivator in creating PyNengo.
+was a primary goal of Nengo 2.0.
 To validate that performance has improved,
-we ran the models described in section~\ref{sec:examples}
+we ran the models described in Section~\ref{sec:examples}
 for various numbers of neurons and dimensions
@@
 on the same machine,
-and plot the mean time elapsed.
+and plot the mean time elapsed
+in Figure~\ref{fig:benchmarks}.
 In most cases, the coefficient of variation
@@
 run on multiple cores using message passing,
-the reference simulator of PyNengo
+the reference simulator of Nengo
 also only uses one CPU core.
-The results also suggest that PyNengo's
-simulators are faster than JavaNengo's simulator.
+The results further suggest that Nengo 2.0's
+simulators are faster than Nengo 1.4's simulator.
 
@@
 implement circular convolution.
-In this case, only versions of Nengo
-were tested.
+In this case, only versions of Nengo were tested.
 Instead of running each simulation multiple times,
@@
 for large models, the OpenCL simulator
-performs much faster than JavaNengo;
+performs much faster than Nengo 1.4;
 in particular, a Radeon 7970 GPU performs
@@
 faster than real time,
-and 50 times faster than JavaNengo.
+and 50 times faster than Nengo 1.4.
 In the 50-dimensional case,
 the Radeon 7970 GPU is 200 times faster
-than JavaNengo.
-Additionally, although both JavaNengo
-and the OpenCL simulator on a CPU
+than Nengo 1.4.
+Additionally, although both Nengo 1.4
+and the OpenCL simulator on CPUs
 use all available CPU cores,
-PyNengo's OpenCL simulator is significantly faster.
+Nengo's OpenCL simulator is many times faster.
 
@@
 There are many other neural simulators
-dedicated to building large-scale neural models;
-however, Nengo is unique in being built
+dedicated to building large-scale neural models
+\citep{eppler2008,goodman2009,hines2009},
+and many tools for simulating cognitive phenomena
+with various levels of biologically plausibility
+\citep{cooper1998,sun2001,anderson2004,franklin2007,
+  aisa2008,dekamps2008,laird2012}.
+However, Nengo is unique in being built
 on a theoretical framework
-that has been validated
+that has enabled a cognitive architecture
+(the Semantic Pointer Architecture)
+that maintains a high level
+of biological plausibility,
+and has been validated
 through the Spaun model and other past work.
 
-The most closely related project is PyNN
-\citep{davison2008},
-which provides a high-level scripting
-interface similar to the
-high-level object model in PyNengo.
-Rather than implementing its
-own simulator, PyNN uses four existing
-simulators with Python bindings,
-such as NEST.
+The most closely related projects
+in terms of software design
+are PyNN \citep{davison2008}
+and Topographica \citep{bednar2009},
+both of which provide a high-level scripting
+interface to low-level neural simulators.
+PyNN in particular has a similar to the
+high-level object model in Nengo,
+and provides a convenient interface
+to the three most widely used
+neural simulators, according
+to a survey by \citet{hanke2011}.
 
@@
 two ensembles;
-e.g., \texttt{DecodedConnection(A, B,
+e.g., \texttt{nengo.Connection(A, B,
 function=square)} connects ensemble A
@@
 
-As shown in Figure~\ref{fig:benchmarks},
-Nengo is able to simulate many models
-faster than each of PyNN's simulators.
+On the neural simulator side,
+we have shown that both Nengo's
+reference simulator and OpenCL simulator
+are able to simulate
+two benchmark models much faster
+than Brian, NEST and NEURON
+(see Figure~\ref{fig:benchmarks}).
 This is, in part,
@@
 rather than storing the entire matrix.
-However, PyNN is able to simulate
+However, these simulators are able to simulate
 many detailed neuron models
 and learning rules,
-and has access to the wealth of models
-already created for each of its
-accessible simulators.
-Because Nengo is in an earlier development stage,
+and have access to
+a wealth of existing neuron models
+and learning rules.
+Because Nengo 2.0 is in an earlier development stage,
 many of these detailed neuron models
 and learning rules remain to be added.
-PyNN is therefore currently better suited to
-simulate detailed neural models.
+Neural simulators like Brian, NEST,
+and NEURON are therefore currently better suited for
+simulating a wider range of single cell models,
+while Nengo is designed for large networks
+of simple neural models
+that are connected together according to
+the principles of the Neural Engineering Framework,
+and can simulate these types of models efficiently.
+
+One key difference between Nengo's simulators
+and traditional neural simulators
+is the target platform.
+While NEST and NEURON
+can be run on commodity hardware,
+networks of modest size are
+typically simulated in high-performance computing
+environments by using
+the Message-Passing Interface (MPI).
+Nengo enables large-scale neural simulation
+on commodity hardware,
+allowing researchers and hobbyists
+without access to large computing clusters
+the ability to experiment with
+theoretical neuroscience
+as is currently practiced
+in cutting edge research.
+In particular, GPUs are a
+powerful, low-cost computing resource
+that are available in most
+modern workstations.
+The OpenCL simulator makes full use
+of GPUs, while the previously discussed simulators
+currently do not.
 
@@
 
-PyNengo is a young project that
-has started with a deliberately minimal base
-in order to make development easier than in
-previous implementations.
-Additionally, many neural modellers
-are already using Python,
-as evidenced by the existence of
-PyNN, Brian, and Python bindings to other simulators.
-Nengo will be able to benefit
-from improvements to scientific Python tools.
-It is also extendable by modellers with
-varying technical skills;
-creating a new network is simple,
-while creating a new simulator is complex, but possible.
-By enabling outside contributions,
-and providing a user-friendly API
-and many example models,
-we hope to foster a rich community of modellers
-contributing networks, neuron types, learning rules,
-and simulators to Nengo.
-Internally, our short-term goal
-is to implement the JavaNengo use cases
-not currently covered by PyNengo.
+Our short-term goal
+is to implement the Nengo 1.4 use cases
+not currently covered by Nengo 2.0.
+While we have ensured that all of the models
+currently used by Nengo 1.4 tutorials
+can be run in Nengo 2.0,
+several large models, like Spaun,
+include custom extensions written in Java.
+We will incorporate useful extensions
+in Nengo's API directly,
+and reimplement more specific extensions
+to use Nengo's API.
+
+We are also developing two simulators
+that will take the same NEF model description
+as the existing simulators,
+but will target two pieces
+of neuromorphic hardware
+to achieve greater speed and power efficiency
+than the OpenCL simulator.
+
 Our long-term goal is to
@@
 interactively inspect simulations.
+Nengo 1.4 includes a graphical interface
+that includes an interactive simulation inspector.
+We will use our experience building that interface
+to construct an improved interface
+for Nengo 2.0.
+
+Additionally, we hope that in the future
+some work in Nengo will be
+done through outside contributions.
+Nengo 2.0 is a complete rewrite that
+has started with a deliberately minimal base
+and a well-defined API
+in order to make development easier than in
+previous versions.
 
@@
 
-PyNengo is the next generation of Nengo.
-Though the project is still young,
-it can already simulate most models
-that have been built using the NEF.
+Nengo 2.0 is the next generation of Nengo.
+It has been rewritten from scratch,
+but can already simulate most models
+that have been built using Nengo 1.4.
 It does this with 11\% as many lines of code
@@
 it can simulate circular convolution models
-50-200 times faster than JavaNengo,
+50-200 times faster than Nengo 1.4,
 which itself is faster than alternative simulators
@@
 the NEF as a theory of neural computation;
-PyNengo makes those models
+Nengo makes those models
 accessible to anyone with
 a modern computer.
+
+\appendix
+\section*{Appendix: Neural Engineering Framework details}
+
+This description of the Neural Engineering Framework
+is adapted from the supplementary material
+of \citep{eliasmith2012} with permission.
+
+The Neural Engineering Framework (NEF; \citealp{eliasmith2003})
+provides a set of methods
+for building biologically plausible models
+based on a functional specification of a neural system.
+The central idea behind the NEF
+is that a group of spiking neurons
+can represent a vector space over time,
+and that connections between
+groups of neurons
+can compute functions on those vectors.
+The NEF provides a set of methods
+for determining what the connections
+need to be to compute a given function
+on the vector space
+represented by a group of neurons.
+
+Suppose we wish to compute the function
+$\mathbf{y}=f(\mathbf{x})$,
+where vector space $\mathbf{x}$
+is represented in population A,
+and vector space $\mathbf{y}$
+is represented in population B.
+To do so, the NEF assumes that
+each neuron in A and B has a
+``preferred direction'' or ``encoding'' vector.
+The preferred direction vector
+is the vector (i.e. direction in the vector space)
+for which that neuron will fire most strongly.
+The spiking activity of every neuron
+in population A can be written as
+\begin{equation} \label{eq:enc}
+  a_{i}(\mathbf{x})=G_{i}[\alpha_{i}\mathbf{e}_{i}\mathbf{x}+J_{i}^{bias}],
+\end{equation}
+where $a_{i}$ is the spike train
+of the $i$th neuron in the population,
+$G$ is the spiking neural nonlinearity,
+$\alpha$ is the gain of the neuron,
+$\mathbf{e}$ is the preferred direction
+(or ``encoding'') vector,
+and $J^{bias}$ is a bias current
+to account for background activity of the neuron.
+Notably, the elements in the square brackets
+determine the current flowing into the cell,
+which then drives the spiking
+of the chosen single cell model $G$.
+Equation \eqref{eq:enc} describes
+how a vector space
+is encoded into neural spikes.
+This equation is depicted
+for a 1-dimensional vector space
+in Figure~\ref{fig:nef}A.
+
+The NEF proposes that linear decoders
+can be found to provide an appropriate
+estimate of any vector $\mathbf{x}$
+given the neural activities from
+the encoding equation.
+We can write this as a decoding equation:
+\begin{equation} \label{eq:dec}
+  \hat{\mathbf{x}}=\sum_{i}^{N}a_{i}\mathbf{(x)}\mathbf{d}_{i},
+\end{equation}
+where $N$ is the number of neurons in the group,
+$\mathbf{d}_{i}$ are the linear decoders,
+and $\hat{\mathbf{x}}$ is the estimate of
+the input driving the neurons.
+
+The NEF determines this complementary decoding
+for any given encoding.
+Specifically, this decoding is found
+using a least-squares optimization:
+\begin{equation} \label{eq:error}
+  E=\frac{1}{2}\int[\mathbf{x}-\sum_{i}a_{i}(\mathbf{x)}\mathbf{d}_{i}]^{2}d\mathbf{x},
+\end{equation}
+where $\mathbf{d}_{i}$ are the decoding vectors
+over which this error is minimized.
+
+In effect, this optimization process
+replaces learning in most other
+approaches to constructing neural networks.
+This optimization is not
+biologically plausible on its own,
+although networks generated in
+this manner can also be learned
+with a spike-based rule \cite{macneil2011}.
+
+The decoding process is depicted
+in Figure~\ref{fig:nef}B,
+where the optimal linear decoders
+have been found and used for eight neurons.
+Notably, this kind of temporal decoding
+requires an assumption about
+the nature of the temporal filter being used.
+Here we assume that post-synaptic currents
+are such filters, and set the time constants
+to reflect the kind of neurotransmitter receptors
+in the connection
+(e.g., AMPA receptors have short time constants, $\sim10$ ms,
+and NMDA receptors have longer time constants, $\sim50$ ms).
+
+Such temporal filters map
+to biophysical processes
+once we connect groups of neurons together.
+Defining the encoding and decoding for
+groups A and B using equations
+\eqref{eq:enc} and \eqref{eq:dec}
+provides a means of connecting groups.
+For example, we can substitute
+the decoding of A into the encoding of B,
+thereby deriving connection weights
+\begin{equation} \label{eq:weights}
+  \omega_{ij}=\mathbf{d}_{i}\alpha_{j}\mathbf{e}_{j},
+\end{equation}
+where $i$ indexes the neurons in group A
+and $j$ indexes the neurons in B.
+These weights will compute the function
+$\mathbf{y}=\mathbf{x}$
+(where $\mathbf{y}$ is the vector space
+represented in B and $\mathbf{x}$
+is the vector space represented in A).
+For the more general case,
+it is possible to solve for decoders
+$\mathbf{d}_{i}^{f}$
+for any function by substituting
+$f(\mathbf{x})$ for $\mathbf{x}$
+in equation \eqref{eq:error}, i.e., solving
+\begin{equation} \label{eq:error-fcns}
+E=\frac{1}{2}\int[f(\mathbf{x})-\sum_{i}a_{i}(\mathbf{x)}\mathbf{d}_{i}^{f}]^{2}d\mathbf{x}.
+\end{equation}
+
+In addition, if the function to be computed is linear,
+the relevant linear operator can be introduced
+into Equation \eqref{eq:weights}.
+The resulting general weight equation
+for computing any combination
+of linear and nonlinear functions becomes:
+\begin{equation} \label{eq:weights-general}
+  \omega_{ij}=\alpha_{j}\mathbf{d}_{i}^{f}\mathbf{L}\mathbf{e}_{j}
+\end{equation}
+for any nonlinear function $f$ and
+$N_{B}\times N_{A}$ linear operator $\mathbf{L}$.
+Computing the linear function
+$\mathbf{y}=\mathbf{-x}$ and
+computing the nonlinear function
+which is the element-wise
+square of the vector $\mathbf{x}$
+(i.e., $\mathbf{y}=[x_{1}^{2},x_{2}^{2},...,x_{n}^{2}]$
+is shown in Figure~\ref{fig:nef}C.
+
+This brief discussion is insufficient
+to fully introduce the generality of the NEF.
+However, showing how to compute linear
+and nonlinear functions of vector spaces
+is sufficient for many neural computations.
+As these same methods can be used
+to compute connection weights for recurrent connections,
+the NEF also allows for the neural implementation
+of a wide variety of linear and nonlinear
+dynamical systems in recurrent networks.
+A simple harmonic oscillator is shown in Figure~\ref{fig:nef}D.
 
@@
 
-TB led development of the
-PyNengo object model and scripting interface,
+TB led development of
+Nengo's object model and scripting interface,
 wrote the text of the paper,
@@
 JB led development of
-the PyNengo reference simulator and OpenCL simulator,
+the Nengo reference simulator and OpenCL simulator,
 edited text, created an early version
@@
 EH contributed significantly
-to PyNengo and both of its simulators, and edited text.
+to Nengo and both of its simulators, and edited text.
 TD led development
 of a Theano-backed
-version of PyNengo that
+version of Nengo that
 identified issues with Theano,
 and provided the base for
-the version of PyNengo described in this paper.
+the version of Nengo described in this paper.
 TCS contributed
-to PyNengo, helped implement
+to Nengo, helped implement
 the PyNN scripts used in
-sections~\ref{sec:examples} and \ref{sec:benchmark},
-and implemented the JavaNengo scripting interface
-on which PyNengo is based.
+Sections~\ref{sec:examples} and \ref{sec:benchmark},
+and implemented the Nengo 1.4 scripting interface
+on which the Nengo object model is based.
 DR contributed
-to PyNengo and the reference simulator,
+to Nengo and the reference simulator,
 and edited text.
-XC contributed to PyNengo
+XC contributed to Nengo
 and the reference simulator.
@@
 CE oversaw all development,
-contributed to PyNengo,
+contributed to Nengo,
+wrote the NEF appendix,
 and co-created the NEF with Charles Anderson.
@@
 for editing this paper,
-contributing to PyNengo
-and the PyNengo OpenCL simulator
-and for creating the JavaNengo simulator
-that has been useful
+contributing to Nengo
+and the Nengo OpenCL simulator
+and for creating Nengo 1.4,
+which has been indispensable
 for the past six years,
@@
 who have contributed
-to PyNengo by providing examples,
+to Nengo by providing examples,
 unit tests, and bugfixes.
+Finally, we thank the two reviewers
+who have provided constructive
+feedback on this paper.
 
@@
    signals being represented by population of neurons
-   can be thought of as state variables in a dynamical system.
-   }
+   can be thought of as state variables in a dynamical system.}
 \end{figure}
@@
  \textbf{\refstepcounter{figure}\label{fig:sim} Figure \arabic{figure}.}{
-   Detailed breakdown of the PyNengo reference simulator
+   Detailed breakdown of the Nengo reference simulator
    running a simple model for two timesteps.
@@
    implementing the dynamics of the Lorenz attractor.
-   Note that this populations does not receive
+   Note that this population does not receive
    any input that might drive its initial value;
@@
    on the example models described
-   in section~\ref{sec:examples}.
+   in Section~\ref{sec:examples}.
    In (A) and (B), all of the simulators
-   except the PyNengo OpenCL simulator
+   except the Nengo OpenCL simulator
    were run on an Intel Core i7-965.
-   JavaNengo used all 4 cores of this processor;
+   Nengo 1.4 used all 4 cores of this processor;
    all other simulator used only 1 core.
-   The PyNengo OpenCL simulator
+   The Nengo OpenCL simulator
    was run on an NVidia GTX280 GPU.
@@
    at a 1 millisecond timestep.
-   For all model sizes, PyNengo simulators
-   are faster than JavaNengo,
+   For all model sizes, Nengo simulators
+   are faster than Nengo 1.4,
    which is significantly faster than NEURON and NEST,
@@
    The full Brian results are not shown;
-   for the largest model, Brian takes $\sim$768 seconds.
+   for the largest model,
+   the Brian simulation takes $\sim$768 seconds.
    \textbf{(B)} Benchmark results from simulating
@@
    The full results for Brian and NEST are not shown;
-   for the largest model, they take $\sim$1467
-   and $\sim$601 seconds respectively.
+   for the largest model, simulations in Brian and NEST
+   take $\sim$1467 and $\sim$601 seconds, respectively.
    \textbf{(C)} Benchmark results from simulating
@@
    For the blue lines, the simulator used
-   was the PyNengo OpenCL simulator.
-   The CPU used for JavaNengo and
-   the PyNengo reference simulator
+   was the Nengo OpenCL simulator.
+   The CPU used for Nengo 1.4 and
+   the Nengo reference simulator
    was an Intel Core i7-3770;
-   all 4 cores were used by JavaNengo,
-   while PyNengo's reference simulator
+   all 4 cores were used by Nengo 1.4,
+   while Nengo's reference simulator
    only used one core.
@@
    the OpenCL simulator is faster
-   than the JavaNengo simulator,
-   which is faster than the PyNengo
+   than the Nengo 1.4 simulator,
+   which is faster than the Nengo
    reference simulator.
@@
    up to 50 dimensions.
-   The full results for JavaNengo
+   The full results for Nengo 1.4
    are not shown; for the largest model,
-   it takes $\sim$45 seconds.
- }
+   simulation with Nengo 1.4
+   takes $\sim$45 seconds.}
 \end{figure}

### 2013-12-05 scripts/base.py
Updated figures to Nengo object syntax

@@ -2,12 +2,11 @@ import numpy as np
 
 from nengo.helpers import sorted_neurons
 
-def sorted_spikes(sim, pop, iterations=100, every=None):
-    t = sim.data(sim.model.t)
-    indices = sorted_neurons(sim.model.get(pop), iterations=iterations)
+def sorted_spikes(t, ens, spike_data, iterations=100, every=None):
+    indices = sorted_neurons(ens, iterations=iterations)
     if every is not None:
         indices = indices[::every]
-    spikes = [t[sim.data(pop + ".spikes")[:,i] > 0].flatten() for i in indices]
+    spikes = [t[spike_data[:,i] > 0].flatten() for i in indices]
     for ix in xrange(len(spikes)):
         if spikes[ix].shape == (0,):
             spikes[ix] = np.array([-1])

### 2013-12-05 scripts/plot_bench_cchannel.py

@@ -13,36 +13,36 @@ data = OrderedDict()
 data['Brian'] = OrderedDict()
 data['NEURON'] = OrderedDict()
 data['NEST'] = OrderedDict()
-data['JavaNengo'] = OrderedDict()
-data['PyNengo\nOpenCL'] = OrderedDict()
-data['PyNengo\nreference'] = OrderedDict()
+data['Nengo 1.4'] = OrderedDict()
+data['Nengo\nOpenCL'] = OrderedDict()
+data['Nengo\nreference'] = OrderedDict()
 colors = ['#238B45', '#984EA3', '#FF7F00', 'k', '#08519C', '#E41A1C']
 
-data['PyNengo\nOpenCL'][(100, 1)] = [
+data['Nengo\nOpenCL'][(100, 1)] = [
     0.794183015823,0.797194957733,0.798620939255,0.796317100525,0.799592971802]
-data['PyNengo\nOpenCL'][(250, 1)] = [
+data['Nengo\nOpenCL'][(250, 1)] = [
     0.836793899536,0.837684869766,0.840667963028,0.837545871735,0.837010145187]
-data['PyNengo\nOpenCL'][(500, 10)] = [
+data['Nengo\nOpenCL'][(500, 10)] = [
     1.11762309074,1.11406683922,1.11392784119,1.11361694336,1.11517691612]
-data['PyNengo\nOpenCL'][(1000, 50)] = [
+data['Nengo\nOpenCL'][(1000, 50)] = [
     4.37966895103,4.36161804199,4.36030006409,4.36526703835,4.36328911781]
 
-data['PyNengo\nreference'][(100, 1)] = [
+data['Nengo\nreference'][(100, 1)] = [
     1.79912400246,1.72637414932,1.77085995674,1.72308516502,1.75330209732]
-data['PyNengo\nreference'][(250, 1)] = [
+data['Nengo\nreference'][(250, 1)] = [
     1.97227215767,1.98903083801,1.96172499657,2.02151703835,2.0146241188]
-data['PyNengo\nreference'][(500, 10)] = [
+data['Nengo\nreference'][(500, 10)] = [
     2.4279820919,2.41257190704,2.44041395187,2.39347505569,2.40128898621]
-data['PyNengo\nreference'][(1000, 50)] = [
+data['Nengo\nreference'][(1000, 50)] = [
     3.12998580933,3.08235788345,3.18121695518,3.12098288536,3.13230895996]
 
-data['JavaNengo'][(100, 1)] = [
[…]

### 2013-12-05 scripts/plot_bench_cconv.py

@@ -10,8 +10,8 @@ neurons = {
 
 data = OrderedDict()
 
-data['JavaNengo'] = OrderedDict()
-data['PyNengo\nreference'] = OrderedDict()
+data['Nengo 1.4'] = OrderedDict()
+data['Nengo\nreference'] = OrderedDict()
 data['Intel Core\ni7-3770'] = OrderedDict()
 data['Intel Xeon\nE5540'] = OrderedDict()
 data['Intel Xeon\nE5-2620'] = OrderedDict()
@@ -21,13 +21,13 @@ data['ATI Radeon\nHD7970'] = OrderedDict()
 colors = ['k', '#E41A1C', '#023858', '#045A8D', '#0570B0',
           '#3690C0', '#74A9CF', '#A6BDDB']
 
-data['JavaNengo'][100] = 9
-data['JavaNengo'][200] = 18
-data['JavaNengo'][500] = 45
-data['PyNengo\nreference'][5] = 2.62999796867
-data['PyNengo\nreference'][10] = 5.21562504768
-data['PyNengo\nreference'][20] = 10.5295498371
-data['PyNengo\nreference'][50] = 27.0069601536
+data['Nengo 1.4'][100] = 9
+data['Nengo 1.4'][200] = 18
+data['Nengo 1.4'][500] = 45
+data['Nengo\nreference'][5] = 2.62999796867
+data['Nengo\nreference'][10] = 5.21562504768
+data['Nengo\nreference'][20] = 10.5295498371
+data['Nengo\nreference'][50] = 27.0069601536
 data['Intel Core\ni7-3770'][5] = 0.168757915497
 data['Intel Core\ni7-3770'][10] = 0.256566047668
 data['Intel Core\ni7-3770'][20] = 0.482143878937
@@ -93,7 +93,7 @@ for ix, name in enumerate(data.keys()):
     elif "reference" in name:
         yoffset = -0.68
 
-    if name == "JavaNengo":
+    if name == "Nengo 1.4":
         plt.text(n[1] + 12000, data[name].values()[1], name, color=colors[ix],
                  ha='left', va='center', fontsize='large')
     else:

### 2013-12-05 scripts/plot_bench_lorenz.py

@@ -13,36 +13,36 @@ data = OrderedDict()
 data['Brian'] = OrderedDict()
 data['NEST'] = OrderedDict()
 data['NEURON'] = OrderedDict()
-data['JavaNengo'] = OrderedDict()
-data['PyNengo\nreference'] = OrderedDict()
-data['PyNengo\nOpenCL'] = OrderedDict()
+data['Nengo 1.4'] = OrderedDict()
+data['Nengo\nreference'] = OrderedDict()
+data['Nengo\nOpenCL'] = OrderedDict()
 colors = ['#238B45', '#FF7F00', '#984EA3', 'k', '#E41A1C', '#08519C']
 
-data['PyNengo\nOpenCL'][100] = [
+data['Nengo\nOpenCL'][100] = [
     0.823611974716,0.827681064606,0.824057102203,0.826804876328,0.825788021088]
-data['PyNengo\nOpenCL'][500] = [
+data['Nengo\nOpenCL'][500] = [
     1.03150510788,1.03527617455,1.03847002983,1.04004907608,1.08543992043]
-data['PyNengo\nOpenCL'][1000] = [
+data['Nengo\nOpenCL'][1000] = [
     1.27273511887,1.27891302109,1.26637387276,1.27474498749,1.26921701431]
-data['PyNengo\nOpenCL'][2000] = [
+data['Nengo\nOpenCL'][2000] = [
     1.73656892776,1.73761916161,1.72962117195,1.72728586197,1.73043680191]
 
-data['PyNengo\nreference'][100] = [
+data['Nengo\nreference'][100] = [
     1.09255409241,1.09260296822,1.09535098076,1.04782700539,1.09734082222]
-data['PyNengo\nreference'][500] = [
+data['Nengo\nreference'][500] = [
     1.37857794762,1.36459708214,1.35432505608,1.36031198502,1.35320615768]
-data['PyNengo\nreference'][1000] = [
+data['Nengo\nreference'][1000] = [
     1.62529683113,1.61883807182,1.64285898209,1.61444401741,1.63075304031]
-data['PyNengo\nreference'][2000] = [
+data['Nengo\nreference'][2000] = [
     2.17722296715,2.18516111374,2.20153999329,2.20053505898,2.19265985489]
 
-data['JavaNengo'][100] = [
+data['Nengo 1.4'][100] = [
     3.18400001526,3.3599998951,3.22500014305,3.28499984741,3.30400013924]
-data['JavaNengo'][500] = [
[…]

### 2013-12-05 scripts/plot_cconv.py

@@ -13,27 +13,25 @@ a_in = np.asarray([-0.21, 0.5, 0.12, 0.06])
 b_in = np.asarray([-0.18, 0.28, 0.18, -0.52])
 
 model = nengo.Model("Circular convolution")
-model.make_node("Input A", a_in)
-model.make_ensemble("A", nengo.LIF(512), 4)
-model.connect("Input A", "A")
+in_a = nengo.Node(a_in)
+a = nengo.Ensemble(nengo.LIF(512), 4, label="A")
+nengo.Connection(in_a, a)
+in_b = nengo.Node(b_in)
+b = nengo.Ensemble(nengo.LIF(512), 4, label="B")
+nengo.Connection(in_b, b)
 
-model.make_node("Input B", b_in)
-model.make_ensemble("B", nengo.LIF(512), 4)
-model.connect("Input B", "B")
+cconv = CircularConvolution(neurons=nengo.LIF(1032), dimensions=4)
+nengo.DecodedConnection(a, cconv.A)
+nengo.DecodedConnection(b, cconv.B)
 
-cconv = model.add(CircularConvolution(
-    "Circular Convolution", neurons=nengo.LIF(1032), dimensions=4))
-model.connect("A", cconv.A)
-model.connect("B", cconv.B)
+result = nengo.Ensemble(nengo.LIF(512), 4, label="Result")
+nengo.Connection(cconv.output, result)
 
-model.make_ensemble("Result", nengo.LIF(512), 4)
-model.connect(cconv, "Result")
+a_val = nengo.Probe(a, 'decoded_output', filter=0.02)
+b_val = nengo.Probe(b, 'decoded_output', filter=0.02)
+res_val = nengo.Probe(result, 'decoded_output', filter=0.02)
 
-model.probe("A", filter=0.02)
-model.probe("B", filter=0.02)
-model.probe("Result", filter=0.02)
-
-sim = model.simulator()
+sim = nengo.Simulator(model)
 sim.run(0.5)
 
 actual_out = circconv(a_in, b_in)
@@ -47,7 +45,7 @@ def plot(actual, ens):
     plt.gca().xaxis.set_ticks_position('bottom')
     plt.gca().yaxis.tick_right()
     plt.gca().yaxis.set_label_position("right")
-    plt.plot(sim.data(model.t), sim.data(ens))
[…]

### 2013-12-05 scripts/plot_comm_channel.py

@@ -8,25 +8,28 @@ from base import sorted_spikes
 
 model = nengo.Model("Communication Channel")
 
-model.make_node("Input", output=white_noise(1, 5, seed=60))
+in_ = nengo.Node(output=white_noise(1, 5, seed=60))
 
-model.make_ensemble("A", nengo.LIF(30), 1)
-model.make_ensemble("B", nengo.LIF(30), 1)
+a = nengo.Ensemble(nengo.LIF(30), 1, label="A")
+b = nengo.Ensemble(nengo.LIF(30), 1, label="B")
 
-model.connect("Input", "A")
-model.connect("A", "B")
+nengo.Connection(in_, a)
+nengo.DecodedConnection(a, b)
 
-model.probe("Input")
-model.probe("A", filter=0.01)
-model.probe("A.spikes")
-model.probe("B", filter=0.01)
-model.probe("B.spikes")
+in_p = nengo.Probe(in_, 'output')
+a_val = nengo.Probe(a, 'decoded_output', filter=0.01)
+a_spikes = nengo.Probe(a, 'spikes')
+b_val = nengo.Probe(b, 'decoded_output', filter=0.01)
+b_spikes = nengo.Probe(b, 'spikes')
 
-sim = model.simulator()
+sim = nengo.Simulator(model)
 sim.run(1)
 
-a_spikes = sorted_spikes(sim, "A")
-b_spikes = sorted_spikes(sim, "B")
+t = sim.data(model.t_probe)
+sim_a = next(ens for ens in sim.model.objs if ens.label == "A")
+a_spikes = sorted_spikes(t, sim_a, sim.data(a_spikes))
+sim_b = next(ens for ens in sim.model.objs if ens.label == "B")
+b_spikes = sorted_spikes(t, sim_b, sim.data(b_spikes))
 
 def adjust(ax):
     ax.spines['top'].set_visible(False)
@@ -40,7 +43,7 @@ mm_to_inches = 0.0393701
 figsize = (85. * mm_to_inches / 2., (39.31 + 4) * mm_to_inches)
 plt.figure(figsize=figsize)
 ax = plt.subplot(3,1,1)
-ax.plot(sim.data(model.t), sim.data("Input"), color='k')
+ax.plot(t, sim.data(in_p), color='k')
 ax.text(0.5, 1.0, "Input", ha='center', va='center', fontsize='large')
 adjust(ax)
[…]

### 2013-12-05 scripts/plot_lorenz.py

@@ -12,7 +12,7 @@ beta = 8.0/3
 rho = 28
 
 model = nengo.Model('Lorenz attractor')
-model.make_ensemble('State', nengo.LIF(2000), 3, radius=60)
+state = nengo.Ensemble(nengo.LIF(2000), 3, radius=60, label="State")
 
 def feedback(x):
     dx0 = -sigma * x[0] + sigma * x[1]
@@ -23,26 +23,27 @@ def feedback(x):
             dx1 * tau + x[1],
             dx2 * tau + x[2]]
 
-model.connect('State', 'State',
-              function=feedback, filter=tau)
-model.probe('State', filter=tau)
-model.probe('State.spikes')  # Very expensive!!
+nengo.DecodedConnection(state, state, function=feedback, filter=tau)
+st_val = nengo.Probe(state, 'decoded_output', filter=tau)
+st_spikes = nengo.Probe(state, 'spikes')  # Very expensive!!
 
-sim = model.simulator()
+sim = nengo.Simulator(model)
 sim.run(6)
 
-state_spikes = sorted_spikes(sim, "State", iterations=250, every=80)
+sim_state = next(ens for ens in sim.model.objs if ens.label == "State")
+state_spikes = sorted_spikes(sim.data(model.t_probe), sim_state,
+                             sim.data(st_spikes), iterations=250, every=80)
 
 mm_to_inches = 0.0393701
 figsize = (100. * mm_to_inches, 80. * mm_to_inches)
 plt.figure(figsize=figsize)
 ax = plt.subplot2grid((2,3), (0,0), colspan=2, rowspan=2, projection='3d')
-ax.plot(sim.data('State')[:,0], sim.data('State')[:,1], sim.data('State')[:,2],
+ax.plot(sim.data(st_val)[:,0], sim.data(st_val)[:,1], sim.data(st_val)[:,2],
         color='k')
 ax.dist = 9
 
 ax = plt.subplot2grid((2,3), (0,2))
-ax.plot(sim.data(model.t), sim.data('State'))
+ax.plot(sim.data(model.t_probe), sim.data(st_val))
 ax.set_xlim(0, 6)
 ax.set_ylabel("Amplitude")
 ax.spines['top'].set_visible(False)

### 2013-12-05 scripts/plot_nef_summary.py

@@ -9,43 +9,42 @@ from base import sorted_spikes
 ## Encoding
 
 model = nengo.Model("Encoding")
-model.make_node("Input signal", output=lambda t: t * 2 - 1)
-model.probe("Input signal")
+in_ = nengo.Node(output=lambda t: t * 2 - 1)
+in_p = nengo.Probe(in_, 'output')
 
 intercepts = np.linspace(-0.9, 0.9, 8)
 encoders = np.asarray([1,-1,1,-1,1,-1,1,-1])
 intercepts *= encoders
 encoders.shape = (8,1)
-model.make_ensemble("A", nengo.LIF(8), 1, intercepts=intercepts,
-                    max_rates=Uniform(80, 100), encoders=encoders)
-model.connect("Input signal", "A")
-model.probe("A.spikes")
-model.probe("A", filter=0.03)
+a = nengo.Ensemble(nengo.LIF(8), 1, intercepts=intercepts,
+                   max_rates=Uniform(80, 100), encoders=encoders)
+nengo.Connection(in_, a)
+a_spikes = nengo.Probe(a, 'spikes')
+a_val = nengo.Probe(a, 'decoded_output', filter=0.03)
 
-model.make_node("Dummy", output=lambda t: 0)
-cos = model.connect("A", "Dummy", function=np.cos, filter=0.03)
-model.probe(cos)
+dummy = nengo.Node()
+nengo.DecodedConnection(a, dummy, function=np.cos, filter=0.03)
+cos_out = nengo.Probe(dummy, 'output')
 
 ## Encoding plot
 mm_to_inches = 0.0393701
 figsize = (180. * mm_to_inches / 4.0, 180. * mm_to_inches / 2.0)
 
-sim = model.simulator()
+sim = nengo.Simulator(model)
 sim.run(1)
 
-t = sim.data(model.t)
-spikes = [t[sim.data("A.spikes")[:,i] > 0].flatten()
-          for i in xrange(model.get("A").n_neurons)]
-A = sim.model.get("A")
-A.eval_points.sort(axis=0)
-J = np.dot(A.eval_points, A.encoders.T)
-activities = A.neurons.rates(J)
+a = sim.get(a)
+t = sim.data(model.t_probe)
+spikes = [t[sim.data(a_spikes)[:,i] > 0].flatten() for i in xrange(a.n_neurons)]
+a.eval_points.sort(axis=0)
[…]

### 2013-12-18 nengo.tex
Submission finalized, references fixed

@@
 has been done previously
-by PyNN \citep{davison2008}.
+by PyNN \citep{davison2009}.
 In PyNN, the same Python script
@@
 the same functionality
-with Brian \citep{goodman2009}, NEURON \citep{hines2009},
-and NEST \citep{eppler2008}.\footnote{We
+with Brian \citep{goodman2008}, NEURON \citep{hines2009},
+and NEST \citep{eppler2009}.\footnote{We
   were unable to compile PCSim \citep{pecevski2009} on the
@@
 dedicated to building large-scale neural models
-\citep{eppler2008,goodman2009,hines2009},
+\citep{eppler2009,goodman2008,hines2009},
 and many tools for simulating cognitive phenomena
@@
 in terms of software design
-are PyNN \citep{davison2008}
+are PyNN \citep{davison2009}
 and Topographica \citep{bednar2009},

### 2014-03-19 abstract.tex
Added abstract
Just for archival sake!

\documentclass[12pt]{article}

\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{fullpage}
\usepackage{multicol}
\usepackage{natbib}
\usepackage[textwidth=0.8in, shadow]{todonotes}

\setlength{\bibsep}{0.0pt}

% \title{Reimplementing Nengo in Python for speed and simplicity}

\begin{document}

\thispagestyle{empty} % Suppress page number

\begin{centering}
{\large \bf Reimplementing Nengo in Python for speed and simplicity}

\vspace{12pt}
{Trevor Bekolay, James Bergstra, Xuan Choo, Travis DeWolf,
Eric Hunsberger, \\ Daniel Rasmussen, Terrence Stewart, Chris Eliasmith}

{\it Centre for Theoretical Neuroscience, University of Waterloo}

\vspace{12pt}
\end{centering}

  %Nengo has recently been used to build Spaun, 
  Nengo is the neural simulation software package
  recently used to build Spaun,
  the largest functional brain model to date~\cite{spaun}.
  Despite Nengo being open source, it has not been
  widely used or extended. Two reasons
  for this are the complexity of
  the implementation and the speed of larger simulations.
  In this paper, we describe a new version of Nengo
  which has been reimplemented in Python to
  address these problems.

  Nengo was originally a Java application that
  was extended to use Python scripting through Jython,
  a Java implementation of the Python language~\cite{nengo}.
  While this enabled models to be created
  with simpler syntax, 
  the approach had several weaknesses.
  % there were several weaknesses of the approach. 
  The scripting layer
  could not leverage the Python ecosystem
  because most scientific packages call C or Fortran code.
  Nengo developers had to be familiar with both Java and Python.
  The latest stable version of Jython
  lags behind CPython, so language features like
  decorators could not be used until recently.
  The new implementation of Nengo targets CPython,
  and therefore can take advantage of modern language features
  and libraries like NumPy and Matplotlib.
  The API was also redesigned to provide
  a much simpler interface for modelers.
  Currently, Nengo implements the majority
  of the commonly used features of the Java version,
  with an order of magnitude less code
  (previously $\sim$30,000 lines of Java;
  currently $\sim$2000 lines of Python).

  Nengo now leverages the Theano package~\cite{theano}
  to speed up the creation and simulation of
  Nengo models.
  Theano offers a simple interface for
  describing computations on matrices.
  It builds a graph of those computations,
  simplifies wherever possible, and then transparently
  generates and runs optimized C code.
  Theano can target multi-core CPUs, GPUs,
  and other local and remote computing resources
  with minimal modeler effort.
  Preliminary speed tests have shown that simple models
  can be simulated 9--12~times faster
  using a single CPU core compared to using multiple cores
  with the Java version.
  Simulating a leaky integrate-and-fire neuron in Theano
  can be done $\sim$15~times faster on a GPU than
  on a CPU, and so we predict $\sim$100 times speedups
  for some Nengo models using a GPU.

  %Nengo is now built such that
  %it can use multiple backends to construct and simulate
  %neural models. 
  Nengo now accomodates multiple backends 
  to construct and simulate neural models.
  This will allow, for example,
  a backend that generates PyNN scripts~\cite{pynn}.
  Such a backend would enable other modelers to create
  large-scale functional models by using Nengo
  as a ``neural compiler'' that translates mathematical
  models to biologically realistic spiking neuron models
  without requiring them to learn a new simulation framework.

\begingroup
\renewcommand{\section}[2]{}% Removes the References title
{\scriptsize
\begin{multicols}{2}
\bibliographystyle{unsrt}
\bibliography{biblio.bib}
\end{multicols}}
\endgroup

\end{document}
